### 2019-04-22 models/gcm/gcm.Rmd
data go up

---
title: "GCM with Lookup Tables"
author: "Peter Racz"
date: "21/04/2019"
output: html_document
---

```{r echo = F, message = F, warning = F}
knitr::opts_chunk$set(fig.width=10, fig.height=10, fig.path='Figs/', eval=FALSE, echo=TRUE, warning = FALSE, message=FALSE, tidy=TRUE)

setwd('~/Work/NZILBB/RaczBecknerHayPierrehumbert2019/models/gcm')
```

This is the Generalised Context Model [GCM] code for the paper "Morphological convergence as on-line lexical analogy" [from now on: ESP paper].

This md goes through the logic of creating the datasets with the various GCM metrics. But the files pulled in by the grabdata function are in fact the outputs of this md. Beware.

The GCM uses distances between pairs of words and sums them over different domains (e.g. a word's distance to words in class A and class B).

We use two types of distance, edit distance (Levenshtein distance) and word-aligned natural-class based distance [from Adam Albright](http://web.mit.edu/albright/www/software/Alignment.zip).

Adam's perl script finds the best alignment between two words:

| s | p | r | I | N |
|:-:|:-:|:-:|:-:|:-:|
| - | - | r | I | N |

And then uses overlapping natural classes (from Frisch, Pierrehumbert and Broe 2004) to calculate distance between the two strings.

It is not very fast, and we have a lot of comparisons to make. Instead of computing these online every time the GCM is fit, I create a **lookup table** of every combination of target and test items for our experiments and then match values from this table for each model fit using basically glorified sql.

The lookup table looks like this:

| target word | training word | distance |
|:-----------:|:-------------:|:--------:|
| target1     | training1     | dist val |
| target1     | training2     | dist val |
| target2     | training1     | dist val |
| ...         |               |          |

It's long.

```{r }
source('~/Work/NZILBB/RaczBecknerHayPierrehumbert2019/functions/grabfunctions.R')

source('~/Work/NZILBB/RaczBecknerHayPierrehumbert2019/functions/grabdata.R')

lookup = read_csv('~/Work/NZILBB/RaczBecknerHayPierrehumbert2019/models/gcm/feature_distance_lookup_tidy.txt')

```

## 1. Creating the lookup tables

We create all combinations of forms. This is test words x test words / training words. We write the combinations into a file.

```{r create_lookup_table1, eval=F}

# test words
test = baseline.verbs$disc
# training words: verbs in celex in moder classes
training1 = celex %>% 
  filter(!is.na(moder)) %>% 
  select(ahpa) %>% # using Albright and Hayes' transcription
  pull
# training words: other regular verbs (monsyllabic)
c2 = celex %>% 
  filter(
    category == 'regular',
    is.na(moder),
    !is.na(ahpa)
    ) %>% 
  mutate(
    nchar = nchar(ahpa),
    vowels = str_extract_all(ahpa, '[euVoEIi2U{Q34@a6}]'),
    nvowels = map(vowels, ~ length(.))
  ) %>% 
  select(-vowels) %>% 
  unnest()

training2 = c2 %>% 
  filter(
    nvowels == 1
    ) %>% 
  select(ahpa) %>% 
  distinct() %>% 
  pull

c3 = celex %>% 
  filter(
    category == 'irregular',
    is.na(moder),
    !is.na(ahpa)
    ) %>% 
  mutate(
    nchar = nchar(ahpa),
    vowels = str_extract_all(ahpa, '[euVoEIi2U{Q34@a6}]'),
    nvowels = map(vowels, ~ length(.))
  ) %>% 
  select(-vowels) %>% 
  unnest() %>% 
  filter(word != 'forsake', nvowels == 1) # a survivor

training3 = c3$ahpa

training = c(training1, training2, training3, test)

all_combinations = expand.grid(test, training) %>% 
  distinct() %>% 
  filter(Var1 != Var2)

write_delim(all_combinations, 'all_combinations.txt', delim = '\t', col_names = F)
# this takes about twelve hours on my hardware:
# system('bash alignment_runner_ideal.sh > all_combinations_out.txt')

```

We grab the combinations and create a lookup table.

```{r create_lookup_table2, eval=F}
# of course, what I do above is not what I did in practice. in practice, I fit four different sets of words because I was in the process of figuring out what's going on and what I need. The first set had word pairs repeat several times because of a join. Etc.
# see pocketgcm_create_lookup_table.R

# I'm not sharing these files, but you can see the code here to get an idea.

# batch 1
distances1 = read_csv('~/Work/NZILBB/convergence_esp/ESP2019/pocketgcm/Alignment/out.txt', col_names = F)
names(distances1) = 'feature.distance'

pairs1 = read_delim('~/Work/NZILBB/convergence_esp/ESP2019/pocketgcm/Alignment/everybody.txt', '\t', col_names = F)
names(pairs1) = c('target', 'training')

nrow(pairs1)==nrow(distances1)

lookup1 = cbind(distances1, pairs1)
lookup1 = lookup1 %>% distinct() # I made a mistake in setting up the first batch. luckily, I made it for both distances and pairs
names(lookup1)

# batch 2
distances2 = read_csv('~/Work/NZILBB/convergence_esp/ESP2019/pocketgcm/Alignment/out2.txt', col_names = F)
names(distances2) = 'feature.distance'

pairs2 = read_delim('~/Work/NZILBB/convergence_esp/ESP2019/pocketgcm/Alignment/everybody2.txt', '\t', col_names = F)
names(pairs2) = c('target', 'training')

nrow(pairs2)==nrow(distances2)

lookup2 = cbind(distances2, pairs2)
names(lookup2)

# batch 3
distances3 = read_csv('~/Work/NZILBB/convergence_esp/ESP2019/pocketgcm/Alignment/out3.txt', col_names = F)
names(distances3) = 'feature.distance'

pairs3 = read_delim('~/Work/NZILBB/convergence_esp/ESP2019/pocketgcm/Alignment/everybody3.txt', '\t', col_names = F)
names(pairs3) = c('target', 'training')

nrow(pairs3)==nrow(distances3)

lookup3 = cbind(distances3, pairs3)
names(lookup3)

# batch 4
distances4 = read_csv('~/Work/NZILBB/convergence_esp/ESP2019/pocketgcm/Alignment/out4.txt', col_names = F)
names(distances4) = 'feature.distance'

pairs4 = read_delim('~/Work/NZILBB/convergence_esp/ESP2019/pocketgcm/Alignment/everybody4.txt', '\t', col_names = F)
names(pairs4) = c('target', 'training')

nrow(pairs4)==nrow(distances4)

lookup4 = cbind(distances4, pairs4)

lookups_raw = rbind(lookup1,lookup2,lookup3, lookup4)
lookups_raw %>% head

write_csv(lookups_raw, 'feature_distance_lookup_raw.txt')

```

Verbs are categorised according to their phonological schema. We call these categories "Moder" classes, both to honour Carol Lynn Moder who largely influenced them and because I have old code with slightly different categories and I want backwards compatibility.

I'm putting the moder class definitions here for posterity. Note that this bit does not interact with anything else in the code.

```{r moder_class_definitions, eval=F}
c$moder = NA
# DROVE
c$moder = ifelse(
  c$monosyl==T
  & c$class == 'irregular'
  & str_detect(c$present, "[i2][zvdltnk]$") # including "steal"
  & str_detect(c$past, "5[zvdltnk]$"),
    'DROVE', c$moder)
# SANG
c$moder = ifelse(
  c$monosyl==T
  & c$class == 'irregular'
  & str_detect(c$present, "I(m|N|Nk)$")
  & str_detect(c$past, "\\{(m|N|Nk)$"),
    'SANG', c$moder)
# BURNT    
c$moder = ifelse(
  c$monosyl==T
  & c$class == 'irregular'
  & str_detect(c$present, "[3EI][nl]$")
  & str_detect(c$past, "[3EI][nl]t$"),
    'BURNT', c$moder)
# KEPT
c$moder = ifelse(
  c$monosyl==T
  & c$class == 'irregular'
  & str_detect(c$present, "i[lpmn]$")
  & str_detect(c$past, 'E[lpmn]t$'),
    'KEPT', c$moder)
```

We add a couple things to the lookup table, including the Moder category.

```{r create_lookup_table3, eval=F}

lookup = read_csv('feature_distance_lookup_raw.txt')

# we add raw, uncompromised edit distance
lookup$edit.distance = stringdist::stringdist(lookup$target, lookup$training, method = 'lv')

# then, we match up training forms with their verb categories (drove/sang/kept/burnt) and gcm categories (regular/irregular), paying attention to using the A&H transcription

celex2 = celex %>% 
  # filter(!(is.na(moder) & category == 'irregular')) %>%
  filter(!is.na(ahpa)) %>% # celex has more rows than albright & hayes celex. we want the latter.
  mutate( # we want monosyllabic
    nchar = nchar(ahpa),
    vowels = str_extract_all(ahpa, '[euVoEIi2U{Q34@a6}]'),
    nvowels = map(vowels, ~ length(.))
  ) %>% 
  select(-vowels) %>% 
  unnest()

celex2 = celex2 %>% 
  filter(nvowels == 1) %>%  
  select(
    moder, 
    ahpa,
    category
    ) %>% 
  rename(
    training = ahpa,
    moder.training = moder,
    category.training = category
    )

# this is now the following monosyllabic words in the a&H celex:
# - irregular and regular verbs in sang/kept/burnt/drove
# - regular verbs not in a moder class
# - irregular verbs not in a moder class

# some verbs are in celex as both regular and irregular. these verbs are in the lookup only once.
# so we expect the lookup to get longer after the join.
lookup = left_join(lookup, celex2, by = 'training')
# lookup[lookup$training == 'rIN',] # see?

# there are 256 extra rows here, one for each target, that we get rid of for now.
lookup = lookup %>%
  filter(!is.na(training))

# lookup %>% 
  # filter(is.na(moder.training)) %>% View

# then, we match up test forms with their categories.

bv2 = baseline.verbs %>% 
  select(disc, moder) %>% 
  rename(
    target = disc,
    moder.target = moder
  )

lookup = left_join(lookup, bv2, by = 'target')

# a couple things got in here twice, these are mostly regular verbs from celex.
# nrow(lookup)
# nrow(distinct(lookup))
# grr = lookup[duplicated(lookup),] %>% select(training, category.training, moder.training) %>% distinct()

lookup = lookup %>% distinct()

# we need to flag if a training verb is from celex (relevant for baseline) or we made it up (relevant for esp)
target.verbs = unique(lookup$target)
lookup$training.type = ifelse(lookup$training %in% target.verbs, 'esp', 'baseline')

# now we need moder.training for test verbs AS training verbs

# then, we match up test forms AS training forms with their categories.

bv3 = baseline.verbs %>% 
  select(disc, moder) %>% 
  rename(
    training = disc,
    moder.training = moder
  )

lookup_baseline = lookup[lookup$training.type == 'baseline',]
lookup_esp = lookup[lookup$training.type == 'esp',]
# nrow(lookup) == (nrow(lookup_baseline) + nrow(lookup_esp))
lookup_esp$moder.training = NULL
lookup_esp = left_join(lookup_esp, bv3, by = 'training')
# nrow(lookup) == (nrow(lookup_baseline) + nrow(lookup_esp))

# grr = lookup_baseline[lookup_baseline$target == 'blIm',]
# grr %>% count(moder.training, category.training)
# View(grr)
lookup_baseline = lookup_baseline[!(is.na(lookup_baseline$moder.training) & is.na(lookup_baseline$category.training)),]
# grr = lookup_baseline[lookup_baseline$target == 'blIm',]
# grr %>% count(moder.training, category.training) # sounds about right
# View(grr)
# grr = lookup_esp[lookup_esp$target == 'blIm',]
# grr %>% count(moder.training, category.training) # sounds about right too

lookup = rbind(lookup_baseline, lookup_esp)
# let's check one target
# grr = lookup[lookup$target == 'brip',]
# View(grr)
# burnt/drove/kept/sang: irregular/ regular celex + target. regular non-moder verbs from celex.

write_csv(lookup, 'feature_distance_lookup_tidy.txt')
```

Our lookup table has been made. Let's hope it's not the way Joe Pesci was made in Goodfellas.

## 2a. Replicating the Baseline-GCM in the paper.

Training sets are regular / irregular. They consist of verbs in Celex that match the Moder class of the target. Note that one verb can be in both the training and the test sets if it's listed as both a regular and an irregular in Celex.

```{r baseline_gcm1, eval=F}
# training sets
# SANG & 13 & 8 \\
# BURNT & 43 & 6\\
# KEPT & 34 & 12\\
# DROVE & 93 & 15\\

baseline.gcm.paper = baselineGCMpaperSetup(lookup = lookup)

# baseline.gcm.paper[baseline.gcm.paper$target == 'drin',] %>% arrange(category.training) %>% View
# baseline.gcm.paper[baseline.gcm.paper$target == 'brip',] %>% count(moder.training, category.training)
# baseline.gcm.paper[baseline.gcm.paper$target == 'biv',] %>% arrange(category.training) %>% View
# looks about right.

baseline.gcm.paper = doGCM(restricted.set = baseline.gcm.paper, gcm.id = 'baseline_gcm_wf_lookup_moder', var_s = 0.3, var_p = 1)

```

## 2b. Refitting the Baseline-GCM in the paper using a larger regular set.

Reviewer points out that we are tilting the field in the GCM's favour by restricting its training set to verbs that basically look exactly like the verb schemata we are interested in. One of their alternative suggestions is to keep Moder classes but include *all* regular verbs in the regular category.

We implement this below with one restriction: verbs have to be monosyllabic. (Irregular verbs in the set already are. Polysyllabic regular verbs will be very far away from the targets in any case.)

```{r baseline_gcm2, eval=F}
# training sets
# SANG & 13 & 1221 \\
# BURNT & 43 & 1219\\
# KEPT & 34 & 1225\\
# DROVE & 93 & 1228\\

# lookup[is.na(lookup$category.training),] %>% 
#   select(training) %>% 
#   distinct() %>% 
#   View

baseline.gcm.reviewer = baselineGCMreviewerSetup(lookup)

# baseline.gcm.reviewer[baseline.gcm.reviewer$target == 'brip',] %>% count(category.training)  
# baseline.gcm.reviewer[baseline.gcm.reviewer$target == 'zEl',] %>% count(category.training)      
# checks out

baseline.gcm.reviewer = doGCM(restricted.set = baseline.gcm.reviewer, gcm.id = 'baseline_gcm_wf_lookup_reviewer', var_s = 0.3, var_p = 1)

```

Let's check results.

```{r check_baselines, eval=F}
baseline.verbs = left_join(baseline.verbs, baseline.gcm.paper, by = 'disc')
baseline.verbs = left_join(baseline.verbs, baseline.gcm.reviewer, by = 'disc')

# celex_gcm_nof is the edit-distance gcm. 
# it's not implemented here but you could tweak doGCM() to use edit rather than feature distance.
# celex_gcm_wf is from an old run. there are some issues with transcriptions. but you see it is very similar in accuracy.
with(baseline.verbs, cor(reg.av, celex_gcm_nof)) # .558
with(baseline.verbs, cor(reg.av, celex_gcm_wf)) # .545
with(baseline.verbs, cor(reg.av, baseline_gcm_wf_lookup_moder)) # .531
with(baseline.verbs, cor(reg.av, baseline_gcm_wf_lookup_reviewer)) # .455

```

## 2c. Tuning s for the reviewer GCM

We suspect that the drop in accuracy stems from the fact that there are loads more verbs in the regular category now.

We tune the s parameter and use two measures: pearson's rho with baseline averages and index of concordance with individual responses (regular/irregular). We could also fit hierarchical generalised linear models and compare their goodness of fit; but it would take longer.

```{r tune_s_baseline, eval=F}

baseline.verbs = baseline %>% 
  group_by(
    word, 
    disc
    ) %>% 
  summarise(
    reg.av = mean(regular)
  )

# r.val1 = tuneS(lookup = lookup, baseline.verbs = baseline.verbs, var_s = 0.3)
# checks out

var_s_range = c(seq(0.1, 0.9, 0.1), 0.95, 0.99)

r.vals = map(
  var_s_range, ~ tuneS(lookup = lookup, var_s = ., var_p = 1) %>% 
    getR(baseline.verbs = baseline.verbs)
  )

r.vals %>% unlist

C.vals = map(
  var_s_range, ~ tuneS(lookup = lookup, var_s = ., var_p = 1) %>% 
    getC(baseline = baseline)
  )

C.vals %>% unlist

# s = .9 wins based on two metrics. s = .95/.99 seems too high for me, with diminishing returns.

```

We tune *s* way up then, and see what it does to the ESP experiment.

## 2d. We file the baseline value

```{r fit_baseline, eval=F}
from.lexicon = baselineGCMreviewerSetup(lookup = lookup)

baseline.gcm.reviewer = doGCM(restricted.set = from.lexicon, gcm.id = 'baseline_gcm_wf_lookup_reviewer', var_s = 0.9, var_p = 1)

# nrow(baseline)
baseline = left_join(baseline, baseline.gcm.reviewer, by = 'disc')
# nrow(baseline)
baseline.verbs = left_join(baseline.verbs, baseline.gcm.reviewer, by = 'disc')
with(baseline.verbs, cor(reg.av, baseline_gcm_wf_lookup_reviewer)) # .61

```

## 3a. Fitting the reviewer GCM (I hope this catches on) on the ESP experiment

We now have a GCM that's trained on irregular and regular verbs in the target verb's moder class, as well as miscellaneous regular verbs. We tuned it.

We loop through the participants of the esp experiment. For each participant, we filter the lookup table to only contain target forms from test2 and training forms from celex (following the reviewer GCM setup) and esp robot answers with appropriate regular / irregular categories.

```{r esp_fit, eval=F}

my.participant_ids = unique(espdata$participant_id)

my.esp.gcms = map(my.participant_ids, ~ runGuyLookup(espdata = espdata, lookup = lookup, my.participant_id = ., gcm.id = 'esp_gcm_lookup', var_s = 0.9, var_p = 1))

my.esp.gcm.df = do.call('rbind', my.esp.gcms)

# my.esp.gcms2 = map(my.participant_ids, ~ runGuyLookup(espdata = espdata, lookup = lookup, my.participant_id = ., gcm.id = 'esp_gcm_lookup', var_s = 0.3, var_p = 1))
# 
# my.esp.gcm.df2 = do.call('rbind', my.esp.gcms2)
test2 = left_join(test2, my.esp.gcm.df, by = c('disc', 'participant_id'))

```

## 3b. Checking accuracy

```{r eval=F}

# my.esp.gcm.df2 = my.esp.gcm.df2 %>% 
  # rename(esp_gcm_lookup2 = esp_gcm_lookup)
# test2$esp_gcm_lookup2 = NULL
# test2 = left_join(test2, my.esp.gcm.df2, by = c('disc', 'participant_id'))

# esp_gcm_wf had similar issues as celex_gcm_wf: transcriptions were off here and there. you see that the new esp_gcm, where these are resolved, does better.
with(test2, Hmisc::somers2(esp_gcm_wf,resp_post_reg)['C']) # .62
with(test2, Hmisc::somers2(esp_gcm_lookup,resp_post_reg)['C']) # .68
# with(test2, Hmisc::somers2(esp_gcm_lookup2,resp_post_reg)['C']) # .62

test2sum = test2 %>% 
  group_by(
    participant_id,
    reg_rate,
    lex_typicality
  ) %>% 
  summarise(
    mean.resp = mean(na.omit(resp_post_reg)),
    mean.gcm = mean(esp_gcm_lookup)
  )

ggplot(test2sum, aes(mean.resp, mean.gcm, colour = lex_typicality)) +
  geom_point() +
  stat_smooth(method = 'lm') +
  facet_wrap(~ reg_rate)

```

## 4. Time to tidy and save datasets.

```{r eval=F}
# you don't want to run this, the data pulled in by grabdata are its outputs: it will complain about columns missing.

baseline = baseline %>% 
  rename(
    baseline_mgl_features = celex_mgl_wf, # we didn't change the Minimal Generalisation Learner so we keep its predictions
    baseline_mgl_edits = celex_mgl_nof, #...also from its edit-distance run.
    baseline_gcm_edits = celex_gcm_nof,
    baseline_gcm_features = baseline_gcm_wf_lookup_reviewer
  ) %>% 
  select(
    -celex_gcm_wf
  )

write_csv(baseline, 'convergence_paper_baseline_data.txt')

espdata = espdata %>% # we had a heap of various gcm fits. we don't keep them. we only want the ones we report in the paper.
  select(
    -verb_baseline_gcm3_reg,
    -verb_baseline_gcm2_reg,
    -verb_baseline_gcm_reg,
    -verb_esp_gcm_reg,
    -verb_esp_mgl_reg
  )

write_csv(espdata, 'convergence_paper_esp_data.txt')

test2 = test2 %>% 
  select(
    # -esp_gcm_lookup2,
    -celex_mgl_wf,
    -celex_mgl_nof,
    -celex_gcm_wf,
    -celex_gcm_nof,
    -esp_gcm_wf
  ) %>% 
  rename(
    individual_gcm_edits = esp_gcm_nof,
    individual_gcm_features = esp_gcm_lookup,
    individual_mgl_edits = esp_mgl_nof,
    individual_mgl_features = esp_mgl_wf
  )

bsum = baseline %>% 
  select(
    disc,
    baseline_gcm_edits,
    baseline_mgl_edits,
    baseline_gcm_features,
    baseline_mgl_features
    ) %>% 
  distinct()

test2 = left_join(test2, bsum, by = 'disc')

write_csv(test2, 'convergence_paper_esp_test2_predictions.txt')

```

## 5. Check the reviewer critique of moder classes on baseline data

Reviewer points out that the moder classes are inconsistent and makes suggestions.

- KEPT class: "Why are {l,p,n,m} members included, but {v} is excluded?" (see [d]: not added because irregular past tense has to end in [t])
- SANG class: "A similar question could be asked about the SANG class, which includes I->ae before a nasal, but excludes I->ae before other C's (sit~sat) and other V->ae (run-ran)."
- DROVE class: "But then why are rise~rose and weave~wove unified, to the exclusion of choose~chose and break~broke, which also have [oʊ] in the past?  (And, potentially, also wear~wore, which is transcribed as [ɔ] in CELEX, correctly reflecting British English, but seems very close to an 'o-past' in American English.)"

We update moder classes based on these suggestions and refit the model to see if this changes the fit.

```{r moder_class_definitions2, eval=F}

reviewer2Moder = function(celex){

celex = celex %>% 
  mutate(
    moder = case_when(
      str_detect(ahpa, '[e]') ~ 'DROVE',
      word == 'run' ~ 'SANG',
      word == 'leave' ~ 'KEPT',
      T ~ moder
      )
  )

return(celex)
}

# celex %>% filter(str_detect(ahpa, 'e')) %>% View
# celex %>% filter(str_detect(ahpa, 'n$')) %>% View
# celex %>% filter(str_detect(ahpa, 'iv$')) %>% View

```

And now we go through the baseline data again.

```{r definitions_testing, eval=F}
celex2 = celex %>% 
  reviewer2Moder()

celex2 = celex2 %>% 
  select(ahpa, moder) %>% 
  rename(
    training = ahpa,
    moder.training = moder
  ) %>% 
  filter(!is.na(training))

lookup_a = lookup %>% filter(training %in% celex2$training)
lookup_b = lookup %>% filter(!training %in% celex2$training)
# nrow(lookup) == nrow(lookup_a) + nrow(lookup_b)
lookup_a = lookup_a %>% 
  select(
    -moder.training
  ) #%>%

lookup_a = left_join(lookup_a, celex2, by = 'training')
lookup2 = rbind(lookup_a, lookup_b)

bgcm2 = baselineGCMreviewerSetup(lookup2)

# bgcm2[bgcm2$target == 'brin',] %>% View
# bgcm2[bgcm2$target == 'brin',] %>% count(moder.training, category.training) # ya

bgcm2 = doGCM(restricted.set = bgcm2, gcm.id = 'rev2gcm', var_s = 0.9, var_p = 1)

baseline$rev2gcm = NULL
baseline = left_join(baseline, bgcm2, by = 'disc')

with(baseline, Hmisc::somers2(baseline_gcm_features,regular)['C']) # .58
with(baseline, Hmisc::somers2(rev2gcm,regular)['C']) # .59

baseline.verbs = left_join(baseline.verbs, bgcm2, by = 'disc')
with(baseline.verbs, cor(reg.av, baseline_gcm_wf_lookup_reviewer)) # .61
with(baseline.verbs, cor(reg.av, rev2gcm)) # .69

```

There is a numeric improvement, more so in pearson's rho than in C. Is it robust? We fit two mixed models, one with weights from the GCM based on the reported moder classes, one with weights from a GCM based on the updated classes. We try to predict trial responses, grouped under participants. We scale the two predictors. We don't want to have the two predictors in one model, as they are very collinear.

```{r definitions_testing2, eval=F}
library(itsadug)

str(baseline %>% select(regular, baseline_gcm_features, rev2gcm, subject))
baseline$reg.resp = as.numeric(baseline$regular)
baseline$subject = as.factor(baseline$subject)

baseline = baseline %>% 
  mutate(
    c.baseline_gcm_features = scales::rescale(baseline_gcm_features),
    c.rev2gcm = scales::rescale(rev2gcm)
  )

fit1 = bam(reg.resp ~ c.baseline_gcm_features + s(subject, bs = 're'), data = baseline, family = binomial, method = 'ML')

fit2 = bam(reg.resp ~ c.rev2gcm + s(subject, bs = 're'), data = baseline, family = binomial, method = 'ML')

compareML(fit1, fit2)

# small difference in ML scores.

```

The difference in model ML scores is small, so I assume that m2 does not provide a significantly better fit on the baseline data than m1. This is not to say that critiques of the consistency of the moder classes are invalid, but it suggests that these might not make a major difference.

### 2019-04-22 models/paper/convergence_paper_models.Rmd
data go up

---
title: "Stats for the convergence paper"
author: "Peter Racz"
date: "21/04/2019"
output: html_document
---

```{r echo = F, message = F, warning = F}
knitr::opts_chunk$set(fig.width=10, fig.height=10, fig.path='Figs/', eval=FALSE, echo=TRUE, warning = FALSE, message=FALSE, tidy=TRUE) # eval = FALSE and how

library(brms)
library(lme4)
library(tidyverse)
options(mc.cores=parallel::detectCores())
# options(mc.cores=1)
Sys.setenv(TZ="Europe/Rome") # a szivemben örök tavasz van

setwd('~/Work/NZILBB/RaczBecknerHayPierrehumbert2019/models/paper')
```

These are the models and stats for the convergence paper, going through it by section number.

```{r }
source('~/Work/NZILBB/RaczBecknerHayPierrehumbert2019/functions/grabdata.R')

```

## Section 4.2

```{r }
baseline %>% group_by(word) %>% summarise(regular = mean(regular)) %>% filter(word == 'splive') 
```

## Section 4.3

```{r }
espdata %>% select(participant_id,lex_typicality,reg_rate) %>% unique %>% group_by(lex_typicality,reg_rate) %>% summarise(n = n()) 
```

## Section 5

Fits on what people are doing across conditions in test2. No algorithmic learning here.

Our weapon of choice is typially lme4. But it won't converge with full random effects structure. So we turn to brms.

```{r }
test2$lex_typicality = factor(test2$lex_typicality, levels = c('typical', 'random', 'atypical'))
test2$reg_rate = factor(test2$reg_rate, levels = c('-40%', 'nc', '+40%'))
test2$baseline_av_z = scales::rescale(test2$baseline_av)
test2$pre_reg_av_z = scales::rescale(test2$pre_reg_av)

fit1 = brm(resp_post_reg ~ 1 + baseline_av_z * lex_typicality + pre_reg_av_z + reg_rate +
    (1 | participant_id) + 
    (1 | word), 
  data = test2, 
  family = bernoulli, 
  control=list(adapt_delta=0.99, stepsize = 0.01, max_treedepth =15))

fit1 = add_criterion(fit1, criterion = 'loo')

save(fit1, file = '~/Work/NZILBB/RaczBecknerHayPierrehumbert2019/models/paper/bayes_fit1.rda')

#####################################################

fit2 = brm(resp_post_reg ~ 1 + baseline_av_z * lex_typicality + pre_reg_av_z + reg_rate +
    (1 + baseline_av_z | participant_id) + 
    (1 | word), 
  data = test2, 
  family = bernoulli, 
  control=list(adapt_delta=0.99, stepsize = 0.01, max_treedepth =15))

fit2 = add_criterion(fit2, criterion = 'loo')

save(fit2, file = '~/Work/NZILBB/RaczBecknerHayPierrehumbert2019/models/paper/bayes_fit2.rda')

#####################################################

fit3 = brm(resp_post_reg ~ 1 + baseline_av_z * lex_typicality + pre_reg_av_z + reg_rate +
    (1 + baseline_av_z | participant_id) + 
    (1 + pre_reg_av_z | word), 
  data = test2, 
  family = bernoulli, 
  control=list(adapt_delta=0.99, stepsize = 0.01, max_treedepth =15))

fit3 = add_criterion(fit3, criterion = 'loo')

# fit3b = add_criterion(fit3b, criterion = 'loo')

save(fit3, file = '~/Work/NZILBB/RaczBecknerHayPierrehumbert2019/models/paper/bayes_fit3.rda')

#####################################################

fit4 = brm(resp_post_reg ~ 1 + baseline_av_z * lex_typicality + pre_reg_av_z + reg_rate +
    (1 + baseline_av_z | participant_id) + 
    (1 + pre_reg_av_z + lex_typicality + reg_rate | word), 
  data = test2, 
  family = bernoulli, 
  control=list(adapt_delta=0.99, stepsize = 0.01, max_treedepth =15))

fit4 = add_criterion(fit4, criterion = 'loo')

save(fit4, file = '~/Work/NZILBB/RaczBecknerHayPierrehumbert2019/models/paper/bayes_fit4.rda')

#####################################################

loo.comparisons = loo_compare(fit1,fit2,fit3,fit4)
save(loo.comparisons, file = '~/Work/NZILBB/RaczBecknerHayPierrehumbert2019/models/paper/loo_comparisons.rda')

options(mc.cores=1)

#####################################################
```

```{r }

load('~/Work/NZILBB/RaczBecknerHayPierrehumbert2019/models/paper/bayes_fit3.rda')
bayes_factor()

load('~/Work/NZILBB/RaczBecknerHayPierrehumbert2019/models/paper/loo_comparisons.rda')
loo.comparisons

# https://arxiv.org/abs/1507.04544 s5.2
```

We definitely want some random slopes (fit1 is worse than the rest). Other than that, the various constellations do not improve the model's expected log pointwise predictive density.

It looks like the model reported in the paper does not ignore too much relevant variation in the random effects.

## Bayes factors!

```{r eval=F}
# random vs atypical

test2b = test2 %>% 
  filter(
    lex_typicality != 'typical'
  )

fit3b = brm(resp_post_reg ~ 1 + baseline_av_z * lex_typicality + pre_reg_av_z + reg_rate +
    (1 + baseline_av_z | participant_id) + 
    (1 + pre_reg_av_z | word), 
  data = test2b, 
  family = bernoulli, 
  control=list(adapt_delta=0.99, stepsize = 0.01, max_treedepth =15), save_all_pars = T)

fit3c = brm(resp_post_reg ~ 1 + baseline_av_z + pre_reg_av_z + reg_rate +
    (1 + baseline_av_z | participant_id) + 
    (1 + pre_reg_av_z | word), 
  data = test2b, 
  family = bernoulli, 
  control=list(adapt_delta=0.99, stepsize = 0.01, max_treedepth =15), save_all_pars = T)

save(fit3b, file = 'fit3b.rda')
save(fit3c, file = 'fit3c.rda')

```

```{r }
load('fit3b.rda')
load('fit3c.rda')

bf = bayes_factor(fit3c, fit3b)
bf$bf
# time to ask for help.
```

## Section 6

```{r }
baseline$r.baseline_gcm_features = scales::rescale(baseline$baseline_gcm_features, to = c(0,1))
baseline$r.baseline_mgl_features = scales::rescale(baseline$baseline_mgl_features, to = c(0,1))

baseline %>% select(word,r.baseline_gcm_features,baseline_gcm_edits) %>% filter(word == 'splive') %>% unique
baseline %>% select(word,r.baseline_mgl_features,baseline_mgl_edits) %>% filter(word == 'splive') %>% unique

```

## Section 6.3

```{r }
# don't want to load Hmisc, messes with select()
with(baseline, Hmisc::somers2(baseline_gcm_features,regular)['C']) # .581

with(baseline, Hmisc::somers2(baseline_mgl_features,regular)['C']) # .586
```

## jumping to Appendix E!

We scale gcm and mgl scores for the regression models. We residualise to see if the baseline gcm and mgl explain variation not explained by the other one.

```{r eval=F}

baseline = baseline %>% 
  mutate(
    c.baseline_gcm_features = scales::rescale(baseline_gcm_features),
    c.baseline_mgl_features = scales::rescale(baseline_mgl_features)
  )

baseline_sum = baseline %>% 
  select(word,baseline_gcm_features,c.baseline_gcm_features,baseline_mgl_features,c.baseline_mgl_features) %>% 
  unique

with(baseline_sum, cor(baseline_gcm_features,baseline_mgl_features))
with(baseline_sum, cor(c.baseline_gcm_features,c.baseline_mgl_features))

m1 <- lm(c.baseline_gcm_features ~ c.baseline_mgl_features, data = baseline_sum)
m2 <- lm(c.baseline_mgl_features ~ c.baseline_gcm_features, data = baseline_sum)

summary(m1)
summary(m2)

baseline_sum$gcm_residual <- residuals(m1)
baseline_sum$mgl_residual <- residuals(m2)

baseline_sum = baseline_sum %>% 
  select(gcm_residual,mgl_residual,word)

baseline2 = left_join(baseline,baseline_sum, by = 'word')

m3 = glmer(regular ~ c.baseline_mgl_features + gcm_residual + (1 | subject) + (1|word), data = baseline2, family = binomial (link = logit), control=glmerControl(optimizer="bobyqa"))

save(m3, file = '~/Work/NZILBB/RaczBecknerHayPierrehumbert2019/models/paper/m3.rda')

m4 = glmer(regular ~ c.baseline_gcm_features + mgl_residual + (1 | subject) + (1|word), data = baseline2, family = binomial (link = logit), control=glmerControl(optimizer="bobyqa"))

save(m4, file = '~/Work/NZILBB/RaczBecknerHayPierrehumbert2019/models/paper/m4.rda')
```

```{r }

load('~/Work/NZILBB/RaczBecknerHayPierrehumbert2019/models/paper/m4.rda')
load('~/Work/NZILBB/RaczBecknerHayPierrehumbert2019/models/paper/m3.rda')

summary(m3)
#                         Estimate Std. Error z value Pr(>|z|)    
# (Intercept)             -0.44918    0.11202  -4.010 6.08e-05 ***
# c.baseline_mgl_features  1.27846    0.07525  16.989  < 2e-16 ***
# gcm_residual             1.37767    0.17171   8.023 1.03e-15 ***
summary(m4)
#                         Estimate Std. Error z value Pr(>|z|)    
# (Intercept)             -1.12780    0.14259  -7.909 2.58e-15 ***
# c.baseline_gcm_features  2.29014    0.15126  15.140  < 2e-16 ***
# mgl_residual             0.95005    0.08549  11.113  < 2e-16 ***
```

They do. The abstract generalisations and the word-based comparisons explain non-overlapping domains of variation in participant responses.

## Section 6.4

## jumping to Appendix F; Table 12

```{r }

with(baseline, Hmisc::somers2(baseline_gcm_features,regular)['C']) # .58
with(baseline, Hmisc::somers2(baseline_mgl_features,regular)['C']) # .59

with(test2, Hmisc::somers2(baseline_gcm_features,resp_post_reg)['C']) # .6
with(test2, Hmisc::somers2(individual_gcm_features,resp_post_reg)['C']) # .68

with(test2, Hmisc::somers2(baseline_mgl_features,resp_post_reg)['C']) # .61
with(test2, Hmisc::somers2(individual_mgl_features,resp_post_reg)['C']) # .55
```

## jumping to Appendix F.2

Now we check, for GCM/MGL, if the individual model explains variation beyond the baseline model in the esp test2.

```{r eval=F}
test2 = test2 %>% 
  mutate(
    c.baseline_gcm_features = scales::rescale(baseline_gcm_features),
    c.baseline_mgl_features = scales::rescale(baseline_mgl_features),
    c.individual_gcm_features = scales::rescale(individual_gcm_features),
    c.individual_mgl_features = scales::rescale(individual_mgl_features),
  )

test2$gcm_wf_score_diff <- test2$c.individual_gcm_features - test2$c.baseline_gcm_features
test2$mgl_wf_score_diff <- test2$c.individual_mgl_features - test2$c.baseline_mgl_features

m5 <- glmer(resp_post_reg ~ c.baseline_gcm_features + gcm_wf_score_diff + pre_reg_av + (1 + c.baseline_gcm_features + gcm_wf_score_diff | participant_id) + (1 | word), data = test2, family = binomial (link = logit), control=glmerControl(optimizer="bobyqa"))
save(m5, file = '~/Work/NZILBB/RaczBecknerHayPierrehumbert2019/models/paper/m5.rda')

m6 <- glmer(resp_post_reg ~ c.baseline_mgl_features + mgl_wf_score_diff + pre_reg_av + (1 + c.baseline_mgl_features + mgl_wf_score_diff | participant_id) + (1 | word), data = test2, family = binomial (link = logit), control=glmerControl(optimizer="bobyqa"))
save(m6, file = '~/Work/NZILBB/RaczBecknerHayPierrehumbert2019/models/paper/m6.rda')

# Third

m7 <- glmer(resp_post_reg ~ c.individual_gcm_features + c.baseline_mgl_features + pre_reg_av + (1 + c.individual_gcm_features + c.baseline_mgl_features | participant_id) + (1 | word), data = test2, family = binomial (link = logit), control=glmerControl(optimizer="bobyqa"))
save(m7, file = '~/Work/NZILBB/RaczBecknerHayPierrehumbert2019/models/paper/m7.rda')

```

```{r }
load('~/Work/NZILBB/RaczBecknerHayPierrehumbert2019/models/paper/m5.rda')
load('~/Work/NZILBB/RaczBecknerHayPierrehumbert2019/models/paper/m6.rda')
load('~/Work/NZILBB/RaczBecknerHayPierrehumbert2019/models/paper/m7.rda')

summary(m5)
#                  Estimate Std. Error z value Pr(>|z|)    
# (Intercept)              -5.5677     0.3724 -14.949  < 2e-16 ***
# c.baseline_gcm_features   4.3687     0.4249  10.281  < 2e-16 ***
# gcm_wf_score_diff         2.6117     0.4696   5.562 2.67e-08 ***
# pre_reg_av                5.2887     0.6022   8.782  < 2e-16 ***
summary(m6)
#                        Estimate Std. Error z value Pr(>|z|)    
# (Intercept)              -5.0641     0.3374 -15.010   <2e-16 ***
# c.baseline_mgl_features   2.0335     0.2079   9.782   <2e-16 ***
# mgl_wf_score_diff         0.2085     0.1723   1.210    0.226    
# pre_reg_av                7.4492     0.5996  12.425   <2e-16 ***
summary(m7)
#                      Estimate Std. Error z value Pr(>|z|)    
# (Intercept)                -5.6326     0.3364 -16.743  < 2e-16 ***
# c.individual_gcm_features   2.7857     0.3776   7.378 1.61e-13 ***
# c.baseline_mgl_features     1.5998     0.1709   9.361  < 2e-16 ***
# pre_reg_av                  5.4374     0.5842   9.307  < 2e-16 ***
summary(m7)$coef %>% round(3) %>% xtable
# vif.mer(m7)
summary(rePCA(m7))

test2 %>% select(word,moder,verb_baseline_mgl_reg) %>% filter(moder == 'KEPT' & verb_baseline_mgl_reg == 1) %>% unique %>% arrange(word) %>% select(word) %>% pull %>% length
# 28, not 57
```

It looks like the best model has the predictions of the individual gcm and the baseline mgl.

## Marginal and conditional r squared for the two models

```{r eval=F}
load('esp_fit.rda')
formula(fit)
MuMIn::r.squaredGLMM(fit) # 0.27 0.45
MuMIn::r.squaredGLMM(m7) # 0.22 0.47 
test2$m7.pred = predict(m7, type = 'response')
with(test2, Hmisc::somers2(m7.pred, resp_post_reg)) # .84
```

## Figure 4

```{r }
# baseline, post-test, type
# type: means, gcm, mgl
b1 = baseline %>% 
  group_by(
    word,
    baseline_gcm_features,
    baseline_mgl_features
    ) %>% 
  summarise(reg.av = mean(regular)) %>% 
  ungroup()

b1 = b1 %>% 
  mutate(
    baseline_gcm_features = scales::rescale(baseline_gcm_features, to = c(0,1)),
    baseline_mgl_features = scales::rescale(baseline_mgl_features, to = c(0,1)),
    reg.av = scales::rescale(reg.av, to = c(0,1))
  )

t2 = test2 %>% 
  group_by(
    word, 
    lex_typicality
  ) %>% 
  summarise(
    post.reg.av = mean(resp_post_reg),
    individual_gcm_features = mean(individual_gcm_features),
    individual_mgl_features = mean(individual_mgl_features),
  ) %>% 
  ungroup() %>% 
  mutate(
    individual_gcm_features = scales::rescale(individual_gcm_features, to = c(0,1)),
    individual_mgl_features = scales::rescale(individual_mgl_features, to = c(0,1)),
    post.reg.av = scales::rescale(post.reg.av, to = c(0,1))
  )

t3 = t2 %>% left_join(b1, by = 'word') %>% 
  rename(
    individual.mgl = individual_mgl_features,
    individual.gcm = individual_gcm_features,
    baseline.mgl = baseline_mgl_features,
    baseline.gcm = baseline_gcm_features
  )

t3la = t3 %>% 
  select(
    word,
    lex_typicality,
    baseline.gcm,
    baseline.mgl,
    reg.av
  ) %>% 
  rename(
    GCM = baseline.gcm,
    MGL = baseline.mgl,
    participants = reg.av
  ) %>% 
  gather(type, baseline.averages, -word, -lex_typicality)

t3lb = t3 %>% 
  select(
    word,
    lex_typicality,
    individual.gcm,
    individual.mgl,
    post.reg.av
  ) %>% 
  rename(
    GCM = individual.gcm,
    MGL = individual.mgl,
    participants = post.reg.av
  ) %>% 
  gather(type, individual.averages, -word, -lex_typicality)

t4 = full_join(t3la,t3lb, by = c("word", "lex_typicality", "type"))

t4$type = factor(t4$type, levels = c('participants', 'GCM', 'MGL'))

ggplot(t4, aes(x = baseline.averages, y = individual.averages, colour = lex_typicality, shape = lex_typicality)) +
  guides(size = F) +
  geom_point() +
  geom_smooth(se = F, method = 'loess', formula = 'y ~ x') +
  facet_wrap( ~ type)
ggsave('interactionplot.pdf', device = cairo_pdf, width = 12.5, height = 4)

```

### 2019-04-22 functions/grabdata.R

@@ -1,0 +1,40 @@
+# training set
+celex = read_csv('~/Work/NZILBB/RaczBecknerHayPierrehumbert2019/data/celex_verbs_with_categories_csv_fixed_moder.txt')
+
+celex = celex %>% 
+  select(
+    present.o,
+    present,
+    moder,
+    regular
+  ) %>% 
+  rename(
+    word = present.o,
+    disc.old = present
+  ) %>%
+  mutate(
+    category = case_when(
+      regular == 'Y' ~ 'regular',
+      regular == 'N' ~ 'irregular'
+    )
+  )  %>% 
+  filter(!is.na(category))
+
+# fixing transcription
+
+transcription.matcher = read_csv('~/Work/NZILBB/RaczBecknerHayPierrehumbert2019/data/transcription_matcher.txt')
+
+celex = left_join(celex, transcription.matcher)
+
+celex[celex$disc == 'riN',]
+
+# test set
+baseline = read_csv('~/Work/NZILBB/RaczBecknerHayPierrehumbert2019/data/convergence_paper_baseline_data.txt')
+
+
+# esp data
+espdata = read_csv('~/Work/NZILBB/RaczBecknerHayPierrehumbert2019/data/convergence_paper_esp_data.txt')
+
+# esp only post-test with gcm weights
+test2 = read_csv('~/Work/NZILBB/RaczBecknerHayPierrehumbert2019/data/convergence_paper_esp_test2_predictions.txt')
+

### 2019-04-22 functions/grabfunctions.R

@@ -1,0 +1,198 @@
+# get similarity from whatever distance you choose to use
+getSimilarity = function(feature.distance, var_s, var_p){#the def values in the paper
+  pairwise.similarity = exp ( -feature.distance / var_s )^var_p 
+  return(pairwise.similarity)
+}
+
+# set up GCM training thing as in paper
+
+baselineGCMpaperSetup = function(lookup){
+# first thing we need to do is restrict training set to moder verbs in celex.
+  target.verbs = unique(lookup$target)
+
+  baseline.gcm.paper = lookup %>% 
+    filter(
+      !training %in% target.verbs,
+      moder.target == moder.training # we only keep comparisons to training verbs in the same moder class
+    )
+
+return(baseline.gcm.paper)
+}
+
+# fit GCM where regular class has ordinary regular verbs
+baselineGCMreviewerSetup = function(lookup){
+
+  target.verbs = unique(lookup$target)
+
+  baseline.gcm.reviewer1 = lookup %>% 
+    filter(
+      !training %in% target.verbs, # no need for these
+      moder.training == moder.target # we compare target verbs to real verbs in their category...
+    )
+
+  baseline.gcm.reviewer2 = lookup %>% 
+    filter(
+      is.na(moder.training) & category.training == 'regular'
+    ) %>% 
+    mutate(
+      moder.training = moder.target # each "other" regular occurs with each target. all we need to do is flag it as belonging to the target's class.
+    )
+
+  baseline.gcm.reviewer = rbind(baseline.gcm.reviewer1, baseline.gcm.reviewer2)
+  
+return(baseline.gcm.reviewer)
+}
+
+# fit GCM on custom lookup table made by previous function
+# for each verb class
+#   for each target
+#     sum over total
+#     sum over regular
+#     sum over irregular
+#     get similarity to class
+
+doGCM = function(restricted.set,gcm.id, var_s, var_p){
+# this function expects a lookup table where each target is matched with all relevant training words.
+# training words are regular or irregular.
+  restricted.set = restricted.set %>% 
+    mutate(
+      f.pairwise.similarity = getSimilarity(feature.distance, var_s = var_s, var_p = var_p)
+    )
+  
+  restricted.set = restricted.set %>% 
+    group_by(target) %>% 
[…]

### 2019-04-22 models/mgl/mgl_processor.R

@@ -1,0 +1,66 @@
+# code to process an MGL sum file
+
+library(dplyr)
+library(stringr)
+library(knitr)
+
+
+handles <- read.delim("256verbs_print_forms.txt", sep=",") %>% select(base.form,category,base.print,regular.form,irregular.form) %>% rename(present.form = base.form)
+# get formatting right
+handles$irregular.form <- sub("5","o",handles$irregular.form)
+handles$regular.form <- sub("Id","@d",handles$regular.form)
+
+
+getMGLRules <- function(output){
+output2 <- output
+
+output2$past.form <- sub("»", "", output2$V5)
+output2$present.form <- sub("»", "", output2$V3)
+
+output2$alternation <- paste(as.character(output2$V7),as.character(output2$V8),as.character(output2$V9), sep="")
+output2$Pres <- output2$V12
+output2$Pfeat <- output2$V13
+output2$environment <- paste(as.character(output2$V14),as.character(output2$V15),as.character(output$V16), sep = ' ')
+output2$Qfeat <- output2$V17
+output2$Qres <- output2$V18
+output2$scope <- output2$V19
+output2$hits <- output2$V20
+# output2$rule2 <- paste(as.character(output2$rule),as.character(output2$structural_description), sep= ' / ')
+output2$reliability <- as.numeric(as.character(output2$V21))
+output2$confidence <- as.numeric(as.character(output2$V22))
+output2$related_forms <- output2$V23
+output2$exceptions <- output2$V24
+
+# output2$past.form <- output2$V5
+# output2$present.form <- output2$V3
+# It does the thing where it inserts an onset-rhyme boundary marker
+
+output3 <- output2[,c("present.form","past.form","alternation","Pres","Pfeat","environment","Qfeat","Qres","scope","hits","reliability","confidence","related_forms","exceptions")]
+output3 <- output3[-1,]
+
+output3 <- output3[order(-output3$confidence),]
+
+foc.dat <- merge(output3,handles)
+
+foc.dat <- foc.dat[order(-foc.dat$confidence),]
+foc.dat <- foc.dat[order(foc.dat$present.form),]
+
+foc.irreg.rules <- subset(foc.dat, as.character(past.form) == as.character(irregular.form))
+foc.reg.rules <- subset(foc.dat, as.character(past.form) == as.character(regular.form))
[…]

### 2019-05-28 models/paper/convergence_paper_models.Rmd
fixing paths

@@
 
-setwd('~/Work/NZILBB/RaczBecknerHayPierrehumbert2019/models/paper')
+try(setwd('../../models/paper'))
+try(setwd('~/Github/RaczBecknerHayPierrehumbert2019/models/paper'))
 ```
@@
 ```{r }
-source('~/Work/NZILBB/RaczBecknerHayPierrehumbert2019/functions/grabdata.R')
+source('../../functions/grabdata.R')
 
@@
 
-save(fit1, file = '~/Work/NZILBB/RaczBecknerHayPierrehumbert2019/models/paper/bayes_fit1.rda')
+save(fit1, file = '../../models/paper/bayes_fit1.rda')
 
@@
 
-save(fit2, file = '~/Work/NZILBB/RaczBecknerHayPierrehumbert2019/models/paper/bayes_fit2.rda')
+save(fit2, file = '../../models/paper/bayes_fit2.rda')
 
@@
 
-save(fit3, file = '~/Work/NZILBB/RaczBecknerHayPierrehumbert2019/models/paper/bayes_fit3.rda')
+save(fit3, file = '../../models/paper/bayes_fit3.rda')
 
@@
 
-save(fit4, file = '~/Work/NZILBB/RaczBecknerHayPierrehumbert2019/models/paper/bayes_fit4.rda')
+save(fit4, file = '../../models/paper/bayes_fit4.rda')
 
@@
 loo.comparisons = loo_compare(fit1,fit2,fit3,fit4)
-save(loo.comparisons, file = '~/Work/NZILBB/RaczBecknerHayPierrehumbert2019/models/paper/loo_comparisons.rda')
+save(loo.comparisons, file = '../../models/paper/loo_comparisons.rda')
 
@@
 
-load('~/Work/NZILBB/RaczBecknerHayPierrehumbert2019/models/paper/bayes_fit3.rda')
+load('../../models/paper/bayes_fit3.rda')
 bayes_factor()
 
-load('~/Work/NZILBB/RaczBecknerHayPierrehumbert2019/models/paper/loo_comparisons.rda')
+load('../../models/paper/loo_comparisons.rda')
 loo.comparisons
@@
 
-save(m3, file = '~/Work/NZILBB/RaczBecknerHayPierrehumbert2019/models/paper/m3.rda')
+save(m3, file = '../../models/paper/m3.rda')
 
@@
 
-save(m4, file = '~/Work/NZILBB/RaczBecknerHayPierrehumbert2019/models/paper/m4.rda')
+save(m4, file = '../../models/paper/m4.rda')
 ```
@@
 
-load('~/Work/NZILBB/RaczBecknerHayPierrehumbert2019/models/paper/m4.rda')
-load('~/Work/NZILBB/RaczBecknerHayPierrehumbert2019/models/paper/m3.rda')
+load('../../models/paper/m4.rda')
+load('../../models/paper/m3.rda')
 
@@
 m5 <- glmer(resp_post_reg ~ c.baseline_gcm_features + gcm_wf_score_diff + pre_reg_av + (1 + c.baseline_gcm_features + gcm_wf_score_diff | participant_id) + (1 | word), data = test2, family = binomial (link = logit), control=glmerControl(optimizer="bobyqa"))
-save(m5, file = '~/Work/NZILBB/RaczBecknerHayPierrehumbert2019/models/paper/m5.rda')
+save(m5, file = '../../models/paper/m5.rda')
 
 m6 <- glmer(resp_post_reg ~ c.baseline_mgl_features + mgl_wf_score_diff + pre_reg_av + (1 + c.baseline_mgl_features + mgl_wf_score_diff | participant_id) + (1 | word), data = test2, family = binomial (link = logit), control=glmerControl(optimizer="bobyqa"))
-save(m6, file = '~/Work/NZILBB/RaczBecknerHayPierrehumbert2019/models/paper/m6.rda')
+save(m6, file = '../../models/paper/m6.rda')
 
@@
 m7 <- glmer(resp_post_reg ~ c.individual_gcm_features + c.baseline_mgl_features + pre_reg_av + (1 + c.individual_gcm_features + c.baseline_mgl_features | participant_id) + (1 | word), data = test2, family = binomial (link = logit), control=glmerControl(optimizer="bobyqa"))
-save(m7, file = '~/Work/NZILBB/RaczBecknerHayPierrehumbert2019/models/paper/m7.rda')
+save(m7, file = '../../models/paper/m7.rda')
 
@@
 ```{r }
-load('~/Work/NZILBB/RaczBecknerHayPierrehumbert2019/models/paper/m5.rda')
-load('~/Work/NZILBB/RaczBecknerHayPierrehumbert2019/models/paper/m6.rda')
-load('~/Work/NZILBB/RaczBecknerHayPierrehumbert2019/models/paper/m7.rda')
+load('../../models/paper/m5.rda')
+load('../../models/paper/m6.rda')
+load('../../models/paper/m7.rda')
 
@@
 t2 = test2 %>%
+  mutate(lex_typicality2 = 
+           case_when(
+             lex_typicality == 'atypical' ~ 'reversed',
+             lex_typicality == 'typical' ~ 'typical',
+             lex_typicality == 'random' ~ 'random',
+           ),
+         lex_typicality2 = factor(lex_typicality2, levels = c('typical', 'reversed', 'random'))
+         ) %>% 
   group_by(
     word, 
-    lex_typicality
+    lex_typicality2
   ) %>% 
@@
     word,
-    lex_typicality,
+    lex_typicality2,
     baseline.gcm,
@@
   ) %>% 
-  gather(type, baseline.averages, -word, -lex_typicality)
+  gather(type, baseline.averages, -word, -lex_typicality2)
 
@@
     word,
-    lex_typicality,
+    lex_typicality2,
     individual.gcm,
@@
   ) %>% 
-  gather(type, individual.averages, -word, -lex_typicality)
+  gather(type, individual.averages, -word, -lex_typicality2)
 
-t4 = full_join(t3la,t3lb, by = c("word", "lex_typicality", "type"))
+t4 = full_join(t3la,t3lb, by = c("word", "lex_typicality2", "type"))
 
@@
 
-ggplot(t4, aes(x = baseline.averages, y = individual.averages, colour = lex_typicality, shape = lex_typicality)) +
+ggplot(t4, aes(x = baseline.averages, y = individual.averages, colour = lex_typicality2, shape = lex_typicality2)) +
   guides(size = F) +
@@
   geom_smooth(se = F, method = 'loess', formula = 'y ~ x') +
-  facet_wrap( ~ type)
-ggsave('interactionplot.pdf', device = cairo_pdf, width = 12.5, height = 4)
+  facet_wrap( ~ type) +
+  scale_shape_discrete(name="Lexical\ntypicality") +
+  scale_colour_discrete(name="Lexical\ntypicality") +
+  xlab('verb baseline average') +
+  ylab('verb ESP average')
+# ggsave('interactionplot.pdf', device = cairo_pdf, width = 12.5, height = 4)
+ggsave('fig4_2019.pdf', width = 12.5, height = 4)
 

### 2019-05-28 functions/grabdata.R

@@ -1,5 +1,11 @@
+try(setwd('~/Work/NZILBB/RaczBecknerHayPierrehumbert2019'))
+try(setwd('~/Github/RaczBecknerHayPierrehumbert2019'))
+
+
 # training set
-celex = read_csv('~/Work/NZILBB/RaczBecknerHayPierrehumbert2019/data/celex_verbs_with_categories_csv_fixed_moder.txt')
+
+
+celex = read_csv('data/celex_verbs_with_categories_csv_fixed_moder.txt')
 
 celex = celex %>% 
   select(
@@ -22,19 +28,19 @@ celex = celex %>%
 
 # fixing transcription
 
-transcription.matcher = read_csv('~/Work/NZILBB/RaczBecknerHayPierrehumbert2019/data/transcription_matcher.txt')
+transcription.matcher = read_csv('data/transcription_matcher.txt')
 
 celex = left_join(celex, transcription.matcher)
 
 celex[celex$disc == 'riN',]
 
 # test set
-baseline = read_csv('~/Work/NZILBB/RaczBecknerHayPierrehumbert2019/data/convergence_paper_baseline_data.txt')
+baseline = read_csv('data/convergence_paper_baseline_data.txt')
 
 
 # esp data
-espdata = read_csv('~/Work/NZILBB/RaczBecknerHayPierrehumbert2019/data/convergence_paper_esp_data.txt')
+espdata = read_csv('data/convergence_paper_esp_data.txt')
 
 # esp only post-test with gcm weights
-test2 = read_csv('~/Work/NZILBB/RaczBecknerHayPierrehumbert2019/data/convergence_paper_esp_test2_predictions.txt')
+test2 = read_csv('data/convergence_paper_esp_test2_predictions.txt')
 

### 2019-07-01 ll_response_to_reviewers_2.txt
huh

Dear Pavel

We would like to submit the revised version of our ms ID 18-ES-2560.R1, titled `Not all indexical cues are equal:  Differential sensitivity to dimensions of indexical meaning in an artificial language' to Language Learning.

The word "social" was changed to "indexical" in the title. This is in line with revisions in the previous round, emphasising that the paper considers distinctions that are too simple to be considered properly "social".

In this round, in response to reviewer comments, we have reworked the introductory sections of the paper, completely rewritten the methods section, and reran the statistical analyses and correspondingly rewritten the results section as well.

We removed our "research questions" to focus on the main hypotheses, added a section to spell out the rationale of the current study, and merged the discussion and general discussion sections.

Changes in the revised text are in blue -- a blue section title indicates that a complete section has been entirely rewritten.

Below, we respond to your point-by-point summary of the reviews and then go on to respond to the reviewers in turn.

> 1. As discussed by all reviewers, improve the flow and the logic of the writing in the introductory section of your manuscript and especially with respect to your research questions. The reviewers indicated that (a) the literature review did not prepare the reader adequately to understand your study, (b) your research questions were unclear, and (c) the overall logic of literature review was hard to follow.

We changed the introductory and background sections in three major ways. First, we set up the previous and the current study immediately after the starting paragraph to provide context to what follows. Second, we removed the "research questions" and instead set up the introduction + background to lead up to the rationale for the current study and, subsequently, to the main hypotheses. Third, we wrote a number of linking paragraphs in the background section. These (hopefully) spell out why we need to overview specific domains of category learning and language variation and how results from these fields inform our study.

> 2. As particularly discussed by Reviewers 1 and 5, please ensure that the manuscript reports full methods, materials, and procedure information that would make it a standalone publication and would not require the reader to first consult your previously published work. And again, do provide a clear, coherent account of how the current dataset extends and builds on your previous publication. This information should ideally be used to motivate the current study.

We now discuss the previous study three times: First, we set it up immediately at the beginning to motivate the current study. Second, we discuss its results in detail at the end of the background section to segue into the rationale for our current study. Third, we return to it in the Discussion in order to highlight what the conversation on contextual language learning can gain from the present study.

We were rather alarmed that Reviewer 5 had to revisit the 2017 paper to understand the design of the current one. This was clearly not our intent. We now used the methods section of the 2017 paper as a basis to remodel our methods section.

> 3. Address carefully the Design issues raised by Reviewer 5. I will leave it up to you whether you would wish to remove any data from analysis (point 5 in this reviewer’s feedback), but I strongly encourage you to consider using Levenshtein’s distance between images as a predictor in your analyses (point 6 in this reviewer’s feedback).

We adopted every single suggestion by Reviewer 5 -- we discuss these changes one by one in our reply to this reviewer.

> 4. Most importantly, you must address all statistical/analytical concerns expressed by Reviewer 5 (under Issues with statistical modelling). I find all of these concerns reasonable, and would like to reiterate that reporting effect size values and confidence intervals is mandatory for us (point 5 under statistical/analytical concerns by Reviewer 5).

Model comparisons are now reported in the methods section and estimated 95% confidence intervals are reported in the results section. We hope that this framing is now sufficient to provide enough information to the reader without resorting to p values.

The data analysis section might be overly verbose at this point. For each model, it lists all relevant fits along with the statistics of how well they fit the data. One alternative could be to shorten this by reporting on how comparisons were made and what the precise criteria were for picking a best model. Another alternative would be to shorten the data analysis section this way and report all the comparisons in an online-only appendix. We would leave this for the Editor to decide.

> 5. Finally, address all remaining issues by all reviewers.

Please find below.

Reviewer: 1
Recommendation: Minor Revision
Comments to the Author:

> My first issue with the introduction is that the first research question (“Can we reliably associate any non-linguistic context whatsoever with linguistic constructions?”) still feels too broad considering the question the experiment answers. What exactly is the answer? Do we conclude that for 18% of participants, at least, the answer is yes? (But then, what if there are harder contexts to learn than orientation?) Or do we conclude that, because most participants failed to learn the association with orientation, that it can’t be reliably associated with linguistic constructions? (But then, most participants failed on gender too, and we already know that that can be associated with linguistic constructions.) The second part of the same question (“If not, do our prior expectations regarding what contexts are generally socially relevant play into whether such an association can be established?”)  also seems slightly problematic. I’d agree that participants presumably arrive with prior expectations as to what contexts are generally socially relevant, but that isn’t actually measured here; it’s simply inferred based on what humans seem to associate linguistic constructions to in natural language. The hypotheses given on page 10 seem a much better way to set up the goals of the study.

The reviewer has a point here. Our intention was to set up a broad field of interest using research questions and then narrow this down using hypotheses. It seems that the research questions do not particularly aid in setting up the broad field of interest and instead confuse readers.

Instead, we now start the paper with the current study, and then very explicitly state that, in order for the reader to follow why this is interesting or how this works, we need to set up a number of previous results from specific domains. We then do this, in a way that is more verbose. This segues into a new section, the rationale for the current study, which concludes with the hypotheses. We explicitly revisit the hypotheses in the results section and the discussion.

> My other issue is more minor and concerned with how the introduction flows. It’s generally much better, but could still be improved. Two examples stand out. First, the section on contextual learning segues from discussion of context aiding memory to context aiding prediction, to the involvement of context in learning, to the role of context in processing. That’s fine in itself, but the transitions are not quite seamless. For instance, it’s not at all clear in the text how the Qian et al. (2014) study is really a non-linguistic equivalent of the immediately preceding studies. The sociolinguistic learning section has a related but slightly different problem: It feels like a somewhat disjointed list of studies.

This reviewer is not alone in finding the introduction disjointed. We wrote a number of linking paragraphs that hopefully clarify why we want to talk about a certain thing and how that relates to previous things. We also hope that, as a result, the reader will not feel that details are missing, because the new flow makes the levels of resolution clearer. For instance, we expanded the sociolinguistic learning section to explain how, in our narrative, one study follows from the other. The section now sets up more clearly that broad contextual variation can be learned in the lab, and that variation based on gender is picked up in these tasks by both children and adults. This leads us to our 2017 study, which, in turn, leads to our present study.

> The Method and Results sections continue to suffer from problems with clarity.  Part of the problem, I think, is that it feels as if there are two many paragraph breaks, which contributes to a disjointed feel. But there are other problems of unclarity. For instance, on page 10, the conversation partner is introduced in a way that takes the reader by surprise. It would be clearer if the paragraph said something like, “In the first, training phase, a participant has to progress on a path, learning a simple artificial language from [IN ITALICS] conversation partners met along the way. In this language, suffixes that represent a grammatical relation (either the diminutive or the plural) are contextually cued: they depend on the conversation partner, present in each trial.” (Another issue here, incidentally – what precisely constitutes a trial?) There are a few cases like this. I feel as if the Method section would benefit from being reread by someone entirely unfamiliar with the study.  The same problems continue in the Results section, which I found hard to follow. I would suggest the same remedy.

We have completely rewritten the Methods and Results sections. The Methods section now starts with a description of one specific experimental setup, ignoring the rest of the details, which are added later. The results section is now simplified -- a large amount of detail is moved to the data analysis section, and the main results and how these survive additional scrutiny are spelt out more clearly at the end.

> Minor issues (many of which contribute to clarity issues mentioned above):
> 1. There is a formatting error in the abstract (? instead of, I assume, dashes)

Corrected.

> 2. The present tense should be replaced by the past tense in many places throughout the manuscript. Its use in reporting published work (e.g., “Docherty, Langstrof, and Foulkes (2013) investigate…”) is common but explicitly discouraged in APA style.  Its use in describing experimental method or reporting results is a little odder and slightly jarring.

We re-tensed the paper to simple past. The one exception is the first description of the experiment, where the past tense seemed overly distracting.

> 3. Page 6: “Speakers of different dialects will focus on and learn different linguistic details (Cohn, Ham, & Podesva, 1999), and speaker awareness of contextual information on all levels can be very imprecise”. This sentence seems to require a “but” not an “and”, as there appears to be a contrast between the two halves. The “and” thus creates a slightly jarring effect.

The reviewer is right that this is not quite a logical relation: speakers learn different things. at the same time, speakers are imprecise. We corrected this to "while speaker awareness..."

> 4. Page 6: “While these factors will have complex social and linguistic representations, they will remain robust and socially salient”. Is there a source for this?

Upon revisiting, this sentence feels like a red herring; in the end, all factors have complex social and linguistic representations, while the previous sentence cites sources on the salience of these specific factors. We removed the sentence.

> 5. Page 9: A conjunction is missing at least twice (in the first sentence and in the sentence beginning “It utilises…”)

Corrected.

> 6. Page 9: “adults […] discard contexts that are deemed irrelevant.” By whom? By the adults themselves? I’m also not sure “deem” is the right term here, since (to me at least) it implies conscious deliberation.

This is a good point -- we removed "deem" (as we make no claims about deeming in the preceding literature review).

> 7. Page 10: The distinction between “socially salient” and “accidental” seems like a category error. Can’t accidental things be socially salient (and can’t non-accidental things be insalient)?

This is indeed unclear. We replaced "accidental" with "irrelevant".

> 8. Why is Figure 3 referenced before Figure 1?

Corrected.

> 9. Page 12: “we repeated two conditions using plural instead of diminutive as the morphological category.” This needs to be clarified! “Repeat” can mean a lot of things. On a related note, the asterisk in Table 2 needs to be explained in the caption.

We added a clarificatory sentence here: "we replicated two conditions using plural instead of diminutive as the morphological category. In these replicated conditions, participants perform the exact same task except that diminutive images are replaced with plural ones." Also clarified the caption.

> 10. Page 15: “Learning extends to generalisation” seems an odd way to express this.

True. All we need to claim is that learning "is followed by" generalisation, so we corrected to this phrasing.

> 11. Page 18: “Training is faster with age, ethnicity, and the fastest with gender.” There seems to be a missing conjunction here again.

This sentence has been removed during the ms' reorganisation.

> 12. Page 18: “it seems that ignoring the view cue is about as difficult as ignoring the gender cue. I think this would be clearer if something were added along the lines of, “In spite of the differences between the different cues when they were main cues…”

This sentence has been removed during the ms' reorganisation. Hopefully the reworked Results section is easier to follow.

> 13. The reasoning on pages 21–22 starting with “Second, the overall…” is hard to follow and could do with being clarified.

This sentence, too, has been removed during the ms' reorganisation.

> 14. Page 22: I think  “derivative morphology” should be “derivational morphology”. I wonder if perhaps this distinction isn’t in danger of being language-specific. I am unaware of any language in which plural suffixes are derivational and diminutives are inflectional, but this may simply reflect my ignorance. In Welsh, for instance, equivalent forms to passive participles in other languages are formed by a clearly derivational affix, which seems to surprise some people. There are also languages that don’t have inflectional number marking; do some of these have derivational markers used in some cases?
15. Page 23: “for example, in French, incorrect marking of the gender of the addressee in a possessive construction will lead to incomprehension.” Perhaps I am misunderstanding what is being claimed about French, but what I imagine from this example is the contrast between “ton [masc.] cheval” and “ta [f.] voiture”; but this distinction depends on the gender of the possessed, not the addressee.

We believe that a major take-away from the Reviewer's point here is that any discussions of morphology will remain tenuous at best, given that the study itself barely touches upon typological questions. What we want to ultimately say is that the task works with at least two different morphological contrasts (not only with the diminutive but also with the plural), much like what we find in our 2017 study. So we removed the discussion of inflectional/derivational morphology.

> 16. Christiansen and Kirby (2003) was written before the experimental turn in language evolution, and thus seems clearly the wrong citation on page 26. The authors might like to consider (for instance): 
> * Scott-Phillips, Thomas C., and Simon Kirby. "Language evolution in the laboratory." Trends in cognitive sciences 14.9 (2010): 411-417.
> * Roberts, Gareth. "The linguist’s Drosophila: Experiments in language change." Linguistics Vanguard 3.1 (2017).

We removed the Christiansen and Kirby citation and instead cite these two works (as we should have in the first place).

Reviewer: 5
Recommendation: Major Revision
Comments to the Author:

> Since I am not an expert in sociolinguistic, I will leave it to the other reviewer(s) to comment on whether the revised manuscript addresses the theoretical issues (i.e., the discussion of the relevant literature, the type of sociolinguistic variation targeted in the study, etc.) raised by the previous round of reviews. My immediate feeling is that not all of the points and requests were fully addressed, and that further revision may be needed to address especially the following issues: (1) the research questions and hypotheses are still not entirely clear, and could be streamlined further; (2) I find the logic of the study still hard to follow at times; (3) the discussion could still do a better job at discussing the limitation of the study and possible alternative interpretations of the results.

We thank the Reviewer for their scrutiny. We adopted all their suggestions and discuss the resulting changes in turn.

> 1)      In general, the paper has a feeling of “unfinishedness” to it. This is especially true of the Introduction and General Discussion, but it also permeates the Methods and Results section. Readers of Language Learning that are not entirely familiar with the reviewed literature (like me) could benefit from a few more details on the studies that are presented (e.g., on p. 5 and 6). Expanding on some of the background literature and on the hypotheses and design of the current study, as well and working on the general cohesiveness of the narrative, would make the submission much stronger.

Following similar comments from all reviewers in this round, the ms has been completely reworked up to the methods section. Now, it introduces the study and its 2017 precedent immediately, explains which literature has to be discussed to set up the premises for the study, provides more links between the various studies to show how they inform the same questions, and adds a rationale for the current study, leading up to the hypotheses tested. We also reorganised the Discussion and the General Discussion, which now start with a summary of the results in the context of the starting hypotheses, follows with how this connects to the 2017 study, and then goes on to speculate a bit on how the results should be interpreted. While we think that speculation has its place in a discussion, we have to admit that the lack of structure in the last version of this ms made speculation overly salient and rampant. Hopefully the rewrites rectify this issue.

> 2)      I think that the entire Methods section (p. 10-17) is quite difficult to follow. Despite several rereads, I had to rely on the Methods section from Rácz, Hay, and Pierrehumbert (2017), in order to fully understand the procedure. I think that the authors should rework the methods section drawing from their 2017 paper, which does a much better job at explaining the experimental task. I like the procedure and I think the design is for the most part quite elegant, but it gets somehow penalized by the poor structure of the section. A (quite extensive) rewrite of the Methods section should easily solve the problem. As also pointed out by Reviewer 1 (first round of reviews), I feel that the authors could still be more explicit in how this study expands on their 2017 paper both methodologically and theoretically.

We found it alarming that the Reviewer had to go back to the 2017 paper to understand what we are doing here. This clearly was not our intent. We have rewritten considerable parts of the methods section. It now starts by setting up a singular case of the design (gender versus spatial orientation), and all the additional details (the use of additional conversation partners, the precise materials, etc) are set up later. We hope that this makes the design easier to follow.

> 3)      APA-related issue: Figures and Tables are not referenced in the right order in the text. Either fix the text or change the order of pictures (e.g., Fig. 3 is mentioned before Fig. 1 and 2, see p. 11) and tables accordingly.

Corrected.

> 4)      Given that there is only one experiment in the study, I do not think we need both a Discussion and a General Discussion. Please merge the two sections together.

Corrected. We clipped the "discussion" and merged it with the "general discussion"

> Design:
5)      I think it is problematic that diminutives and plurals were not represented in the material in equal measure. Plural suffixes were only tested on a subset of the participants because the diminutive is a more “common and iconic pattern that is easier to interpret visually” (I personally believe that the singular-plural distinction is also quite easy to represent visually). Unfortunately, this decision seems a bit a priori, and I cannot help but think that using an equal ratio of diminutive and plural suffixes would have made for a stronger design. I am aware that fixing this issue would require running additional experiments, and I do not think this should be the case here. However, a less painful option that the authors could consider is to remove the plural suffix trials from the data and only focus on the diminutives, as this would make for a cleaner presentation of both hypotheses and results.

We followed the Reviewer's suggestion and now only report the diminutive data in detail in the results section.

We make a separate point about plurals versus diminutives: In our 2017 paper, we show that participant accuracy across the gender / view distinction is similar with plural and with diminutive suffixes. Here we point to this result and add that the same is the case for the gender / ethnicity distinction.

In the data analysis section, we provide the model comparison that supports this conclusion.

> 6)      I also have a slight issue with the visuals of the conversation partners (Fig. 3). It seems to me that the characters in the first set (row 1 and 2) are more visually homogeneous than characters in the second set (row 3 and 4). That is, while it is easy to categorize the characters in the top two rows as differing on three characteristics (gender, age, orientation), there is a lot more going on in the bottom rows (different hair color, different clothing, and in general a number of small details such as creases on dresses and buttons on shirts, which are not there in the first set). So, the bottom set is more visually clattered than the top, and I am wondering whether this may be affecting performance (which would explain e.g., the weaker effect of ethnicity). One way of addressing this problem would be to re-run the models that predict response accuracy based on the Levenshtein distance between images, but this time adding the different pictures as predictors in the model (could be either fixed or random effect; also note that this only works if the Levenshtein distance is calculated on the color version of the images (e.g., RGB values) rather than after reducing the images to grey scale values, since much of the “visual clutter” is driven by the colors, I think).

We reran the comparison of perceptual versus conceptual distance using the model suggested by the reviewer. We also moved the entire discussion of distances to the data analysis section and only point to it towards the end of the results section to reduce clutter. We provide the model comparison that supports our conclusion, namely that conceptual distance outweights perceptual distance in accounting for participant success in the test trials.

We respectfully disagree with the reviewer's suggestion about using coloured images as a basis of comparison. We would bring two arguments for this. First, hosting the setup on Mechanical Turk means that we have no control over the "equipment": participants' screens can look like anything, and the resulting colour spectra are likely highly variable. What will then remain constant is the colour contrasts. Second, the hue distinctions are fairly visible in the black and white images as well (perhaps even more so). These aspects together incline us to use black and white distances.

> Issues with statistical modelling (pages 16-22):
Although the authors have in part addressed the comments from the first round of reviews (the analyses are now more consistent with the hypotheses of the study), some major issues remain:

> 1)      The specifications of the statistical models (both linear and binomial regressions) are not clearly stated in the text. Rather, the structures of the models are almost only inferable by looking at the output of the model presented in tables 4 and 6. When working with linear mixed-effects model, and especially when using the lme4 package in R, it is customary to report the full specification of the models that have been run, either in the form of a mathematical formula, or using R syntax. For instance, p. 17: “For the test model, we considered interactions of across-participant and within-participant predictors. Relevant random slopes were also tested”. Given what I have from p. 15 and Table 4 (across-participant predictors are participant Gender and Age, within-participant predictors are Cue[View, Gender, Ethnicity, Age], Pattern, and Familiarity), I imagine that the structure of the model must have been something along these lines: Correct ~ GenderP+AgeP+(Cue*Familiarity)+(1 | Participant): that is, accuracy is predicted by the main effect of across-participant variables and the interaction effects between Cue and Familiarity, with random intercepts for participants. However, this information is not provided to me, therefore I cannot be sure that this is indeed the correct model. Moreover, the text references testing “relevant random slopes” (i.e., 1 + ????? | Participant), but no information is provided on which variables given as random slopes for participants, and there is no way to infer them from the model output.

The Reviewer raises a number of important issues in points 1-5 that we respond to under 5.

> 2)      Page 17: “We used goodness-of-fit testing (with a significance threshold of p < 0.05) and the Akaike Information Criterion for model selection and reported the best model”. I like the use of model selection, but we need to know which models have been tested, and what the outcome of the test was for the individual models.

> 3)      As the authors also point out on p. 16, count data are not normally distributed, nor are their residuals. I appreciate the authors effort to keep the modelling simple and clear, but using a linear regression here is not ideal and can possibly have generated spurious results. The authors should instead fit the data using a Poisson distribution with a log link function, which expresses the probability of count data (this is easily doable in lme4 using glmr() and specifying family as Poisson and link as log). Given that 24 was the minimum N of trials, it may even be an idea to rescale the count data, so that 24 = 0 (in this way the data would be counting the N of *additional* trials needed on top of the minimum possible N). Beware that using the log link will rescale the outcome variable, meaning that the Estimate values in the output table will not be in the original scale (0-1 probability) anymore (you can get back to the original scale using the inverted log function).

> 4)      This is probably the biggest issue with the paper: the statistical models reported in the paper have very little to none use when it comes to hypothesis testing. No p-values nor confidence intervals or effect size values are reported, and marginal effect plots are only reported for one specific contrast (Cue*Familiarity in one specific model (Fig. 6). Although I am sympathetic with the idea of moving away from null-hypothesis significance testing, we unfortunately cannot draw any conclusions about the hypotheses of the study from the reported data. Unless I am missing something, the authors base their conclusions exclusively on Estimate values (but we have no information about the size/significance of these effects) and on visual inspection of the plot (but these plots are based on the raw data, and not on the statistical models, and therefore give no information about the nature of the effects – plotting the fitted data would have made matters slightly better). The authors do mention the use of model selection (e.g., on p. 6: “… the gender effect seen in training is not supported by model comparison”), which can be used for hypothesis testing, but these results are not reported anywhere. Therefore, I am not sure of what to do of any of the results as they are currently reported. I would recommend that the authors report p-values for all predictor estimates, for instance by using the “lmerTest” package (Kuznetsova, Brockhoff, & Christensen, 2017). If the authors would rather use model comparison, then the different models (as well as the test results) should be reported explicitly (see point 2 above).

> 5)       As a consequence of the issues expressed in point 4, the paper in its current form does not conform to the Statistical Guidelines of the journal: neither confidence intervals nor effect size values are reported. This should of course be corrected.

We considerably reworked the data analysis and results sections in response to the above comments. The changes are as follows:

- For each model discussed in the results section, the data analysis section now lists the specific model formulae (along with random effects) fit in the process of model comparison, as well as the models' goodness-of-fit statistics. So, for instance: in the data analysis, we say that in order to see whether X has an effect, we tested (i) "outcome ~ X + Y" and (ii) "outcome ~ Y". These models have the following goodness-of-fit statistics. This suggests that (i) provides a significantly better fit. This is what we report in the results. In the results, we report (i) and provide 95% CI-s for X and Y. In other cases, we do not report the model in the results in detail, we only point to it in the results. (along the logic of "X is not relevant. see data analysis.")
- Beside the formulae, outcomes and predictors are spelt out in the running text in the data analysis section.
- We removed a separate discussion of training results from the paper. This might seem a major change in how results are presented, but the training results are overall in line with the test results. Instead, we now consider training success in modelling the test data (see below).
- Where we report estimates in the results section, we now provide 95% Wald confidence intervals and give a relatively simple explanation on how they should be interpreted. Hopefully this brings us in line with the statistical guidelines of the journal without reporting p values for predictor estimates.

> 6)      I am a bit puzzled by the inclusion of participant gender and age as fixed effects in the model, since the authors do not have any specific hypotheses about the possible effect of these variables on performance. If they do, then these hypotheses should be stated explicitly on page 15. If they, on the other hand, are only interested in controlling for gender and age differences in the sample, then I am quite convinced that it would suffice having participants as a random effect in the model (as it is already the case, if I understand the model structure correctly).

The Reviewer is right. We by and large did not anticipate gender or age effects and struggled considerably in accounting for them. It is entirely likely that participant age and gender are very relevant aspects of the learning process investigated in tasks of this type, but, if that is the case, this should perhaps be assigned to future research.

> 7)      It seems very likely that the length of training (N of trials needed by the participants to complete the training phase) would be a very strong predictor of performance in the test phase. The authors state this themselves on page 18: “A participant that finished training faster will be more accurate in test (est = -0.03, se = 0.0005, using binomial generalized linear regression)”. However, I cannot seem to find a trace of this main effect anywhere — at least not in Table 6, where I would expect it to be. Without this information, any effect of Cues and Familiarity may be irrelevant, since the effect of training may turn out to explain most of the variance in test performance. Please fix.

We now include training trial counts in our model of the test data, as spelt out in the data analysis section. Participants who are faster in training are also better in test, but the effects of cue type and familiarity remain relevant beyond this. Of course, cue type in training and participant speed in training are strongly correlated, but the variance inflation factors of the model are not high enough for us to worry that our results are artefacts of model fitting.

> 8)      Page 15: “Since we do assume that training is longer in certain across-participant conditions than others, we took every condition separately and then removed the slowest 2.5% based on training trial count threshold”. I do not think this is necessary when using regression models. I would suggest that the authors re-run the analyses without filtering out the slowest 2.5% and see whether they get different results.

If we rerun the best model of the test data reported in the paper using all participants, we find that the results remain very similar -- the z values are all in the same ballpark.

Our reason for excluding outliers is predominantly ontological: given the paradigm (Mechanical Turk), it is clear to us that not all participants paid the same amount of attention, and while our cutoff is an arbitrary (albeit established) one, it allows us to trim those who tried the least. This is the justification for using the same threshold in our 2017 study. To put it in another way: it is true that no participant takes anywhere as long as they would if they picked training trial responses truly at random, but it is still likely that some of them got past training by poor rote learning and sheer force of will, and our outlier thershold helps us remove some of them.

> 9)      Page 19: “3. One, small group have learnt the rule but in reverse…”. Where is this information coming from? This does not seem to be the case by looking at Fig. 5 – there are a very few data points below 25%.

This is a distracting comment that does not really build up anything, so we removed it. Incidentally, there are people who seem to have learnt the rule in reverse, and they are now visible in the figures (see next point), but the fact that they exist is probably statistically not remarkable.

> 10)     Minor comment about Fig. 4 and 5: I like the use of the rug plot to show the raw data points on the y-axis, however it is quite uninformative in its current form, since a lot of the points are overlapping. Consider lowering the alpha value of the geom_rug() in ggplot in order to better show the amount of overlap between data points.

We reworked the plots so that they jitter participants in the violin of the given cue that they trained with. The plots now show both distribution of averages and proportions of participants who "figured it out" versus those who keep guessing in test.

> 11)     Fig. 6: The lack of error bars makes it hard to determine the strength/significance of the difference between familiar and unfamiliar items in relation to Age and Ethnicity. Please fix.

We reworked this plot as well, showing changes for each individual between familiar and novel conversation partners. In addition, we now report confidence intervals for these interactions in the main model.

> 12)     P. 15, second to last paragraph: Strictly speaking it is wrong to list training vs. test as a within-participant factor here, since you are modeling the two phases separately (i.e., considering the two phases as two separate experiments).

This is true. Corrected.

> Minor issues:
> 1)      There are a couple of literal quotes in the Introduction, e.g. on p. 4 and 5. The authors should consider rephrasing these in their own words to make the text flow better.

We kept one literal quote -- here we use a side quote that is not the main conclusion of a paper in a somewhat polemic way, and the verbatim reflects that. We rephrased the others.

> 2)      P. 3, top paragraph: “Non-linguistic contexts also play a crucial role in early…”. This phrase ends a bit abruptly.

Rephrased to "Non-linguistic contexts interact with other aspects of language use: they play a crucial role in early word learning..."

> 3)      P. 7, third paragraph (“The mechanisms through … Chevrot & Foulkes, 2013) is a bit obscure, please expand if relevant, otherwise delete.

This we think is an important point, so we expanded it further: "Indeed, modern theories of the mental lexicon, the storage of linguistic forms, tend to argue that non-linguistic and linguistic contextual information plays a crucial role in how forms are stored and processed, with effects on a range of phenomena from speech perception (Johnson 1997) to priming (De Vaab et al 2007)."

> 4)      P. 9, The Current Study, first paragraph (“Previous research … experimental setting”): References to previous research should appear here.

We now add citations to this paragraph that cross-reference it with work discussed in the previous section.

> 5)      P. 10, second to last paragraph (“Such allomorphies are common across languages … Hungarian”): It would be nice with some examples here, especially for people that do not work with allomorphy.

We reconsidered the discussion of types of allomorphy and came to the conclusion that any discussions of morphology here will remain tenuous at best, given that the study itself barely touches upon typological questions. What we want to ultimately say is that the task works with at least two different morphological contrasts (not only with the diminutive but also with the plural), much like what we find in our 2017 study. So we removed the discussion of morphology.

> 6)      P. 11, second to last sentence (“It is easy…”): awkward phrasing, please rewrite.

We rewrote this section, making the sentence disappear.

> 7)      P. 13: “Choosing a suffix depends on whether … they face is irrelevant” is redundant, consider deleting.

Corrected -- removed the sentence.

> 8)      P. 14: “The platform has been …  Berinsky, Huber, & Lenz, 2012” can be deleted, people have now been using Amazon Mechanical Turk in research for years (I believe that this was probably added in response to a comment by Reviewer 2).

To us, it seems that reader reactions to AMT data are bimodal. Some take it as a perfectly acceptable design decision, while others are more wary. Clearly, Reviewer 5 belongs in the former, Reviewer 2 in the latter category. For the wary reader, we think it is good to include a few references of this sort, if only to show that their epistemological concerns are shared by the authors and that a substantial body of literature exists that investigates precisely these concerns.

> 9)      Same page, “A participant who gets … servers might be unreliable”: also unnecessary, please delete.

Removed.

> 10)     P. 15, “0.72*(24^2.07)~518 trials”: where is this coming from? What are 0.72 and 2.07? I must be missing something here.

We agree that this is distracting, and expanding it would distract further. The point we want to make here is that we checked how long it takes to finish training guessing by the use of simulations, and we now say this in the test.

> 11)     P. 16, top paragraph (“Figure 3 … different individuals”): This could be removed from here and added as caption to Table 2 instead.

Edited.

> 12)     P. 19, top, “However, participants are not normally distributed within each condition”): You mean that average scores are not normally distributed? Were you expecting them to be?

This was a clunky sentence and we removed it during the rewriting of this section. We were in fact surprised that participant test accuracy is so clearly bimodal, especially given that training trial counts are much more normally distributed. We now discuss participant distributions in test more clearly in the results section.

> 13)     P. 22, line 14: “socially irrelevant” -> “less socially relevant” (see point 3 under “General issues”).

Corrected.

Reviewer: 3
Recommendation: Minor Revision
Comments to the Author:

> I think that there may still be a little bit of work guiding the reader at the start however: I understand that it's normal to discuss the literature review before the details of the study itself, but I think the introduction might cover them slightly more as otherwise it's not immediately clear to the reader how some of the literature is going to be relevant.

We reorganised the sections leading up to the Methods. We now set up the current study right at the beginning, then spell out what aspects of human contextual learning need to be discussed to lead up to the hypotheses. Then we review the literature, with added linking paragraphs to explain how the various results belong together. Finally, we lead up to work on contextual language learing, our 2017 study, and how the current paper expands on that study, leading to the hypotheses.

> The results section is clearer, but there are still a lot of tests one after the other and in places I wasn't sure what was being tested and why anymore, so here too a little bit more guidance would be good.

We rewrote the data analysis and results sections. A lot of detail has been moved to the data analysis section. The results section is more terse on certain details that are less relevant. It is also more repetitive. Hopefully this helps bringing the main points of the paper across.

> There are still a few typos and sentences that don't read clearly here and there (e.g. p 4 line 3, it should be a comma, not a full stop ... if not. do our prior...) and the tenses go back and forth in places between past and present (mostly where I'd expect it to stick to the past, e.g. p 14 where the participants are being discussed)

We went through the paper and enforced the past tense, as per APA guidelines. The one exception is the first description of the experiment, where the past tense seemed overly distracting.

> A couple minor points:
> - p .11 figure 3 is mentioned before figure 1 - should the numbering be changed?

Corrected.

> - p 23 French marking of the gender of the addressee for possession - maybe include an example, I couldn't work out what this would be (ton, ta? but they're not marked according to the addressee's gender)

This example was not set up correctly: we replaced it with an example of French adjectives, which are marked by the gender of the modified nouns (contrasting "Je suis désolé" and "Je suis désolée").

### 2019-07-01 models/paper/convergence_paper_models.Rmd
huh

@@
 test2$lex_typicality = factor(test2$lex_typicality, levels = c('typical', 'random', 'atypical'))
-test2$reg_rate = factor(test2$reg_rate, levels = c('-40%', 'nc', '+40%'))
+test2$reg_rate = factor(test2$reg_rate, levels = c('nc', '-40%', '+40%'))
 test2$baseline_av_z = scales::rescale(test2$baseline_av)
@@
 
-load('../../models/paper/bayes_fit3.rda')
+load('models/paper/bayes_fit3.rda')
 bayes_factor()
@@
 loo.comparisons
+
+# the frequentist fit reported in the paper
+# watch out for factor levels
+# lex typ: typical - random - reversed
+# reg shift: nc - -40% - +40%
+lme.fit2 = glmer(resp_post_reg ~ 1 + baseline_av_z * lex_typicality + pre_reg_av_z + reg_rate +
+    (1 + baseline_av_z | participant_id) + 
+    (1  | word), 
+  data = test2, 
+  family = binomial (link = logit),
+  control=glmerControl(optimizer="bobyqa")
+)
+summary(lme.fit2)
+# not the same as model in paper.
 

### 2019-07-02 models/paper/convergence_paper_models.Rmd
effects

--- 2019-05-28 models/paper/convergence_paper_models.Rmd
@@
 library(lme4)
+library(merTools)
 library(tidyverse)
@@
 
-try(setwd('../../models/paper'))
-try(setwd('~/Github/RaczBecknerHayPierrehumbert2019/models/paper'))
+try(setwd('~/Work/NZILBB/RaczBecknerHayPierrehumbert2019'))
+try(setwd('~/Github/RaczBecknerHayPierrehumbert2019'))
 ```
@@
 ```{r }
-source('../../functions/grabdata.R')
+source('functions/grabdata.R')
 
@@
 test2$lex_typicality = factor(test2$lex_typicality, levels = c('typical', 'random', 'atypical'))
-test2$reg_rate = factor(test2$reg_rate, levels = c('-40%', 'nc', '+40%'))
+test2$reg_rate = factor(test2$reg_rate, levels = c('nc', '-40%', '+40%'))
 test2$baseline_av_z = scales::rescale(test2$baseline_av)
 test2$pre_reg_av_z = scales::rescale(test2$pre_reg_av)
+test2$baseline_av_c = test2$baseline_av - mean(test2$baseline_av)
+test2$pre_reg_av_c = test2$pre_reg_av - mean(test2$pre_reg_av)
 
@@
 loo.comparisons
-
-# https://arxiv.org/abs/1507.04544 s5.2
 ```
@@
 
-fit3b = brm(resp_post_reg ~ 1 + baseline_av_z * lex_typicality + pre_reg_av_z + reg_rate +
-    (1 + baseline_av_z | participant_id) + 
-    (1 + pre_reg_av_z | word), 
+fit_power1 = glmer(resp_post_reg ~ 1 + baseline_av_c * lex_typicality + pre_reg_av_c + reg_rate +
+    (1 + baseline_av_c | participant_id) + 
+    (1 | word), 
   data = test2b, 
-  family = bernoulli, 
-  control=list(adapt_delta=0.99, stepsize = 0.01, max_treedepth =15), save_all_pars = T)
+  family = binomial ( link = logit ),
+  control=glmerControl(optimizer="bobyqa")
+  )
 
-fit3c = brm(resp_post_reg ~ 1 + baseline_av_z + pre_reg_av_z + reg_rate +
-    (1 + baseline_av_z | participant_id) + 
-    (1 + pre_reg_av_z | word), 
+fit_power2 = glmer(resp_post_reg ~ 1 + baseline_av_c + lex_typicality + pre_reg_av_c + reg_rate +
+    (1 + baseline_av_c | participant_id) + 
+    (1 | word), 
   data = test2b, 
-  family = bernoulli, 
-  control=list(adapt_delta=0.99, stepsize = 0.01, max_treedepth =15), save_all_pars = T)
+  family = binomial ( link = logit ),
+  control=glmerControl(optimizer="bobyqa")
+  )
 
-save(fit3b, file = 'fit3b.rda')
-save(fit3c, file = 'fit3c.rda')
+# summary(fit_power1)
+# summary(fit_power2)
+
+exp((BIC(fit_power2) - BIC(fit_power1))/2)
+# 0.01179529
+# we favour the null model
 
@@
 
+The lme4 fit reported in the paper is here.
+
 ```{r }
-load('fit3b.rda')
-load('fit3c.rda')
+# the frequentist fit reported in the paper
+# watch out for factor levels
+# lex typ: typical - random - reversed
+# reg shift: nc - -40% - +40%
+lme.fit2 = glmer(resp_post_reg ~ 1 + baseline_av_c * lex_typicality + pre_reg_av_c + reg_rate +
+    (1 + baseline_av_c | participant_id) + 
+    (1  | word), 
+  data = test2, 
+  family = binomial (link = logit),
+  control=glmerControl(optimizer="bobyqa")
+)
+summary(lme.fit2)$coef %>% 
+  xtable
 
-bf = bayes_factor(fit3c, fit3b)
-bf$bf
-# time to ask for help.
+# https://arxiv.org/abs/1507.04544 s5.2
+```
+
+Let's have a few estimates for CI-s.
+
+```{r }
+lme.fit2.preds = predictInterval(merMod = lme.fit2, level = 0.95, n.sims = 1000, stat = 'median', type = 'probability')
+
+test2c = cbind(test2,lme.fit2.preds) 
+test2c %>%   
+  group_by(
+    reg_rate
+  ) %>% 
+  summarise(
+    fit = mean(fit),
+    upr = mean(upr),
+    lwr = mean(lwr)
+  ) %>% 
+  ggplot(aes(x = reg_rate, y = fit)) +
+  geom_point() +
+  geom_linerange(aes(ymin = lwr, ymax = upr))
+
+confint(lme.fit2, method = 'Wald', level = 0.95)
+effects::allEffects(lme.fit2)
 ```

### 2019-12-16 models/paper/convergence_paper_models.Rmd
new figs

@@
 ```{r }
-source('functions/grabdata.R')
+# source('functions/grabdata.R')
+
+celex = read_csv('data/celex_verbs_with_categories_csv_fixed_moder.txt')
+
+celex = celex %>% 
+  select(
+    present.o,
+    present,
+    moder,
+    regular
+  ) %>% 
+  rename(
+    word = present.o,
+    disc.old = present
+  ) %>%
+  mutate(
+    category = case_when(
+      regular == 'Y' ~ 'regular',
+      regular == 'N' ~ 'irregular'
+    )
+  )  %>% 
+  filter(!is.na(category))
+
+# fixing transcription
+
+transcription.matcher = read_csv('data/transcription_matcher.txt')
+
+celex = left_join(celex, transcription.matcher)
+
+celex[celex$disc == 'riN',]
+
+# test set
+baseline = read_csv('data/convergence_paper_baseline_data.txt')
+
+baseline.verbs = read_csv('data/baseline_verbs.txt')
+# esp data
+espdata = read_csv('data/convergence_paper_esp_data.txt')
+
+# esp only post-test with gcm weights
+test2 = read_csv('data/convergence_paper_esp_test2_predictions.txt')
 
@@
 # lex typ: typical - random - reversed
-# reg shift: nc - -40% - +40%
-lme.fit2 = glmer(resp_post_reg ~ 1 + baseline_av_c * lex_typicality + pre_reg_av_c + reg_rate +
+# reg rate: nc - -40% - +40%
+test2$lexical_typicality = test2$lex_typicality
+test2$regularisation_rate = ifelse(
+  as.character(test2$reg_rate) == 'nc', 'no change',
+  as.character(test2$reg_rate)
+  )
+test2$regularisation_rate = factor(test2$regularisation_rate, levels = c('no change', '-40%', '+40%'))
+
+lme.fit2 = glmer(resp_post_reg ~ 1 + baseline_av_c * lexical_typicality + pre_reg_av_c + regularisation_rate +
     (1 + baseline_av_c | participant_id) + 
@@
 )
-summary(lme.fit2)$coef %>% 
-  xtable
+summary(lme.fit2)$coef
+
+plot(effects::effect('regularisation_rate', lme.fit2), main = '', xlab = 'regularisation shift', ylab = 'regular response in post-test')
+plot(effects::effect('baseline_av_c:lexical_typicality', lme.fit2), multiline = T, lines=list(lty=c(1,2,3)), main = '', xlab = 'verb baseline mean', ylab = 'regular response in post-test')
 
@@
 effects::allEffects(lme.fit2)
+
+test2 %>% 
+  mutate(
+    reg_rate2 = factor(reg_rate, levels = c('-40%', 'nc', '+40%'))
+  ) %>% 
+  ggplot(aes(x = reg_rate2, y = post_reg_av, fill = reg_rate2)) +
+  geom_jitter(aes(colour = reg_rate2), width = 0.3) +
+    geom_violin(aes(alpha = 0.5)) +
+    geom_boxplot(width = 0.2) +
+    # stat_summary(fun.y=mean, geom="point", shape=16, size=4) +
+    scale_fill_brewer(palette="Set2") +
+    scale_colour_brewer(palette="Set2") +
+    theme(
+      text = element_text(size=30), 
+      # axis.text.x = element_text(angle=90, vjust=0.5), 
+      # axis.title.x=element_blank(), 
+      legend.position = 'none'
+      ) +
+    xlab('bot regularisation shift') +
+    ylab('participant post-test mean regularisation') +
+    scale_x_discrete(labels=c('-40%', 'no change', '+40%'))
+ggsave('reg_shift_av.pdf', width = 9, height = 6)
+
+test2 %>% 
+  group_by(
+    word,
+    baseline_av,
+    lex_typicality
+  ) %>% 
+  summarise(
+    mean = mean(resp_post_reg)
+  ) %>% 
+  ggplot(aes(x = baseline_av, y = mean, colour = lex_typicality, label = word)) +
+  geom_label() +
+  geom_smooth(method = 'lm', se = F) +
+  scale_colour_brewer(palette="Dark2") +
+    theme(
+      text = element_text(size=30), 
+      # axis.text.x = element_text(angle=90, vjust=0.5), 
+      # axis.title.x=element_blank(), 
+      legend.position = 'none'
+      ) +
+    xlab('verb baseline mean regularisation') +
+    ylab('verb post-test mean regularisation')
+ggsave('lex_typ_av.pdf', width = 9, height = 9)
+
+sum1 = confint(lme.fit2, method = 'Wald') %>% data.frame
+sum2 = summary(lme.fit2)$coef %>% data.frame
+sum1$Term = rownames(sum1)
+sum2$Term = rownames(sum2)
+sum3 = left_join(sum2, sum1) %>% 
+  rename(`Std error` = Std..Error,
+    `z value` = z.value,
+    `2.5%` = X2.5..,
+    `97.5%` = X97.5..
+    )
+sum3 %>% dplyr::select(Term, Estimate, `Std error`, `z value`, `2.5%`, `97.5%`) %>% knitr::kable(digits = 2)
+
 ```
@@
 t3la = t3 %>% 
-  select(
+  dplyr::select(
     word,
@@
 t3lb = t3 %>% 
-  select(
+  dplyr::select(
     word,
@@
 
-ggplot(t4, aes(x = baseline.averages, y = individual.averages, colour = lex_typicality2, shape = lex_typicality2)) +
+ggplot(t4, aes(x = baseline.averages, y = individual.averages, colour = lex_typicality2, shape = lex_typicality2, linetype = lex_typicality2)) +
   guides(size = F) +
@@
   scale_shape_discrete(name="Lexical\ntypicality") +
-  scale_colour_discrete(name="Lexical\ntypicality") +
-  xlab('verb baseline average') +
-  ylab('verb ESP average')
+  # scale_colour_discrete(name="Lexical\ntypicality") +
+  xlab('verb baseline regularization') +
+  ylab('verb ESP regularization') +
+  scale_colour_brewer(palette = 'Set2', name="Lexical\ntypicality")
 # ggsave('interactionplot.pdf', device = cairo_pdf, width = 12.5, height = 4)

### 2019-12-16 functions/grabdata.R

@@ -1,6 +1,5 @@
-try(setwd('~/Work/NZILBB/RaczBecknerHayPierrehumbert2019'))
 try(setwd('~/Github/RaczBecknerHayPierrehumbert2019'))
-
+# try(setwd('~/Work/NZILBB/RaczBecknerHayPierrehumbert2019'))
 
 # training set
 
@@ -37,7 +36,7 @@ celex[celex$disc == 'riN',]
 # test set
 baseline = read_csv('data/convergence_paper_baseline_data.txt')
 
-
+baseline.verbs = read_csv('data/baseline_verbs.txt')
 # esp data
 espdata = read_csv('data/convergence_paper_esp_data.txt')
 

### 2019-12-17 models/paper/convergence_paper_models.Rmd
confint table

@@
 
+# moving factors round
+test2$lex_typicality = factor(test2$lex_typicality, levels = c('typical', 'random', 'atypical'))
+test2$reg_rate = factor(test2$reg_rate, levels = c('nc', '-40%', '+40%'))
+test2$baseline_av_z = scales::rescale(test2$baseline_av)
+test2$pre_reg_av_z = scales::rescale(test2$pre_reg_av)
+test2$baseline_av_c = test2$baseline_av - mean(test2$baseline_av)
+test2$pre_reg_av_c = test2$pre_reg_av - mean(test2$pre_reg_av)
+
+# lex typ: typical - random - reversed
+# reg rate: nc - -40% - +40%
+test2$lexical_typicality = test2$lex_typicality
+test2$regularisation_rate = ifelse(
+  as.character(test2$reg_rate) == 'nc', 'no change',
+  as.character(test2$reg_rate)
+  )
+test2$regularisation_rate = factor(test2$regularisation_rate, levels = c('no change', '-40%', '+40%'))
+
 ```
@@
 ```{r }
-test2$lex_typicality = factor(test2$lex_typicality, levels = c('typical', 'random', 'atypical'))
-test2$reg_rate = factor(test2$reg_rate, levels = c('nc', '-40%', '+40%'))
-test2$baseline_av_z = scales::rescale(test2$baseline_av)
-test2$pre_reg_av_z = scales::rescale(test2$pre_reg_av)
-test2$baseline_av_c = test2$baseline_av - mean(test2$baseline_av)
-test2$pre_reg_av_c = test2$pre_reg_av - mean(test2$pre_reg_av)
-
 fit1 = brm(resp_post_reg ~ 1 + baseline_av_z * lex_typicality + pre_reg_av_z + reg_rate +
@@
 # watch out for factor levels
-# lex typ: typical - random - reversed
-# reg rate: nc - -40% - +40%
-test2$lexical_typicality = test2$lex_typicality
-test2$regularisation_rate = ifelse(
-  as.character(test2$reg_rate) == 'nc', 'no change',
-  as.character(test2$reg_rate)
-  )
-test2$regularisation_rate = factor(test2$regularisation_rate, levels = c('no change', '-40%', '+40%'))
+test2$verb_baseline_average = test2$baseline_av_c
+test2$bot_lexical_typicality = test2$lexical_typicality
+test2$participant_prereg_average = test2$pre_reg_av_c
+test2$bot_regularisation_rate = test2$regularisation_rate
 
-lme.fit2 = glmer(resp_post_reg ~ 1 + baseline_av_c * lexical_typicality + pre_reg_av_c + regularisation_rate +
+lme.fit2 = glmer(resp_post_reg ~ 1 + verb_baseline_average * bot_lexical_typicality + participant_prereg_average + bot_regularisation_rate +
     (1 + baseline_av_c | participant_id) + 
@@
 
-plot(effects::effect('regularisation_rate', lme.fit2), main = '', xlab = 'regularisation shift', ylab = 'regular response in post-test')
-plot(effects::effect('baseline_av_c:lexical_typicality', lme.fit2), multiline = T, lines=list(lty=c(1,2,3)), main = '', xlab = 'verb baseline mean', ylab = 'regular response in post-test')
+plot(effects::effect('bot_regularisation_rate', lme.fit2), main = '', xlab = 'regularisation shift', ylab = 'regular response in post-test')
+plot(effects::effect('verb_baseline_average:bot_lexical_typicality', lme.fit2), multiline = T, lines=list(lty=c(1,2,3)), main = '', xlab = 'verb baseline mean', ylab = 'regular response in post-test')
 
@@
 
-Let's have a few estimates for CI-s.
+Here is a simple confint plot using Wald estimates.
+
+```{r }
+t1 = broom::tidy(lme.fit2)
+
+# we get some wald confidence intervals
+
+confints = confint(lme.fit2, method = 'Wald') %>% broom::tidy()
+confints = confints[-1,]
+names(confints) = c('term', '2.5%', '97.5%')
+t1 = inner_join(t1, confints) %>% 
+  dplyr::select(-p.value,-group)
+
+t1 %>% 
+  mutate(
+    row_id = 9:1, # watch this
+  ) %>% 
+  mutate(
+    term2 = 
+    case_when(
+      term == 'verb_baseline_average' ~ 'verb baseline average',
+      term == 'bot_lexical_typicalityrandom' ~ 'bot lexical typicality = random',
+      term == 'bot_lexical_typicalityatypical' ~ 'bot lexical typicality = atypical',
+      term == 'participant_prereg_average' ~ 'participant prereg average',
+      term == 'bot_regularisation_rate-40%' ~ 'bot regularisation rate = -40%',
+      term == 'bot_regularisation_rate+40%' ~ 'bot regularisation rate = +40%',
+      term == 'verb_baseline_average:bot_lexical_typicalityrandom' ~ 'verb baseline average : bot lexical typicality = random',
+      term == 'verb_baseline_average:bot_lexical_typicalityatypical' ~ 'verb baseline average : bot lexical typicality = atypical',
+      term == '(Intercept)' ~ 'Intercept'
+      )
+    ) %>% 
+  ggplot(aes(x = term2 %>% reorder(row_id), y = estimate)) +
+  geom_point() +
+  geom_linerange(aes(x = term2, ymin = `2.5%`, ymax = `97.5%`)) +
+  geom_hline(aes(yintercept = 0), linetype = 'dashed') +
+  xlab('') +
+  coord_flip() +
+  ggtitle('Estimates and 95% confidence intervals in model for test results')
+ggsave('terms_figure.pdf', width = 9, height = 5)
+
+```
+
+Let's have a few predicted intervals.
 
@@
   geom_linerange(aes(ymin = lwr, ymax = upr))
-
-confint(lme.fit2, method = 'Wald', level = 0.95)
-effects::allEffects(lme.fit2)
 

### 2019-12-18 models/paper/convergence_paper_models.Rmd
fig4

@@
   ylab('verb ESP regularization') +
+  guides(linetype = F) +
   scale_colour_brewer(palette = 'Set2', name="Lexical\ntypicality")

### 2020-01-05 models/gcm/gcm.Rmd
new figs
final submission

@@
 
-setwd('~/Work/NZILBB/RaczBecknerHayPierrehumbert2019/models/gcm')
+setwd('~/Github/RaczBecknerHayPierrehumbert2019/models/gcm')
 ```
@@
 ```{r }
-source('~/Work/NZILBB/RaczBecknerHayPierrehumbert2019/functions/grabfunctions.R')
+source('Github/RaczBecknerHayPierrehumbert2019/functions/grabfunctions.R')
 
-source('~/Work/NZILBB/RaczBecknerHayPierrehumbert2019/functions/grabdata.R')
+source('Github/RaczBecknerHayPierrehumbert2019/functions/grabdata.R')
 
-lookup = read_csv('~/Work/NZILBB/RaczBecknerHayPierrehumbert2019/models/gcm/feature_distance_lookup_tidy.txt')
+lookup = read_csv('Github/RaczBecknerHayPierrehumbert2019/models/gcm/feature_distance_lookup_tidy.txt')
 
@@
 ```{r create_lookup_table2, eval=F}
-# of course, what I do above is not what I did in practice.
-in practice, I fit four different sets of words because I was in the process of figuring out what's going on and what I need.
-The first set had word pairs repeat several times because of a join.
-Etc.
-# see pocketgcm_create_lookup_table.R
-
+# I did this in four batches originally as I was in the process of figuring things out.
 # I'm not sharing these files, but you can see the code here to get an idea.
@@
 # batch 1
-distances1 = read_csv('~/Work/NZILBB/convergence_esp/ESP2019/pocketgcm/Alignment/out.txt', col_names = F)
+distances1 = read_csv('Github/convergence_esp/ESP2019/pocketgcm/Alignment/out.txt', col_names = F)
 names(distances1) = 'feature.distance'
 
-pairs1 = read_delim('~/Work/NZILBB/convergence_esp/ESP2019/pocketgcm/Alignment/everybody.txt', '\t', col_names = F)
+pairs1 = read_delim('Github/convergence_esp/ESP2019/pocketgcm/Alignment/everybody.txt', '\t', col_names = F)
 names(pairs1) = c('target', 'training')
@@
 # batch 2
-distances2 = read_csv('~/Work/NZILBB/convergence_esp/ESP2019/pocketgcm/Alignment/out2.txt', col_names = F)
+distances2 = read_csv('Github/convergence_esp/ESP2019/pocketgcm/Alignment/out2.txt', col_names = F)
 names(distances2) = 'feature.distance'
 
-pairs2 = read_delim('~/Work/NZILBB/convergence_esp/ESP2019/pocketgcm/Alignment/everybody2.txt', '\t', col_names = F)
+pairs2 = read_delim('Github/convergence_esp/ESP2019/pocketgcm/Alignment/everybody2.txt', '\t', col_names = F)
 names(pairs2) = c('target', 'training')
@@
 # batch 3
-distances3 = read_csv('~/Work/NZILBB/convergence_esp/ESP2019/pocketgcm/Alignment/out3.txt', col_names = F)
+distances3 = read_csv('Github/convergence_esp/ESP2019/pocketgcm/Alignment/out3.txt', col_names = F)
 names(distances3) = 'feature.distance'
 
-pairs3 = read_delim('~/Work/NZILBB/convergence_esp/ESP2019/pocketgcm/Alignment/everybody3.txt', '\t', col_names = F)
+pairs3 = read_delim('Github/convergence_esp/ESP2019/pocketgcm/Alignment/everybody3.txt', '\t', col_names = F)
 names(pairs3) = c('target', 'training')
@@
 # batch 4
-distances4 = read_csv('~/Work/NZILBB/convergence_esp/ESP2019/pocketgcm/Alignment/out4.txt', col_names = F)
+distances4 = read_csv('Github/convergence_esp/ESP2019/pocketgcm/Alignment/out4.txt', col_names = F)
 names(distances4) = 'feature.distance'
 
-pairs4 = read_delim('~/Work/NZILBB/convergence_esp/ESP2019/pocketgcm/Alignment/everybody4.txt', '\t', col_names = F)
+pairs4 = read_delim('Github/convergence_esp/ESP2019/pocketgcm/Alignment/everybody4.txt', '\t', col_names = F)
 names(pairs4) = c('target', 'training')
@@
 
+celex2 %>% filter(is.na(moder.training), category.training == 'regular')
+
 # some verbs are in celex as both regular and irregular. these verbs are in the lookup only once.
@@
 
-Our lookup table has been made. Let's hope it's not the way Joe Pesci was made in Goodfellas.
+Our lookup table has been made.
 
@@
 
-Reviewer points out that we are tilting the field in the GCM's favour by restricting its training set to verbs that basically look exactly like the verb schemata we are interested in.
-One of their alternative suggestions is to keep Moder classes but include *all* regular verbs in the regular category.
+One could argue that we are tilting the field in the GCM's favour by restricting its training set to verbs that basically look exactly like the verb schemata we are interested in.
+One alternative suggestion is to keep Moder classes but include *all* regular verbs in the regular category.
 
@@
 
-## 2c. Tuning s for the reviewer GCM
+## 2c. Tuning s for the GCM with the larger regular training set
 
@@
 
-## 3a. Fitting the reviewer GCM (I hope this catches on) on the ESP experiment
+## 3a. Fitting this second GCM on the ESP experiment
 
@@
 
+We are done with the GCM fits on the data.
+
 ## 5. Check the reviewer critique of moder classes on baseline data

### 2020-01-05 models/paper/convergence_paper_models.Rmd
final submission

@@
 
-try(setwd('~/Work/NZILBB/RaczBecknerHayPierrehumbert2019'))
 try(setwd('~/Github/RaczBecknerHayPierrehumbert2019'))
@@
 ```{r }
-# source('functions/grabdata.R')
-
-celex = read_csv('data/celex_verbs_with_categories_csv_fixed_moder.txt')
-
-celex = celex %>% 
-  select(
-    present.o,
-    present,
-    moder,
-    regular
-  ) %>% 
-  rename(
-    word = present.o,
-    disc.old = present
-  ) %>%
-  mutate(
-    category = case_when(
-      regular == 'Y' ~ 'regular',
-      regular == 'N' ~ 'irregular'
-    )
-  )  %>% 
-  filter(!is.na(category))
-
-# fixing transcription
-
-transcription.matcher = read_csv('data/transcription_matcher.txt')
-
-celex = left_join(celex, transcription.matcher)
-
-celex[celex$disc == 'riN',]
-
-# test set
-baseline = read_csv('data/convergence_paper_baseline_data.txt')
-
-baseline.verbs = read_csv('data/baseline_verbs.txt')
-# esp data
-espdata = read_csv('data/convergence_paper_esp_data.txt')
-
-# esp only post-test with gcm weights
-test2 = read_csv('data/convergence_paper_esp_test2_predictions.txt')
-
-# moving factors round
-test2$lex_typicality = factor(test2$lex_typicality, levels = c('typical', 'random', 'atypical'))
-test2$reg_rate = factor(test2$reg_rate, levels = c('nc', '-40%', '+40%'))
-test2$baseline_av_z = scales::rescale(test2$baseline_av)
-test2$pre_reg_av_z = scales::rescale(test2$pre_reg_av)
-test2$baseline_av_c = test2$baseline_av - mean(test2$baseline_av)
-test2$pre_reg_av_c = test2$pre_reg_av - mean(test2$pre_reg_av)
-
-# lex typ: typical - random - reversed
-# reg rate: nc - -40% - +40%
-test2$lexical_typicality = test2$lex_typicality
-test2$regularisation_rate = ifelse(
-  as.character(test2$reg_rate) == 'nc', 'no change',
-  as.character(test2$reg_rate)
-  )
-test2$regularisation_rate = factor(test2$regularisation_rate, levels = c('no change', '-40%', '+40%'))
+source('functions/grabdata.R')
 
@@
 
-## jumping to Appendix E!
+## jumping to SI section E!
 
@@
 
-## jumping to Appendix F; Table 12
+## jumping to SI section F; Table 12
 
@@
 
-## jumping to Appendix F.2
+## jumping to SI section F.2
 

### 2020-01-05 functions/grabdata.R

@@ -1,8 +1,9 @@
 try(setwd('~/Github/RaczBecknerHayPierrehumbert2019'))
 # try(setwd('~/Work/NZILBB/RaczBecknerHayPierrehumbert2019'))
 
+######################################################################
 # training set
-
+######################################################################
 
 celex = read_csv('data/celex_verbs_with_categories_csv_fixed_moder.txt')
 
@@ -31,9 +32,12 @@ transcription.matcher = read_csv('data/transcription_matcher.txt')
 
 celex = left_join(celex, transcription.matcher)
 
-celex[celex$disc == 'riN',]
+# celex[celex$disc == 'riN',]
 
+######################################################################
 # test set
+######################################################################
+
 baseline = read_csv('data/convergence_paper_baseline_data.txt')
 
 baseline.verbs = read_csv('data/baseline_verbs.txt')

### 2020-01-05 models/mgl/mgl_esp_create_input.R

@@ -1,0 +1,140 @@
+library(Hmisc)
+library(dplyr)
+library(stringr)
+library(knitr)
+library(lme4)
+library(RePsychLing)
+library(tidyverse)
+setwd('~/Github/RaczBecknerHayPierrehumbert2019/')
+
+# MGL batch run doesn't pull in features. this has to be done by hand. 
+# I don't want to run 222 participants by hand.
+# Experiment regularisation rate, lexical typicality and list block order create the possible training + test data combinations for the individual MGL-s. 
+# Individual participants will move round slightly more but this is a reasonable approximation (and still takes hours)
+# I can pick the unique combinations of reg rate + lex typ + list_block_order and run one participant each.
+# this doesn't cover all participants, since 
+
+esp <- read_csv('data/convergence_paper_esp_data.txt') %>% filter(phase == 'test2')
+
+participants <- esp %>% select(participant_id,reg_rate,lex_typicality,list_block_order) %>% unique %>% arrange(reg_rate,lex_typicality,list_block_order) %>% mutate(part_type = paste(reg_rate, lex_typicality, list_block_order, sep = ' ')) %>% select(participant_id,part_type)
+participants <- participants[!duplicated(participants$part_type),]
+handles <- read.csv("data/participanthandles.txt")
+participants <- merge(participants,handles)
+
+# now I paste the relevant participants' in files into inputlists_wf_minimal
+
+# now I run the Minimal Generalisation Learner on each, 
+
+# you have to quit the program EVERY TIME after you fit an individual file, otherwise it doesn't work.
+
+# this works: turn mingenlearner off and on again every time, set lower and upper to 55 95 every time.
+
+# I check every sum file to see if it specifies a set of segments for P environment. that is indicative of features being used. (length of fit is also indicative: with features it takes a noticable amount of time, w/o it fits almost instantly)
+# close mgl. check sum output for features. open mgl. set 55 and 95. select next in. run. takes about 50 sec to fit.
+
+# This takes a couple hours.
+
+# I gather predictions.
+
+# participants2 <- participants %>% filter(participant < 100)
[…]

### 2020-01-05 models/mgl/mgl_processor.R

@@ -1,11 +1,9 @@
 # code to process an MGL sum file
+setwd('~/Github/RaczBecknerHayPierrehumbert2019/models/gcm')
 
-library(dplyr)
-library(stringr)
-library(knitr)
+library(tidyverse)
 
-
-handles <- read.delim("256verbs_print_forms.txt", sep=",") %>% select(base.form,category,base.print,regular.form,irregular.form) %>% rename(present.form = base.form)
+handles <- read.delim("data/256verbs_print_forms.txt", sep=",") %>% select(base.form,category,base.print,regular.form,irregular.form) %>% rename(present.form = base.form)
 # get formatting right
 handles$irregular.form <- sub("5","o",handles$irregular.form)
 handles$regular.form <- sub("Id","@d",handles$regular.form)
@@ -64,3 +62,8 @@ rules <- merge(regular.rules,irregular.rules, all = T)
 rules <- rules %>% mutate(ir.conf = ifelse(is.na(ir.conf), 0, ir.conf)) %>% mutate(indiv_mgl_conf = reg.conf / (reg.conf + ir.conf)) %>% select(present.form,indiv_mgl_conf)
 return(rules)
 }
+
+# an example
+
+out = 
+getMGLRules()

### 2020-01-06 models/mgl/mgl_esp_create_input.R
fixes

@@ -1,7 +1,7 @@
 library(Hmisc)
-library(dplyr)
-library(stringr)
-library(knitr)
+# library(dplyr)
+# library(stringr)
+# library(knitr)
 library(lme4)
 library(RePsychLing)
 library(tidyverse)
@@ -138,3 +138,66 @@ esp <- read.csv('data/convergence_paper_esp_data.txt')
 esp2 <- esp %>% mutate(part_type = paste(reg_rate, lex_typicality, list_block_order, sep = ' '))
 # esp_mini <- merge(esp_mini,esp_mini_wf)
 esp2 <- merge(esp2, esp_mini_wf)
+
+###############################################################
+# Do we lose information by not fitting the MGL on all participants?
+###############################################################
+
+# let's fit a simple model to see how predictions match with actual post-test behaviour.
+# we do this for the entire data (fit1/fit2)
+# we do this for the participants where the MGL was hand-fit (fit1b/fit2b)
+
+test2 = read_csv('data/convergence_paper_esp_test2_predictions.txt')
+
+test2b = test2[test2$participant_id %in% participants$participant_id,]
+
+# a simple model of how predictions match post-test behaviour on the entire data
+
+fit1 = glmer(resp_post_reg ~ 1 + 
+    baseline_mgl_features + 
+    baseline_gcm_features + 
+    individual_mgl_features + 
+    individual_gcm_features + 
+    (1 | participant_id) + (1|word), 
+  family = binomial, data = test2, control=glmerControl(optimizer="bobyqa")
+  )
+summary(fit1)
+
+# a simple model of how predictions match post-test behaviour on the participants fit
+
+fit1b = glmer(resp_post_reg ~ 1 + 
+    baseline_mgl_features + 
+    baseline_gcm_features + 
+    individual_mgl_features + 
+    individual_gcm_features + 
+    (1 | participant_id) + (1|word), 
+  family = binomial, data = test2b, control=glmerControl(optimizer="bobyqa")
+  )
+summary(fit1b)
+
+# comparison models
+
+fit2 = glmer(resp_post_reg ~ 1 + 
+    baseline_mgl_features + 
+    # baseline_gcm_features +
+    # individual_mgl_features + 
+    individual_gcm_features + 
+    (1 | participant_id) + (1|word), 
+  family = binomial, data = test2, control=glmerControl(optimizer="bobyqa")
+  )
+summary(fit2)
+
[…]

### 2020-01-22 models/paper/convergence_paper_models.Rmd
clay comments

@@
 
-These are the models and stats for the convergence paper, going through it by section number.
+These are the models and stats reported in the main text of the convergence paper. We go through the main text by section number and replicate the relevant statistics / figures / models here.
+
+This file references both the main text and the supplementary information.
+To make things slightly more complicated, the appendix of the main text and the SI overlap.
+And section lettering is different in the appendix of the main text and the si.
+Here, we use lettering that matches the si.
 
@@
 
-Our weapon of choice is typially lme4. But it won't converge with full random effects structure. So we turn to brms.
+Our weapon of choice is typially lme4.
+But it won't converge with full random effects structure.
+So we turn to brms.
+We assume flat priors and a bernoulli error distribution, which is preferred to the binomial distribution in brms for binary outcomes.
+We compare models using leave-one-out cross-validation.
+
+The Bayesian models are equivalent to the frequentist models reported in the paper, at least conceptually, if not in many other ways.
 
@@
 
-We definitely want some random slopes (fit1 is worse than the rest). Other than that, the various constellations do not improve the model's expected log pointwise predictive density.
+What do we learn from LOO comparisons?
+
+Fit1 is worse than the rest, so random slopes do improve model fit.
+Building up the random effect structure for the word random intercept does not improve the model's expected log pointwise predictive density.
 
@@
 
-## Bayes factors!
+## Bayes factors
+
+We could not reject the null on the absence of difference between our typical and atypical (reversed) conditions.
+Upon reviewer suggestion, we used Bayes factors to test whether this is due to the lack of evidence.
 
@@
 # 0.01179529
-# we favour the null model
+# it is unlikely that there is a difference here
 
@@
 
-The lme4 fit reported in the paper is here.
+The lme4 fit reported in the paper is here:
 
@@
       term == 'verb_baseline_average:bot_lexical_typicalityrandom' ~ 'verb baseline average : bot lexical typicality = random',
-      term == 'verb_baseline_average:bot_lexical_typicalityatypical' ~ 'verb baseline average : bot lexical typicality = atypical',
+      term == 'verb_baseline_average:bot_lexical_typicalityatypical' ~ 'verb baseline average : bot lexical typicality = reversed',
       term == '(Intercept)' ~ 'Intercept'

### 2020-01-28 models/gcm/gcm.Rmd
in progress

@@
 ```{r }
-source('Github/RaczBecknerHayPierrehumbert2019/functions/grabfunctions.R')
+source('~/Github/RaczBecknerHayPierrehumbert2019/functions/grabfunctions.R')
 
-source('Github/RaczBecknerHayPierrehumbert2019/functions/grabdata.R')
+source('~/Github/RaczBecknerHayPierrehumbert2019/functions/grabdata.R')
 
-lookup = read_csv('Github/RaczBecknerHayPierrehumbert2019/models/gcm/feature_distance_lookup_tidy.txt')
+lookup = read_csv('~/Github/RaczBecknerHayPierrehumbert2019/models/gcm/feature_distance_lookup_tidy.txt')
 
@@
 # test words
-test = baseline.verbs$disc
+test = unique(baseline$disc)
 # training words: verbs in celex in moder classes
@@
   select(-vowels) %>% 
-  unnest()
+  unnest(cols = c(nvowels))
 
@@
 # batch 1
-distances1 = read_csv('Github/convergence_esp/ESP2019/pocketgcm/Alignment/out.txt', col_names = F)
+distances1 = read_csv('~/Github/convergence_esp/ESP2019/pocketgcm/Alignment/out.txt', col_names = F)
 names(distances1) = 'feature.distance'
 
-pairs1 = read_delim('Github/convergence_esp/ESP2019/pocketgcm/Alignment/everybody.txt', '\t', col_names = F)
+pairs1 = read_delim('~/Github/convergence_esp/ESP2019/pocketgcm/Alignment/everybody.txt', '\t', col_names = F)
 names(pairs1) = c('target', 'training')
@@
 # batch 2
-distances2 = read_csv('Github/convergence_esp/ESP2019/pocketgcm/Alignment/out2.txt', col_names = F)
+distances2 = read_csv('~/Github/convergence_esp/ESP2019/pocketgcm/Alignment/out2.txt', col_names = F)
 names(distances2) = 'feature.distance'
 
-pairs2 = read_delim('Github/convergence_esp/ESP2019/pocketgcm/Alignment/everybody2.txt', '\t', col_names = F)
+pairs2 = read_delim('~/Github/convergence_esp/ESP2019/pocketgcm/Alignment/everybody2.txt', '\t', col_names = F)
 names(pairs2) = c('target', 'training')
@@
 # batch 3
-distances3 = read_csv('Github/convergence_esp/ESP2019/pocketgcm/Alignment/out3.txt', col_names = F)
+distances3 = read_csv('~/Github/convergence_esp/ESP2019/pocketgcm/Alignment/out3.txt', col_names = F)
 names(distances3) = 'feature.distance'
 
-pairs3 = read_delim('Github/convergence_esp/ESP2019/pocketgcm/Alignment/everybody3.txt', '\t', col_names = F)
+pairs3 = read_delim('~/Github/convergence_esp/ESP2019/pocketgcm/Alignment/everybody3.txt', '\t', col_names = F)
 names(pairs3) = c('target', 'training')
@@
 # batch 4
-distances4 = read_csv('Github/convergence_esp/ESP2019/pocketgcm/Alignment/out4.txt', col_names = F)
+distances4 = read_csv('~/Github/convergence_esp/ESP2019/pocketgcm/Alignment/out4.txt', col_names = F)
 names(distances4) = 'feature.distance'
 
-pairs4 = read_delim('Github/convergence_esp/ESP2019/pocketgcm/Alignment/everybody4.txt', '\t', col_names = F)
+pairs4 = read_delim('~/Github/convergence_esp/ESP2019/pocketgcm/Alignment/everybody4.txt', '\t', col_names = F)
 names(pairs4) = c('target', 'training')

### 2020-01-28 models/paper/convergence_paper_models.Rmd
in progress

@@
 
+espdata[espdata$lex_typicality == 'atypical',]$lex_typicality = 'reversed'
+
+test2[test2$lex_typicality == 'atypical',]$lex_typicality = 'reversed'
+
 ```
@@
 ```{r }
-baseline %>% group_by(word) %>% summarise(regular = mean(regular)) %>% filter(word == 'splive') 
+baseline %>% 
+  group_by(word) %>% 
+  summarise(regular = mean(regular)) %>% 
+  filter(word == 'fim') 
 ```
@@
 ```{r }
-espdata %>% select(participant_id,lex_typicality,reg_rate) %>% unique %>% group_by(lex_typicality,reg_rate) %>% summarise(n = n()) 
+espdata %>% 
+  select(participant_id,lex_typicality,reg_rate) %>% 
+  unique %>% 
+  group_by(lex_typicality,reg_rate) %>% 
+  summarise(n = n()) %>% 
+  pivot_wider(names_from = reg_rate, values_from = n)
 ```
@@
 
-baseline %>% select(word,r.baseline_gcm_features,baseline_gcm_edits) %>% filter(word == 'splive') %>% unique
-baseline %>% select(word,r.baseline_mgl_features,baseline_mgl_edits) %>% filter(word == 'splive') %>% unique
+baseline %>% 
+  select(word,r.baseline_gcm_features,baseline_gcm_edits) %>% 
+  filter(word == 'splive') %>% 
+  unique
+
+baseline %>% 
+  select(word,r.baseline_mgl_features,baseline_mgl_edits) %>% 
+  filter(word == 'splive') %>% 
+  unique
 
@@
 # don't want to load Hmisc, messes with select()
-with(baseline, Hmisc::somers2(baseline_gcm_features,regular)['C']) # .581
+with(baseline, Hmisc::somers2(baseline_gcm_features,regular)['C'])
 
-with(baseline, Hmisc::somers2(baseline_mgl_features,regular)['C']) # .586
+with(baseline, Hmisc::somers2(baseline_mgl_features,regular)['C']) 
 ```
@@
 
-save(m3, file = '../../models/paper/m3.rda')
+save(m3, file = 'models/paper/m3.rda')
 
@@
 
-save(m4, file = '../../models/paper/m4.rda')
+save(m4, file = 'models/paper/m4.rda')
 ```
@@
 
-load('../../models/paper/m4.rda')
-load('../../models/paper/m3.rda')
+load('models/paper/m4.rda')
+load('models/paper/m3.rda')
 
 summary(m3)
-#                         Estimate Std. Error z value Pr(>|z|)    
-# (Intercept)             -0.44918    0.11202  -4.010 6.08e-05 ***
-# c.baseline_mgl_features  1.27846    0.07525  16.989  < 2e-16 ***
-# gcm_residual             1.37767    0.17171   8.023 1.03e-15 ***
+
 summary(m4)
-#                         Estimate Std. Error z value Pr(>|z|)    
-# (Intercept)             -1.12780    0.14259  -7.909 2.58e-15 ***
-# c.baseline_gcm_features  2.29014    0.15126  15.140  < 2e-16 ***
-# mgl_residual             0.95005    0.08549  11.113  < 2e-16 ***
+
 ```
@@
 
-with(baseline, Hmisc::somers2(baseline_gcm_features,regular)['C']) # .58
-with(baseline, Hmisc::somers2(baseline_mgl_features,regular)['C']) # .59
+# first method
 
-with(test2, Hmisc::somers2(baseline_gcm_features,resp_post_reg)['C']) # .6
-with(test2, Hmisc::somers2(individual_gcm_features,resp_post_reg)['C']) # .68
+with(baseline, Hmisc::somers2(baseline_gcm_features,regular)['C']) 
+with(baseline, Hmisc::somers2(baseline_mgl_features,regular)['C'])
 
-with(test2, Hmisc::somers2(baseline_mgl_features,resp_post_reg)['C']) # .61
-with(test2, Hmisc::somers2(individual_mgl_features,resp_post_reg)['C']) # .55
+with(test2, Hmisc::somers2(baseline_gcm_features,resp_post_reg)['C'])
+with(test2, Hmisc::somers2(individual_gcm_features,resp_post_reg)['C'])
+
+with(test2, Hmisc::somers2(baseline_mgl_features,resp_post_reg)['C']) 
+with(test2, Hmisc::somers2(individual_mgl_features,resp_post_reg)['C'])
 ```
@@
 ```{r eval=F}
+
+# second method
+
 test2 = test2 %>% 
@@
 m5 <- glmer(resp_post_reg ~ c.baseline_gcm_features + gcm_wf_score_diff + pre_reg_av + (1 + c.baseline_gcm_features + gcm_wf_score_diff | participant_id) + (1 | word), data = test2, family = binomial (link = logit), control=glmerControl(optimizer="bobyqa"))
-save(m5, file = '../../models/paper/m5.rda')
+save(m5, file = 'models/paper/m5.rda')
 
 m6 <- glmer(resp_post_reg ~ c.baseline_mgl_features + mgl_wf_score_diff + pre_reg_av + (1 + c.baseline_mgl_features + mgl_wf_score_diff | participant_id) + (1 | word), data = test2, family = binomial (link = logit), control=glmerControl(optimizer="bobyqa"))
-save(m6, file = '../../models/paper/m6.rda')
+save(m6, file = 'models/paper/m6.rda')
 
-# Third
+# third method
+
+m7a <- glmer(resp_post_reg ~ c.individual_gcm_features + c.baseline_mgl_features + c.individual_mgl_features + c.baseline_gcm_features + pre_reg_av + (1 + c.individual_gcm_features + c.baseline_mgl_features | participant_id) + (1 | word), data = test2, family = binomial (link = logit), control=glmerControl(optimizer="bobyqa"))
+
+# summary(m7a)
 
 m7 <- glmer(resp_post_reg ~ c.individual_gcm_features + c.baseline_mgl_features + pre_reg_av + (1 + c.individual_gcm_features + c.baseline_mgl_features | participant_id) + (1 | word), data = test2, family = binomial (link = logit), control=glmerControl(optimizer="bobyqa"))
-save(m7, file = '../../models/paper/m7.rda')
+save(m7, file = 'models/paper/m7.rda')
 
@@
 ```{r }
-load('../../models/paper/m5.rda')
-load('../../models/paper/m6.rda')
-load('../../models/paper/m7.rda')
+load('models/paper/m5.rda')
+load('models/paper/m6.rda')
+load('models/paper/m7.rda')
 
 summary(m5)
-#                  Estimate Std. Error z value Pr(>|z|)    
-# (Intercept)              -5.5677     0.3724 -14.949  < 2e-16 ***
-# c.baseline_gcm_features   4.3687     0.4249  10.281  < 2e-16 ***
-# gcm_wf_score_diff         2.6117     0.4696   5.562 2.67e-08 ***
-# pre_reg_av                5.2887     0.6022   8.782  < 2e-16 ***
 summary(m6)
-#                        Estimate Std. Error z value Pr(>|z|)    
-# (Intercept)              -5.0641     0.3374 -15.010   <2e-16 ***
-# c.baseline_mgl_features   2.0335     0.2079   9.782   <2e-16 ***
-# mgl_wf_score_diff         0.2085     0.1723   1.210    0.226    
-# pre_reg_av                7.4492     0.5996  12.425   <2e-16 ***
 summary(m7)
-#                      Estimate Std. Error z value Pr(>|z|)    
-# (Intercept)                -5.6326     0.3364 -16.743  < 2e-16 ***
-# c.individual_gcm_features   2.7857     0.3776   7.378 1.61e-13 ***
-# c.baseline_mgl_features     1.5998     0.1709   9.361  < 2e-16 ***
-# pre_reg_av                  5.4374     0.5842   9.307  < 2e-16 ***
 summary(m7)$coef %>% round(3) %>% xtable
@@
 
-test2 %>% select(word,moder,verb_baseline_mgl_reg) %>% filter(moder == 'KEPT' & verb_baseline_mgl_reg == 1) %>% unique %>% arrange(word) %>% select(word) %>% pull %>% length
+test2 %>% 
+  select(word,moder,verb_baseline_mgl_reg) %>% 
+  filter(moder == 'KEPT' & verb_baseline_mgl_reg == 1) %>% 
+  unique %>% 
+  arrange(word) %>% 
+  select(word) %>% 
+  pull %>% 
+  length
 # 28, not 57
@@
 ```{r eval=F}
-load('esp_fit.rda')
+load('models/paper/esp_fit.rda')
 formula(fit)
@@
 test2$m7.pred = predict(m7, type = 'response')
-with(test2, Hmisc::somers2(m7.pred, resp_post_reg)) # .84
+with(test2, Hmisc::somers2(m7.pred, resp_post_reg)) 
 ```
@@
              lex_typicality == 'atypical' ~ 'reversed',
+             lex_typicality == 'reversed' ~ 'reversed',
              lex_typicality == 'typical' ~ 'typical',
@@
 # ggsave('interactionplot.pdf', device = cairo_pdf, width = 12.5, height = 4)
-ggsave('fig4_2019.pdf', width = 12.5, height = 4)
+ggsave('fig4_2020.pdf', width = 12.5, height = 4)
 

### 2020-01-28 supplementary_information/si.Rnw
in progress

\documentclass[12pt]{article}
\usepackage[a4paper, margin=1in]{geometry}
\usepackage{graphicx, microtype}
\usepackage{multicol,multirow}
\usepackage{natbib,tipa}

\title{Supplementary Information -- Morphological convergence as on-line lexical analogy}
\author{P\'eter R\'acz, Clay Beckner, Jen Hay, and Janet B.\ Pierrehumbert}

\begin{document}
\SweaveOpts{concordance=TRUE}

<<header, echo=FALSE, eval=TRUE>>=

# libs
library(tidyverse)
library(xtable)

# data
source('~/Github/RaczBecknerHayPierrehumbert2019/functions/grabdata.R')

@

\maketitle

\noindent This is an Sweave file. The Rnw source file contains the code used to compile the pdf output file. It also contains additional code sections that help replicate the results. The files also contain pointers to further detail in the supplementary depository that accompanies the paper.

\noindent In this supplementary information, we detail (\ref{appendixverbs}) the structure of our nonce verb stimuli, the setup of (\ref{appendixgcm}) the Generalized Context Model (GCM) and (\ref{appendixmgl}) the Minimal Generalization Learner (MGL), and (\ref{appendixgcmvsmgl}) how these models compare. 
We illustrate the models using fits on our \emph {baseline} data.

We also discuss (\ref{appendixgcmmglresidualization}) the residualization method used to compare them, (\ref{posttestimplementation}) how well they fit our ESP test, and (\ref{appendixregression}) the regression methods used.

Sections (\ref{appendixgcm}) and (\ref{appendixmgl}) also appear in the main text of  R\'acz et al. (2020), as Appendices A and B. However, this material is included here so that this Supplement can serve as a standalone document for readers who are interested in the technical details of our study.

\section{Regular and irregular verbs in English \label{appendixverbs}}

Four irregular verb classes were defined for our stimuli, based on the vowel alternation and affixation processes that apply to the stem:

\begin{itemize}
    \item DROVE (\textipa{[aI]}/\textipa{[i]} $\rightarrow{}$ \textipa{[oU]})
    \item SANG (\textipa{[I]} $\rightarrow{}$ \textipa{[\ae]})
    \item KEPT (\textipa{[i]} $\rightarrow{}$ \textipa{[E]}Ct)
    \item BURNT (\textipa{[3]}/\textipa{[E]}/\textipa{[I]} $\rightarrow{}$ \textipa{[3]}/\textipa{[E]}/\textipa{[I]}Ct)
\end{itemize}

In the baseline experiment, participants are also tested on monosyllabic verbs that do not change in the irregular past tense (e.g.\ {\em cut}, {\em hit}). These forms are strong outliers and are not reported in the data.

The stimuli for the ESP experiment, 156 nonce verbs across four classes, are shown in Table 1 in the main text. Our formal criteria are based on the behavior of existing verbs in English and their categorization by \cite{bybee1982rules}, \cite{moder1992productivity}, and \cite{albright2003rules}. We made some adjustments to their categories. For instance, our DROVE class is a generalized version of Moder's class 4, described using the \textipa{[aI]}$\rightarrow{}$\textipa{[oU]} alternation but also including {\em weave}. (In contrast, Albright \& Hayes restrict this class to this alternation.) The verb classes could be defined in a number of ways and still be mostly consistent with the work of Moder, for instance. We have trialed a set of possible small changes and found that they have no major effect on our overall results.

<<header, echo=FALSE, eval=FALSE>>=

# table 1
baseline %>% 
  group_by(word) %>% 
  summarise(mean = mean(regular)) %>% 
  arrange(-mean)

@

\section{Implementation of the Generalized Context Model\label{appendixgcm}}

\subsection{Outline}

Our implementation of the GCM evaluates the competition between two categories, {\em regular} and {\em irregular}, for each nonce verb base form. The framework of \cite{nosofsky1990relations} is adapted to morphophonology by using a segmental similarity calculation based on natural classes \citep{frisch1997similarity}. The same treatment of segmental similarity is used in the implementations of the GCM in \cite{albright2003rules} and \cite{dawdy2014learnability}. We build on \cite{dawdy2014learnability} in that we define our categories based on formal similarity.

\begin{quote}
 \texttt{The implementation of the GCM is laid out in models/gcm/gcm.html}.
\end{quote}

\subsection{Training data}

Participants are presented with a sequence of nonce verb base forms, and have to pick either a regular or an irregular past tense form for each. The irregular past tense form is pre-determined by the class for the stem, so that, for a given verb, the participants can only choose between the regular past tense form and the irregular past tense form we assigned to the verb. (So, for instance, for {\em splive}, a verb in the DROVE class, they can choose either {\em splived} or {\em splove}, but not {\em splift} or {\em sploven}, etc.) For a given class (such as DROVE verbs), the GCM has a choice between two sets of verb types.

The irregular set consists of verb types in \textsc{Celex} that form their past tense according to the pattern captured by the class (such as an \{\textipa{[aI]},\textipa{[i]}\} $\rightarrow{}$ \textipa{[oU]} alternation). 
The regular set consists of verb types that have base forms that are similar to these irregular forms but have a regular {\em -ed} past tense form as well as miscellaneous regular verbs -- those that do not belong to our schemata . We narrow the regular set to monosyllabic forms. However, all polysyllabic irregular forms that could serve as a point of comparison are compounds based on monosyllabic forms (an example is {\em overwrite}, a compound form of irregular {\em write}). (A compound form might be more regular than a simplex form, but \textsc{Celex} will list both the regular and the irregular variant in both cases.) Table \ref{regexesforgcm} shows the descriptions of the verb classes using regular expressions.

\begin{table}[ht]
\centering
\begin{tabular}{lllll}
\hline
& \multicolumn{2}{l}{input} & alternation \\
class & regular expression & IPA & \\
\hline
DROVE   & \texttt{[i2][zvdltnk]\$}  & $\{$\textipa{i},\textipa{aI}$\}$ + $\{$z,v,d,l,t,n,k$\}$ \#\# & $\{$\textipa{i},\textipa{aI}$\}$ $\rightarrow{}$ \textipa{oU}\\
SANG    & \texttt{I(m$|$N$|$Nk)\$}  & $\{$\textipa{I}$\}$ + $\{$m,\textipa{N},\textipa{Nk}$\}$ \#\# & \textipa{I} $\rightarrow{}$ \textipa{\ae} \\
KEPT    & \texttt{i[lpnm]\$}       & $\{$\textipa{i}$\}$ + $\{$l,p,n,m$\}$ \#\# & \textipa{i} $\rightarrow{}$ \textipa{E}Ct \\
BURNT   & \texttt{[3EI]nl\$}     & $\{$\textipa{3},\textipa{E},\textipa{I}$\}$ + $\{$n,l$\}$ \#\# & $\{$\textipa{3},\textipa{E},\textipa{I}$\}$ $\rightarrow{}$ $\{$\textipa{3},\textipa{E},\textipa{I}$\}$Ct \\

\hline
\end{tabular}
\caption{Descriptions of verb classes in the GCM. `C' marks any consonant.}
\label{regexesforgcm}
\end{table}

<<celex_counts, echo=F, eval=T>>=

# celex(f>=10) has repetitions:
celex_count = celex$disc %>% unique %>% length

gcm_table = celex %>% # take celex
  filter(!is.na(ahpa)) %>% # remove words not in Albright and Hayes celex
  mutate(
    vowels = str_extract_all(ahpa, '[euVoEIi2U{Q34@a6}]'), # match vowels in the albright and hayes transcription
    nvowels = map(vowels, ~ length(.)) # calculate how many vowels the verb has
  ) %>% 
  select(-vowels) %>% # we don't need vowels now
  unnest(cols = c(nvowels)) %>%  # unnest nvowels
  filter(word != 'forsake', nvowels == 1) %>% # drop any verb with more than 1 vowel also "forsake which survives this
  select(category, moder, disc) %>% # keep category, moder, disc columns (= reg/irreg, drove/sang etc, transcription)
  distinct() %>% # repetitions! 
  count(category, moder) %>% # count drove/sang etc across reg/irreg
  mutate( # if it has no category, it's a misc verb
    moder = replace_na(moder, 'other')
  ) %>% 
  pivot_wider( # stretch table
    names_from = category, values_from = n
  ) %>% 
  rename(
   `verb class` = moder,
    `irregular set` = irregular,
    `regular set` = regular
  )

reg_count = filter(gcm_table, `verb class` == 'other') %>% 
  pull(`regular set`)

gcm_table = gcm_table[1:4,] %>% 
  mutate(
    `miscellaneous regular verbs` = reg_count
  )

gcm_table = rbind(
  gcm_table,
  c('sum', 
    sum(gcm_table$`irregular set`),
    sum(gcm_table$`regular set`),
    unique(gcm_table$`miscellaneous regular verbs`)
    )
)

@

Our starting point for the training set, following \cite{albright2003rules}, is the list of verbs in the \textsc{Celex} corpus (\citealt{baayen1993celex}, based on \citealt{sinclair1987}) with a token frequency of 10 or above, encompassing \Sexpr{celex_count} forms. However, similarity requirements restrict the respective training sets. We use the {\em DISC} transcription in which each contrastive segment of English is represented by a unique character (\texttt{dr2v} equals \textipa{[draIv]}).

Table \ref{gcmsetsize} shows the number of verbs in \textsc{Celex} that were used as training sets for our verb classes. The irregular set consists of forms that match the schema and are irregular. The regular set contains schema matches which are regular, in addition to miscellaneous regulars. The miscellaneous set consists of monosyllabic verbs that do not belong to any of the schemata and are regular. These are included in the regular training set of \textsc{ each} verb class.

The model calculates the similarity of a given nonce verb to the regular and the irregular set. Comparisons to stems in other classes are not calculated, as past tense markings for these classes were not available to participants in the forced-choice tasks.

<<gcmtable, echo=F, eval=T, results=tex>>=
print(xtable(gcm_table, digits=2, label="gcmsetsize", caption="Number of forms in each verb class, GCM training data"))

@

\subsection{Estimation}

To calculate the similarity between two words, we first compute their dissimilarity. This is achieved using the string-edit (Levenshtein) distance, which is the smallest number of changes needed to transform one word into the other. Costs range from 0 (the corresponding segments are identical) to 1 (inserting or deleting an entire segment). Following \cite{albright2003rules} and \cite{dawdy2014learnability}, costs between 0 and 1 are assigned to corresponding segments that are not identical, based on how much the segments differ.

All parts of the word are weighted equally, because although there is evidence that past tense formation in English is predominantly driven by overlaps in word endings, onsets also play a role. (cf. the predominance of \emph {s}(+stop) onsets in irregular verbs forming the past tense with a vowel change, e.g.\ {\em stink}, {\em sink}, etc. -- see \citealt{bybee1983morphological}).

The following transformation, originating with \cite{nosofsky1990relations}, is used to convert dissimilarity into similarity:

\vspace{1cm}
$\eta_{ij} = \exp {(- d_{ij} /s)^p}$
\vspace{1cm}

In the equation above, $\eta_{ij}$ represents the similarity between form $i$ and form $j$, while  $d_{ij}$ is the dissimilarity between the two forms. $s$ and $p$ are free parameters.

We explored a range of parameter settings and use $s = 0.9$ and $p = 1$ which provide the best model fit on the baseline data. (In contrast, Albright and Hayes use $s = 0.4$ and $p = 1$.)

When {\em p} is set to 1, as here, the similarity function is an exponential, rather than a Gaussian, function of the dissimilarity. The weighting parameter {\em s} controls how quickly the similarity decreases as the difference (or distance) between the forms increases. When {\em s} is small, the behavior of the model will be dominated by the small group of instances that differ very little from any given novel form. As it becomes larger, instances that differ more increase their influence on the overall model behavior
 \citep{nosofsky1990relations,nakisa2001dual,albright2003rules,dawdy2014learnability}. Thus, {\em s} effectively controls the size of the set of verbs that will be taken into account in determining the support for the regular versus the irregular outcome.

The overall similarity $S_{iC_J}$ of a test form $i$ to a set
$C_J$ is calculated by summing the similarity $\eta_{ij}$ of each member $j$ of class $C_J$ to the test form $i$, and dividing by the summed similarity $\eta_{ij}$ of each member $k$ of class $C_K$ (the class of all stored forms) to the test form $i$. 
This calculation is summarized in the following equation.

\vspace{1cm}
$S_{iC_J} = \frac{{\sum_{j \in C_J } \eta_{ij} } }{{\sum_{k\in C_K } \eta_{ik} }}$
\vspace{1cm}

\subsection{Output format}

The overall score used in our analyses is the {\em regularity score}, which is the complement to the {\em irregularity score} and reaches a maximum of 1.0 when the output is most likely to be regular. Unlike \cite{dawdy2014learnability}, there is no decision rule on top of the scoring, such that any form that is more likely than not to be regular is predicted to surface as regular all the time. This specific decision rule is statistically optimal, and was imposed in \cite{dawdy2014learnability} in order to determine the ceiling performance for a computational model. The present paper, in contrast, analyzes data aggregated across human participants with differing decision thresholds. As discussed in \cite{schumacher2014reconciling} and \cite{schumacher2017prior}, the input-output relationship in such aggregated data are typically reported to be nearly probability-matching.

We standardize the regularity score to match the range of participant responses: [0,1]. The modified score is interpretable as the probability that the outcome will be regular in aggregated data. It is also appropriate to attribute this type of gradience to people's initial expectations about other people's behavior, on the assumption that people realistically encode the variability they have encountered.

\subsection{Example: {\em splive}}

<<splive1, echo=F, eval=T>>=

irreg_set = gcm_table %>% 
  filter(`verb class` == 'DROVE') %>% 
  pull(`irregular set`) %>% 
  as.double

reg_set = gcm_table %>% 
  filter(`verb class` == 'DROVE') %>% 
  pull(`regular set`) %>% 
  as.double

splive_gcm = test2 %>% 
  mutate(
    c.baseline_gcm_features = scales::rescale(baseline_gcm_features)
  ) %>% 
  filter(word == 'splive') %>% 
  pull(c.baseline_gcm_features) %>% 
  unique %>% 
  round(2)

@

%The nonce form {\em splive} belongs to the DROVE class in our model. The two past forms of {\em splive} in the experiment are regular {\em splived} and irregular {\em splove}. It is compared to \Sexpr{reg_set + reg_count} regular verbs -- these are \Sexpr{reg_set} verbs that match the DROVE schema (e.g.\ {\em side, hive, line}) and \Sexpr{reg_count} miscellaneous verbs. It is also compared to \Sexpr{irreg_set} irregular verbs (e.g.\ {\em drive, stride, smite}) in this class. Overall, it is more similar to the regular set: its {\em regularity score} is \Sexpr{splive_gcm}.

\section{Implementation of the Minimal Generalization \\ Learner}\label{appendixmgl}

\subsection{Outline}

The Minimal Generalization Learner is an algorithm for forming input-output rules of varying generality, which then compete to generate the output.

The Minimal Generalization Learner is implemented here from materials made available by Albright and Hayes \citep{albright2003rules}. These include their Segmental Similarity Calculator, implementing the natural class based similarity metric due to \cite{frisch1997similarity}, also used in the GCM implementation. Due to issues with the MGL code, we had to fit the MGL separately for participants.

\begin{quote}
 \texttt{The implementation of the MGL is laid out in models/mgl/mgl.html}.
\end{quote}

\subsection{Training data}

For our model fitted on our baseline nonce word stimuli, the MGL is trained on regular and irregular English verbs with a minimum frequency cutoff of 10 in \textsc{Celex} \citep{baayen1993celex}, encompassing 4160 past/present verb transcriptions.

The MGL builds rules based on all verb forms in \textsc{Celex} with a token frequency of 10 or above. However the structural descriptions of the resulting rules do not cover all these forms. Table \ref{mglsetsize} shows the number of unique forms covered by the structural descriptions of the `regular' and `irregular' rules that are relevant to each class.

The MGL generates multiple possible past tense forms for each nonce verb. We only consider those rules {\em relevant} that generate the past tense forms that appear in the experiment (e.g.\ {\em splive} : {\em splived} / {\em splove}). There is at most one relevant regular rule and one relevant irregular rule for one verb, but multiple rules can generate the (ir)regular forms for each verb class. We return to this in the next section.

Note that the sets of exceptions and related forms of each rule can overlap, both respectively and with each other. As a consequence, the MGL rules apply to fewer forms than apparent from the table: 456 (instead of 617) in total.

<<mgl_form_counts, eval=T, echo=F>>=

# in order to look up the number of words per rule type, we need to dig up the raw mgl outputs.
# we have a function that turns them into mgl regular weights. but we need to break up that functions part-way and print the output here.

# get the raw output of the baseline feature-distance mgl
mgl_baseline_rules_raw = read_tsv('~/Github/RaczBecknerHayPierrehumbert2019/models/mgl/baseline_mgl/CELEXFull3.sum') %>% 
  mutate(
    form1 = str_replace_all(form1, '»', ''), # get rid of syllable boundary markers
    form2 = str_replace_all(form2, '»', '')
  )

disc_category = test2 %>% # get words and categories from test2. match colnames with above
  distinct(disc, category) %>% 
  mutate(
    form1 = disc
  )

verb_handles = read_csv('~/Github/RaczBecknerHayPierrehumbert2019/data/256verbs_print_forms.txt')

disc_reg_past_irreg_past = verb_handles %>% # get reg and irreg transcriptions from verb handles. match colnames with above.
  select(base.form,regular.form,irregular.form) %>% 
  rename(
    form1 = base.form
  )

mgl_baseline_rules_raw = mgl_baseline_rules_raw %>% # join them up
  left_join(disc_category, by = "form1") %>% 
  left_join(disc_reg_past_irreg_past, by = "form1")

# at this point, we only have categories etc for verbs in our ESP game. but since the interesting rules are those that apply to these forms, we focus on them only. so 156 verbs in total.

mgl_baseline_rules_raw = mgl_baseline_rules_raw %>% 
  filter(
    !is.na(category)
  )

# mgl_count$form1 %>% unique %>% length

# mind you, not all of these verbs have both regular and irregular rules, as discussed in the text. this is handled by the function processing the mgl output -- we won't deal with it here, since it's not relevant

mgl_baseline_rules_raw = mgl_baseline_rules_raw %>%  # group rules per output: reg / irreg!
  mutate(
    rule.type = case_when(
      form2 == regular.form ~ 'regular',
      form2 == irregular.form ~ 'irregular'
    )
  )

mgl_baseline_rules_raw = mgl_baseline_rules_raw %>%
  filter(
!is.na(rule.type)
)

# mgl_count$form1 %>% unique %>% length

# this is where it gets tricky. you can't sum up the hits and scope, because these overlap. you need to bin related forms and expressions per category per rule type and count them.

# first, we split the list-like strings into character vectors

mgl_count = mgl_baseline_rules_raw %>% # clean-up
  mutate(
    `related forms` = str_replace_all(`related forms`, '»', ''),
    exceptions = str_replace_all(exceptions, '»', ''),
    related_list = strsplit(`related forms`, ', '), # split
    exception_list = strsplit(exceptions, ', ')
  )

mgl_count1 = mgl_count %>%
  select(category,rule.type,related_list) %>% 
  unnest(related_list) %>% 
  distinct() %>% 
  count(category, rule.type) %>% 
  mutate(
    type = 'related forms'
  )

mgl_count2 = mgl_count %>%
  select(category,rule.type,exception_list) %>% 
  unnest(exception_list) %>% 
  distinct() %>% 
  count(category, rule.type) %>% 
  mutate(
    type = 'exceptions'
  )

mgl_table = rbind(mgl_count1,mgl_count2) %>% 
  pivot_wider(
    names_from = type, values_from = n
  ) %>% 
  arrange(rule.type, category)

@

<<mgltable, echo=F, eval=T, results=tex>>=
print(xtable(mgl_table, digits=2, label="mglsetsize", caption="Number of forms in each verb class, MGL training data"))

@

\subsection{Estimation}

The MGL begins by considering the relationship between each verb and its past tense as a `rule'. For each pair of verbs in the training data, it then attempts to create a more general rule. It does so by aligning the wordforms and analyzing shared phonetic features. For example,  merging the word-specific rules for {\em ring/rang} and for {\em stink/stank} yields a more general rule that expresses the information that they share: \textipa{[I]} $\rightarrow{}$\textipa{[\ae]} / [+coronal, - cont] \underline{\hspace{0.5cm}} \textipa{[N]}. Each rule inferred in this way is then further generalized on the basis of more comparisons; for instance, taking note of {\em swim/swam} expands the \textipa{[I]} $\rightarrow{}$\textipa{[\ae]} rule to specify that it occurs before all [+nasal] consonants.

The structural description for each rule has a {\em scope}, which is the number of verbs conforming to the description, to which the rule might apply. The number of {\em hits} is the number of such verbs where the rule generates the correct output. In our example, {\em think} and {\em blink} fall in the scope of the rule, but they are not hits, because their past tenses display other patterns ({\em thought} and {\em blinked})
The {\it raw confidence} of the rule is the ratio of hits to scope:

\vspace{1cm}
$\mbox{\it Raw confidence} = \frac{\mbox{\it  hits}} {\mbox{\it  scope}}$
\vspace{1cm}

The raw confidence is 1.0 if the rule applies to all forms that meet its structural description.  It is less than 1.0 if some forms meeting its structural description have past tenses other than that predicted by the structural change. Raw confidence values of 0 are not found, because a rule needs to apply to two or more examples to be posited in the first place.

The MGL raw confidence metric is adjusted on the basis of user-specified confidence limits, to generate an {\em adjusted confidence score} that takes into account the amount and distribution of available data. The MGL's lower limit affects how much confidence is assigned to rules that have a small number of instances; generalizations that are based on a smaller number of word types are penalized. The MGL's upper limit curtails the application of seemingly general rules which are in fact driven by a more specific rule \citep{albright2002modeling}. The MGL is implemented here with its default settings, with the exception of the algorithm's confidence limits. We implement the MGL with lower and upper confidence limits of 55\% and 95\%, respectively, since these values afford the best fit to English verb data in \cite{albright2003rules}.

Note that the MGL algorithm automatically groups together verbs on the basis of shared phonological properties; thus, verbs are most likely to form strong generalizations with other verbs that share the same onset or rhyme. Attempts to merge diverse wordforms under a single generalization would be more likely to incur penalties (i.e.\ exceptions). This feature of the MGL is important for comparing with the methods of the GCM. Both algorithms allow for category-specific similarities to play a role in rule formation.

\subsection{Output format}\label{mgloutput}

Recall that in both our baseline and ESP experiment the trial task is prompted by a stem and offers a choice between a regular form and a specific irregular form, presented orthographically. In order to model this choice, we take the MGL rule for the stem that outputs the regular form (the {\em relevant regular rule}) and the rule that outputs the specific irregular form (the {\em relevant irregular rule}). If several regular / irregular rules generate the same form, we take the one with the highest {\em adjusted confidence}, following \cite{albright2003rules}. We use these rules to calculate the form's {\em relative (adjusted) confidence}.

Out of 156 test verbs in the ESP post-test, the \textsc{Celex}-trained MGL generates a {\em relevant regular rule} for every verb. It does not generate a {\em relevant irregular rule} for 28 verbs. These are all nonce verbs in the \textsc{KEPT} category (see\ Section 4.1 in the main text). In this category, irregular forms are derived from the stem through a vowel change (e.g. {\em greel} $\rightarrow{}$ {\em grelt}). The verbs missing the relevant irregular rule all have bases ending in $<$m$>$, $<$n$>$, or $<$l$>$. This is because, in our implementation, there is an insufficient number of verb types in the training set to support these irregular rules. Decreasing the cutoff criterion for the model leads to the generation of more of the currently `missing' irregular rules, but the overall model fit becomes worse (cf.\ below). Therefore, we keep the cutoff criterion and assume that the adjusted confidence of the irregular rule for these 28 verbs is zero. We then take the relative confidence of the regular rule as compared to the regular {\em and} the irregular rule for each verb and took this as the adjusted regular confidence of the given verb. (If the irregular rule is missing, the value of this adjusted regular confidence is 1.) This is given by the following equation:

\vspace{1cm}
$\mbox{\it Relative (adjusted) confidence} = \frac{\mbox{\it  adjusted confidence of relevant regular rule}} {\mbox{\it  adj.\ conf.\ reg.\ rule } + \mbox{ \it adj.\ conf.\ relevant irreg.\ rule}}$
\vspace{1cm}

This relative adjusted confidence represents the MGL regularity score for an item, to be compared against the regularity score from the GCM (see Section \ref{appendixgcm}).

\subsection{Example: {\em splive}}

<<splive2, echo=F, eval=F>>=

# In the next par, I won't pipe the values into variables in the text, but you can look them up here:
splive_mgl_rules = mgl_baseline_rules_raw %>% 
  filter(
    form1 == 'spl2v'
  )
# regular
splive_mgl_rules[3,]$Pfeat
splive_mgl_rules[3,]$hits
splive_mgl_rules[3,]$scope
splive_mgl_rules[3,]$reliability
splive_mgl_rules[3,]$confidence

# irregular
splive_mgl_rules[1,]$Pfeat
splive_mgl_rules[1,]$hits
splive_mgl_rules[1,]$scope
splive_mgl_rules[1,]$`related forms`
splive_mgl_rules[1,]$exceptions
splive_mgl_rules[1,]$reliability
splive_mgl_rules[1,]$confidence

test2 %>% 
  mutate(
    c.baseline_mgl_features = scales::rescale(baseline_mgl_features)
  ) %>% 
  filter(word == 'splive') %>% 
  select(baseline_mgl_features,c.baseline_mgl_features) %>% 
  distinct()

@

The two past forms of {\em splive} in the experiment are regular {\em splived} and irregular {\em splove}. The {\em relevant regular rule} that generates the regular past tense is `$\emptyset{} \rightarrow{}$ \textipa{[d]} / \{\textipa{D}, \textipa{S}, \textipa{T}, \textipa{Z}, f, s, v, z\} \underline{\hspace{0.5cm}}'. The {\em structural description} indicates that this is a suffixation rule that can apply to forms that {\em end in an anterior fricative} (a natural class in our feature system).  The {\em raw confidence} of this rule is 0.98. This is because this rule applies to most forms in its scope (698/712). The {\em adjusted confidence} is very similar: 0.968. This is because this rule applies to a large number of forms overall. The {\em relevant irregular rule} that generates the irregular form is `\textipa{[aI]} $\rightarrow{}$ \textipa{[oU]} / \{\textipa{D}, \textipa{Z}, \textipa{dZ}, d, l, n, r, z\} \underline{\hspace{0.5cm}}v'. It applies to \textipa{[aI]} in the nucleus {\em preceded by voiced anterior consonant and followed by \textipa{[v]}}. In \textsc{Celex},the rule applies to three forms ({\em drive, strive, dive}) and fails to apply to five ({\em arrive, thrive, contrive, rive, connive}). Its {\em raw confidence} is 0.375. Its {\em adjusted confidence} is slightly lower (0.366). This is because it applies to a smaller number of forms overall. The {\em relative (adjusted) confidence} of the predicted regularity of {\em splive} is $0.98/(0.98+0.37)=0.73$. The respective rescaled value is 0.363.

\section{Further notes on the GCM and the MGL \label{appendixgcmvsmgl}}

The training set for the GCM is focused initially; it is grouped into four verb classes and based on formal similarity with base forms of existing irregular verbs playing a role within each class. This, in effect, assumes lexical gangs as a starting point \citep{alegre1999rule}. In contrast, the MGL starts establishing rules across all forms in the starting dictionary.

As we see, however, the MGL also focuses the training set. The structural descriptions of the forms only cover a fraction of all verbs in the starting dictionary, organized by shared groups of segments. While disjunction increases the power of a theory (increasing the set of acceptable classes), disjunct classes frequently serve as input for phonological processes \citep{mielke2008emergence} . In the MGL, the disjunction emerges despite the lack of an initial specification: while one rule accounts for the specific behavior of a given form, multiple rules cover a given class. The number of regular and irregular rules posited by the MGL for each verb class can be seen in Table \ref{mglruleses}.

<<mgl_rule_counter, eval=T, echo=F>>=

mgl_rule_count = mgl_baseline_rules_raw %>% # take raw rules
  filter( # throw out non-category verbs and non reg/irreg rules (in our definition)
    !is.na(category),
    !is.na(rule.type)
  ) %>% 
  mutate( # make a rule column of the form A -> B / C
    rule = paste0(A, ' ->', B, ' / ', Pfeat)
  ) %>% 
  distinct(rule,category,rule.type) %>% # keep DISTINCT rules
  count(category, rule.type) %>%  # count them
  pivot_wider(names_from = rule.type, values_from = n) %>% # stretch table
  mutate( # make row sum
    sum = irregular + regular
  ) %>% 
  rename(
    `regular rule` = regular,
    `irregular rule` = irregular
  )

# glue on colsums
mgl_rule_count = rbind(mgl_rule_count, c('sum', sum(mgl_rule_count$`irregular rule`), sum(mgl_rule_count$`regular rule`), sum(mgl_rule_count$sum)))

@

<<mgltable, echo=F, eval=T, results=tex>>=
print(xtable(mgl_table, digits=2, label="mglruleses", caption="The number of unique rules posited by the MGL in each verb class"))

@

Table \ref{mglruleseskept} shows the list of (irregular and regular) rules for one of the classes, the KEPT class. We choose this class as it provides a good illustration of the estimation process. An MGL rule describes an $A \rightarrow{} B$ {\em alternation} in an $XAY$ {\em environment}. Since the KEPT rules are all suffixing, the following environment ($Y$) is empty. For the preceding environment ($X$), we give a list of segments that are matched by the structural description if it applies to four segments or fewer. Otherwise, we give a phonological description. So, for instance, the first rule turns \textipa{[il]} to \textipa{[Elt]} following any obstruent. Note that in some cases, the MGL generates multiple rules for the same alternation (such as \textipa{ip} $\rightarrow{}$ \textipa{Ept}). As discussed in Section \ref{mgloutput}, in such instances, we choose the candidate rule with the highest adjusted confidence in order to calculate the MGL regularity score. (Fricatives and nasal consonants constitute a natural class in the segmental feature system used by \citealt{albright2003rules}.)

Inspection of the table shows that the MGL fails to generate some expected irregular rules; in the KEPT class, the MGL omits rules for any stems ending in \textipa{[in]} or \textipa{[im]}. Under the MGL, there is not enough lexical support for irregular rules of these types, and yet participants do irregularize nonce forms like {\em cheen} and {\em kleem}  in the forced choice task. In this respect, the variegated similarity of the non-natural GCM class (which can be used to categorize {\em cheen}) allows for a better approximation of participant behavior. Note also that in some cases, the MGL generates an overly-specific rule; the rule for stems ending in [il] applies only after obstruents, whereas participants also irregularize KEPT stems with a sonorant in the onset ({\em greel}/{\em grelt}).

\begin{table}[ht]
\centering
\begin{tabular}{lll}
  \hline
alternation & preceding environment\\ 
  \hline
\textipa{il} $\rightarrow{}$ \textipa{Elt} & obstruent \\ 
\textipa{ip} $\rightarrow{}$ \textipa{Ept} & \{j, l, r, w\} \\ 
\textipa{ip} $\rightarrow{}$ \textipa{Ept} & dorsal consonant\\
\textipa{ip} $\rightarrow{}$ \textipa{Ept} & l \\ 
\textipa{ip} $\rightarrow{}$ \textipa{Ept} & any consonant \\
\textipa{il} $\rightarrow{}$ \textipa{Elt} & fricative or nasal consonant \\
$\emptyset{}$ $\rightarrow{}$ \textipa{d}    & alveolars \\ 
$\emptyset{}$ $\rightarrow{}$ \textipa{d}    & \{b, m\} \\ 
$\emptyset{}$ $\rightarrow{}$ \textipa{d}    & alveolar fricative or nasal consonant \\ 
$\emptyset{}$ $\rightarrow{}$ \textipa{d}    & \{b, m, p\} \\
   \hline
\end{tabular}
\caption{The unique rules posited for the KEPT class}
\label{mglruleseskept}
\end{table}

The rule structure of the MGL likely follows from its parameter settings, which were tuned to optimize its performance on our fairly irregular-looking nonce word stimuli. It requires further work to determine the extent to which the MGL fit on individuals would improve if the MGL parameters were customized for individual participants.\\

ITT TARTOK

\section{Residualization of the two categorization models on the baseline data}\label{appendixgcmmglresidualization}

Since the regularity scores of the GCM and the MGL for the nonce verbs are correlated  (r $=$ 0.48), we use residualization to see whether the two models make separate significant contributions to explaining variation in the baseline dataset (Section 6.3 in the main text.)

We fit two simple linear regression models; no random effects are explored in these models, since we are merely predicting algorithmic scores for each item. We rescale and center GCM and MGL scores for the model fits.

Model 1 predicts the GCM scores from the MGL scores (M1: GCM regular score  $\sim$ MGL regular score). Model 2 predicts the MGL scores from the GCM scores (M2: MGL regular score  $\sim$ GCM regular score). In two additional models, we then use the residuals from each model (M1, M2) as an estimate of the variation present in one score that cannot be attributed to the other.  First, we use the residuals of Model 1 in another, mixed-effects logistic regression model along with the MGL scores to predict regular responses in the baseline task (M3: regular response  $\sim$ MGL score + GCM residual + (1 + MGL score + GCM residual$|$participant) + (1 $|$ verb)). We then residualize for the MGL. We use the residuals of Model 2 along with the GCM scores in a similar logistic regression model (M4: regular response  $\sim$ GCM score + MGL residual + (1 + GCM score + MGL residual $|$participant)).

 The residuals remain significant predictors in both models (M3: MGL score ($\beta$ $=$ 1.27, p $<$ 0.0001), GCM residual ($\beta$ $=$ 1.37, p $<$ 0.0001); M4: GCM score ($\beta$ $=$ 2.29, p $<$ 0.0001), MGL residual ($\beta$ $=$ 0.95, p $<$ 0.0001). 
 Note that while this paper reports values fit to the baseline experiment data, the same pattern of results also holds for models fit to the pre-test data.

Although residualization is sometimes critiqued when used as a stopgap for collinearity or sign changes in a regression model \citep{wurm2014residualizing}, these concerns do not apply in the present case.  We use residualization as a tool to test whether one metric contributes beyond what is covered by a different metric of the same attribute, as in 
\cite{baayen2006morphological}, and as also acknowledged in \cite{wurm2014residualizing}.

\section{Additional information regarding  the role of the GCM and the MGL in the post-test \label{posttestimplementation}}

\subsection{Concordance Indices}
Table \ref{modellingsummary} summarizes the concordance indices between our data and the various categorization models. The individual-GCM improves on the \textsc{Celex}-GCM in terms of explaining the post-test data.   The individual-MGL does not improve on the \textsc{Celex}-MGL.

\begin{table}[h]

    \centering
    \begin{tabular}{lllll}
            & training data & method & test data & C (test data)\\
            \hline
\textsc{Celex}-GCM & real verbs & `analogy' & verbs in baseline &{\bf 0.58 }\\
\textsc{Celex}-MGL & real verbs & `rules' & verbs in baseline &{\bf  0.59 }\\ \hline
\textsc{Celex}-GCM & real verbs & `analogy' & verbs in post-test &{\bf  0.6}\\
individual-GCM & real verbs + verbs seen in ESP & `analogy' & verbs in post-test &{\bf  0.68}\\

\textsc{Celex}-MGL & real verbs & `rules' & verbs in post-test &{\bf  0.61}\\
individual-MGL & real verbs + verbs seen in ESP & `rules' & verbs in post-test &{\bf  0.55}\\
\hline
    \end{tabular}
    \caption{Summary of concordance indices between our data and the categorization models. The values above the line show the concordance indices for the baseline data, as reported in Section 6.3 of the main text. The values below the line show the indices for the ESP post-test data.}
    \label{modellingsummary}

\end{table}

 Our concordance indices are calculated using the somers2 function in R's \texttt{Hmisc} package \citep{harrell2014hmisc}.

 Note that the indices in Table \ref{modellingsummary} directly calculate C between the relevant algorithm score (MGL or GCM) and the binary post-test outcome (regular or irregular). As a diagnostic for our regression models, we also calculated concordance indices between model predictions and participants' regular/irregular responses. The concordance indices for multiple regression models are, unsurprisingly, considerably higher than the indices calculated on the basis of a single variable.

 The models in Tables 5 and 6 of the main text both have an index of 0.84, indicating an acceptable level of predictive ability \citep{harrell2013regression}.

 Since the random effects likely account for a large amount of this variance, we calculated the marginal r${^2}$ values for both models: these are 0.27 and 0.22, respectively (the conditional values are 0.45 and 0.47). Conditions and learning models explain some amount of the variation, but the data 
 are noisy.

\subsection{Assessing separate contribution of individual models over \textsc{Celex} models}
The text (Section 6.4 in the main text) explains that we calculate the extra information given by the individual models by subtracting the \textsc{Celex} model prediction from the individual model prediction for each item for each participant, and that we find that the extra information from the individual-GCM  models is predictive over and above the \textsc{Celex}-GCM model.

We do this for both the GCM  (Model 5: regular response  $\sim$ \textsc{Celex} GCM score + extra GCM information + participant pre-test mean + (1 + \textsc{Celex} GCM + extra GCM information$|$participant) + (1$|$verb) and the MGL (Model 6: regular response  $\sim$ \textsc{Celex} MGL score + extra MGL information + participant pre-test mean + (1 + \textsc{Celex} MGL + extra MGL information$|$participant)+ (1$|$verb)).

In Model 5, the extra information from the individual-GCM is a significant predictor of post-test responses  ($\beta$ $=$ 2.61, p $<$ 0.001),  above and beyond the \textsc{Celex}-GCM contribution. In similar Model 6, however, the extra information from the individual-MGL is not a significant predictor of post-test responses ($\beta$ $=$ 0.21, p $=$ 0.23). This means that the individual-GCM has some success in modeling post-test behavior when trained with real English verbs and information from the ESP test, whereas adjustments made to the individual-MGL do not improve predictive power.

\section{Regression modeling procedures}\label{appendixregression}

Our regression model selection in this article proceeded as follows. Except where stated otherwise, we fit logistic mixed-effects regression models, using R's lme4 package \citep{bateslme4}. Non-significant predictors are removed from our models, as determined by a series of comparisons of nested models, using likelihood ratio tests in R's anova function. However, our model tables (e.g., Tables 5 and 6 in the main text) include the significance levels generated by lme4 summaries.

We investigate maximal random effects structures, allowing for random intercepts for participants and stimulus items (verbs), as well as random slopes for all within-unit factors. We compare nested random effects structures as above, and retain any random slopes and intercepts which are supported by likelihood ratio tests \citep{baayen2008mixed}.

In the model in Table 5 in the main text, we elect to mean center continuous variables to counteract collinearity that becomes evident when including the interaction between and {\em lexical typicality} .  Although some debate exists regarding the practice of transforming variables \citep{belsley1991conditioning,echambadi2007mean}, centering variables is common practice to remove nonessential ill conditioning, that is, relationships that inevitably exist between main and product terms \citep{aiken1991multiple,jaccard2003interaction, jaeger2008categorical}. We note that if {\em verb baseline mean} is {\em not} centered, collinearity is high (VIF=15.79) in this model. A non-centered model also yields a significant main effect for {\em lexical typicality}, with increased post-test regularization in the random and reversed conditions. However, we elect to take a more conservative approach by not including a potentially spurious main effect. Note that other than this main effect difference, models with and without centering yield results that are qualitatively indistinguishable; all other main effects and interactions are present in either case. The collinearity for the centered model (Table 5 in the main text) is acceptable, with a maximum VIF score of 2.89.

Since it is unclear whether the absence of a difference between random and reversed conditions is due to a lack of data, we refit the model with the baseline average : typicality interaction on the random and reversed conditions only. Then we calculated the Bayes Factor of the null model (with no interaction) over the full model (with interaction) using Bayesian Information Criteria calculated by lme4. The result strongly supports the null model over the full model, indicating that the preference for the null model is {\em not} due to the lack of evidence to reject it.

\bibliographystyle{language}
\bibliography{racz_beckner_hay_pierrehumbert_2020}

\end{document}

### 2020-01-28 supplementary_information/si.tex
in progress

--- 2020-01-28 supplementary_information/si.Rnw
@@
 
+\usepackage{Sweave}
 \begin{document}
-\SweaveOpts{concordance=TRUE}
-
-<<header, echo=FALSE, eval=TRUE>>=
-
-# libs
-library(tidyverse)
-library(xtable)
-
-# data
-source('~/Github/RaczBecknerHayPierrehumbert2019/functions/grabdata.R')
-
-@
+\input{si-concordance}
 
@@
 We have trialed a set of possible small changes and found that they have no major effect on our overall results.
-
-<<header, echo=FALSE, eval=FALSE>>=
-
-# table 1
-baseline %>% 
-  group_by(word) %>% 
-  summarise(mean = mean(regular)) %>% 
-  arrange(-mean)
-
-@
 
@@
 
-<<celex_counts, echo=F, eval=T>>=
-
-# celex(f>=10) has repetitions:
-celex_count = celex$disc %>% unique %>% length
-
-gcm_table = celex %>% # take celex
-  filter(!is.na(ahpa)) %>% # remove words not in Albright and Hayes celex
-  mutate(
-    vowels = str_extract_all(ahpa, '[euVoEIi2U{Q34@a6}]'), # match vowels in the albright and hayes transcription
-    nvowels = map(vowels, ~ length(.)) # calculate how many vowels the verb has
-  ) %>% 
-  select(-vowels) %>% # we don't need vowels now
-  unnest(cols = c(nvowels)) %>%  # unnest nvowels
-  filter(word != 'forsake', nvowels == 1) %>% # drop any verb with more than 1 vowel also "forsake which survives this
-  select(category, moder, disc) %>% # keep category, moder, disc columns (= reg/irreg, drove/sang etc, transcription)
-  distinct() %>% # repetitions! 
-  count(category, moder) %>% # count drove/sang etc across reg/irreg
-  mutate( # if it has no category, it's a misc verb
-    moder = replace_na(moder, 'other')
-  ) %>% 
-  pivot_wider( # stretch table
-    names_from = category, values_from = n
-  ) %>% 
-  rename(
-   `verb class` = moder,
-    `irregular set` = irregular,
-    `regular set` = regular
-  )
-
-reg_count = filter(gcm_table, `verb class` == 'other') %>% 
-  pull(`regular set`)
-
-gcm_table = gcm_table[1:4,] %>% 
-  mutate(
-    `miscellaneous regular verbs` = reg_count
-  )
-
-gcm_table = rbind(
-  gcm_table,
-  c('sum', 
-    sum(gcm_table$`irregular set`),
-    sum(gcm_table$`regular set`),
-    unique(gcm_table$`miscellaneous regular verbs`)
-    )
-)
-
-@
-
-Our starting point for the training set, following \cite{albright2003rules}, is the list of verbs in the \textsc{Celex} corpus (\citealt{baayen1993celex}, based on \citealt{sinclair1987}) with a token frequency of 10 or above, encompassing \Sexpr{celex_count} forms.
+Our starting point for the training set, following \cite{albright2003rules}, is the list of verbs in the \textsc{Celex} corpus (\citealt{baayen1993celex}, based on \citealt{sinclair1987}) with a token frequency of 10 or above, encompassing 3156 forms.
 However, similarity requirements restrict the respective training sets.
@@
 
-<<gcmtable, echo=F, eval=T, results=tex>>=
-print(xtable(gcm_table, digits=2, label="gcmsetsize", caption="Number of forms in each verb class, GCM training data"))
-
-@
-
+% latex table generated in R 3.6.0 by xtable 1.8-4 package
+% Mon Jan 27 16:35:17 2020
+\begin{table}[ht]
+\centering
+\begin{tabular}{rllll}
+  \hline
+ & verb class & irregular set & regular set & miscellaneous regular verbs \\ 
+  \hline
+1 & BURNT & 6 & 42 & 1218 \\ 
+  2 & DROVE & 14 & 83 & 1218 \\ 
+  3 & KEPT & 12 & 31 & 1218 \\ 
+  4 & SANG & 8 & 13 & 1218 \\ 
+  5 & sum & 40 & 169 & 1218 \\ 
+   \hline
+\end{tabular}
+\caption{Number of forms in each verb class, GCM training data} 
+\label{gcmsetsize}
+\end{table}
 \subsection{Estimation}
@@
 
-<<splive1, echo=F, eval=T>>=
-
-irreg_set = gcm_table %>% 
-  filter(`verb class` == 'DROVE') %>% 
-  pull(`irregular set`) %>% 
-  as.double
-
-reg_set = gcm_table %>% 
-  filter(`verb class` == 'DROVE') %>% 
-  pull(`regular set`) %>% 
-  as.double
-
-splive_gcm = test2 %>% 
-  mutate(
-    c.baseline_gcm_features = scales::rescale(baseline_gcm_features)
-  ) %>% 
-  filter(word == 'splive') %>% 
-  pull(c.baseline_gcm_features) %>% 
-  unique %>% 
-  round(2)
-
-@
-
 %The nonce form {\em splive} belongs to the DROVE class in our model.
 The two past forms of {\em splive} in the experiment are regular {\em splived} and irregular {\em splove}.
-It is compared to \Sexpr{reg_set + reg_count} regular verbs -- these are \Sexpr{reg_set} verbs that match the DROVE schema (e.g.\ {\em side, hive, line}) and \Sexpr{reg_count} miscellaneous verbs.
-It is also compared to \Sexpr{irreg_set} irregular verbs (e.g.\ {\em drive, stride, smite}) in this class.
-Overall, it is more similar to the regular set: its {\em regularity score} is \Sexpr{splive_gcm}.
+It is compared to 1301 regular verbs -- these are 83 verbs that match the DROVE schema (e.g.\ {\em side, hive, line}) and 1218 miscellaneous verbs.
+It is also compared to 14 irregular verbs (e.g.\ {\em drive, stride, smite}) in this class.
+Overall, it is more similar to the regular set: its {\em regularity score} is 0.57.
 
@@
 
-<<mgl_form_counts, eval=T, echo=F>>=
-
-# in order to look up the number of words per rule type, we need to dig up the raw mgl outputs.
-# we have a function that turns them into mgl regular weights. but we need to break up that functions part-way and print the output here.
-
-# get the raw output of the baseline feature-distance mgl
-mgl_baseline_rules_raw = read_tsv('~/Github/RaczBecknerHayPierrehumbert2019/models/mgl/baseline_mgl/CELEXFull3.sum') %>% 
-  mutate(
-    form1 = str_replace_all(form1, '»', ''), # get rid of syllable boundary markers
-    form2 = str_replace_all(form2, '»', '')
-  )
-
-disc_category = test2 %>% # get words and categories from test2. match colnames with above
-  distinct(disc, category) %>% 
-  mutate(
-    form1 = disc
-  )
-
-verb_handles = read_csv('~/Github/RaczBecknerHayPierrehumbert2019/data/256verbs_print_forms.txt')
-
-disc_reg_past_irreg_past = verb_handles %>% # get reg and irreg transcriptions from verb handles. match colnames with above.
-  select(base.form,regular.form,irregular.form) %>% 
-  rename(
-    form1 = base.form
-  )
-
-mgl_baseline_rules_raw = mgl_baseline_rules_raw %>% # join them up
-  left_join(disc_category, by = "form1") %>% 
-  left_join(disc_reg_past_irreg_past, by = "form1")
-
-# at this point, we only have categories etc for verbs in our ESP game. but since the interesting rules are those that apply to these forms, we focus on them only. so 156 verbs in total.
-
-mgl_baseline_rules_raw = mgl_baseline_rules_raw %>% 
-  filter(
-    !is.na(category)
-  )
-
-# mgl_count$form1 %>% unique %>% length
-
-# mind you, not all of these verbs have both regular and irregular rules, as discussed in the text.
-this is handled by the function processing the mgl output -- we won't deal with it here, since it's not relevant
-
-mgl_baseline_rules_raw = mgl_baseline_rules_raw %>%  # group rules per output: reg / irreg!
-  mutate(
-    rule.type = case_when(
-      form2 == regular.form ~ 'regular',
-      form2 == irregular.form ~ 'irregular'
-    )
-  )
-
-mgl_baseline_rules_raw = mgl_baseline_rules_raw %>%
-  filter(
-!is.na(rule.type)
-)
-
-# mgl_count$form1 %>% unique %>% length
-
-# this is where it gets tricky. you can't sum up the hits and scope, because these overlap. you need to bin related forms and expressions per category per rule type and count them.
-
-# first, we split the list-like strings into character vectors
-
-mgl_count = mgl_baseline_rules_raw %>% # clean-up
-  mutate(
-    `related forms` = str_replace_all(`related forms`, '»', ''),
-    exceptions = str_replace_all(exceptions, '»', ''),
-    related_list = strsplit(`related forms`, ', '), # split
-    exception_list = strsplit(exceptions, ', ')
-  )
-
-mgl_count1 = mgl_count %>%
-  select(category,rule.type,related_list) %>% 
-  unnest(related_list) %>% 
-  distinct() %>% 
-  count(category, rule.type) %>% 
-  mutate(
-    type = 'related forms'
-  )
-
-mgl_count2 = mgl_count %>%
-  select(category,rule.type,exception_list) %>% 
-  unnest(exception_list) %>% 
-  distinct() %>% 
-  count(category, rule.type) %>% 
-  mutate(
-    type = 'exceptions'
-  )
-
-mgl_table = rbind(mgl_count1,mgl_count2) %>% 
-  pivot_wider(
-    names_from = type, values_from = n
-  ) %>% 
-  arrange(rule.type, category)
-
-@
-
-<<mgltable, echo=F, eval=T, results=tex>>=
-print(xtable(mgl_table, digits=2, label="mglsetsize", caption="Number of forms in each verb class, MGL training data"))
-
-@
-
+% latex table generated in R 3.6.0 by xtable 1.8-4 package
+% Mon Jan 27 16:35:17 2020
+\begin{table}[ht]
+\centering
+\begin{tabular}{rllrr}
+  \hline
+ & category & rule.type & related forms & exceptions \\ 
+  \hline
+1 & burnt & irregular &  24 &  41 \\ 
+  2 & kept & irregular &  10 &   9 \\ 
+  3 & sang & irregular &  11 &  77 \\ 
+  4 & burnt & regular &  38 &  21 \\ 
+  5 & drove & regular &  29 &  27 \\ 
+  6 & kept & regular &  67 &  41 \\ 
+  7 & sang & regular &  47 &  38 \\ 
+   \hline
+\end{tabular}
+\caption{Number of forms in each verb class, MGL training data} 
+\label{mglsetsize}
+\end{table}
 \subsection{Estimation}
@@
 
-<<splive2, echo=F, eval=F>>=
-
-# In the next par, I won't pipe the values into variables in the text, but you can look them up here:
-splive_mgl_rules = mgl_baseline_rules_raw %>% 
-  filter(
-    form1 == 'spl2v'
-  )
-# regular
-splive_mgl_rules[3,]$Pfeat
-splive_mgl_rules[3,]$hits
-splive_mgl_rules[3,]$scope
-splive_mgl_rules[3,]$reliability
-splive_mgl_rules[3,]$confidence
-
-# irregular
-splive_mgl_rules[1,]$Pfeat
-splive_mgl_rules[1,]$hits
-splive_mgl_rules[1,]$scope
-splive_mgl_rules[1,]$`related forms`
-splive_mgl_rules[1,]$exceptions
-splive_mgl_rules[1,]$reliability
-splive_mgl_rules[1,]$confidence
-
-test2 %>% 
-  mutate(
-    c.baseline_mgl_features = scales::rescale(baseline_mgl_features)
-  ) %>% 
-  filter(word == 'splive') %>% 
-  select(baseline_mgl_features,c.baseline_mgl_features) %>% 
-  distinct()
-
-@
-
 The two past forms of {\em splive} in the experiment are regular {\em splived} and irregular {\em splove}.
@@
 
-<<mgl_rule_counter, eval=T, echo=F>>=
-
-mgl_rule_count = mgl_baseline_rules_raw %>% # take raw rules
-  filter( # throw out non-category verbs and non reg/irreg rules (in our definition)
-    !is.na(category),
-    !is.na(rule.type)
-  ) %>% 
-  mutate( # make a rule column of the form A -> B / C
-    rule = paste0(A, ' ->', B, ' / ', Pfeat)
-  ) %>% 
-  distinct(rule,category,rule.type) %>% # keep DISTINCT rules
-  count(category, rule.type) %>%  # count them
-  pivot_wider(names_from = rule.type, values_from = n) %>% # stretch table
-  mutate( # make row sum
-    sum = irregular + regular
-  ) %>% 
-  rename(
-    `regular rule` = regular,
-    `irregular rule` = irregular
-  )
-
-# glue on colsums
-mgl_rule_count = rbind(mgl_rule_count, c('sum', sum(mgl_rule_count$`irregular rule`), sum(mgl_rule_count$`regular rule`), sum(mgl_rule_count$sum)))
-
-@
-
-<<mgltable, echo=F, eval=T, results=tex>>=
-print(xtable(mgl_table, digits=2, label="mglruleses", caption="The number of unique rules posited by the MGL in each verb class"))
-
-@
-
+% latex table generated in R 3.6.0 by xtable 1.8-4 package
+% Mon Jan 27 16:35:17 2020
+\begin{table}[ht]
+\centering
+\begin{tabular}{rllrr}
+  \hline
+ & category & rule.type & related forms & exceptions \\ 
+  \hline
+1 & burnt & irregular &  24 &  41 \\ 
+  2 & kept & irregular &  10 &   9 \\ 
+  3 & sang & irregular &  11 &  77 \\ 
+  4 & burnt & regular &  38 &  21 \\ 
+  5 & drove & regular &  29 &  27 \\ 
+  6 & kept & regular &  67 &  41 \\ 
+  7 & sang & regular &  47 &  38 \\ 
+   \hline
+\end{tabular}
+\caption{The number of unique rules posited by the MGL in each verb class} 
+\label{mglruleses}
+\end{table}
 Table \ref{mglruleseskept} shows the list of (irregular and regular) rules for one of the classes, the KEPT class.

### 2020-01-28 functions/grabdata.R

@@ -1,11 +1,8 @@
-try(setwd('~/Github/RaczBecknerHayPierrehumbert2019'))
-# try(setwd('~/Work/NZILBB/RaczBecknerHayPierrehumbert2019'))
-
 ######################################################################
 # training set
 ######################################################################
 
-celex = read_csv('data/celex_verbs_with_categories_csv_fixed_moder.txt')
+celex = read_csv('~/Github/RaczBecknerHayPierrehumbert2019/data/celex_verbs_with_categories_csv_fixed_moder.txt')
 
 celex = celex %>% 
   select(
@@ -28,7 +25,7 @@ celex = celex %>%
 
 # fixing transcription
 
-transcription.matcher = read_csv('data/transcription_matcher.txt')
+transcription.matcher = read_csv('~/Github/RaczBecknerHayPierrehumbert2019/data/transcription_matcher.txt')
 
 celex = left_join(celex, transcription.matcher)
 
@@ -38,12 +35,11 @@ celex = left_join(celex, transcription.matcher)
 # test set
 ######################################################################
 
-baseline = read_csv('data/convergence_paper_baseline_data.txt')
+baseline = read_csv('~/Github/RaczBecknerHayPierrehumbert2019/data/convergence_paper_baseline_data.txt')
 
-baseline.verbs = read_csv('data/baseline_verbs.txt')
 # esp data
-espdata = read_csv('data/convergence_paper_esp_data.txt')
+espdata = read_csv('~/Github/RaczBecknerHayPierrehumbert2019/data/convergence_paper_esp_data.txt')
 
 # esp only post-test with gcm weights
-test2 = read_csv('data/convergence_paper_esp_test2_predictions.txt')
+test2 = read_csv('~/Github/RaczBecknerHayPierrehumbert2019/data/convergence_paper_esp_test2_predictions.txt')
 

### 2020-01-28 functions/grabfunctions.R

@@ -1,3 +1,7 @@
+###########################################################################
+# GCM functions
+###########################################################################
+
 # get similarity from whatever distance you choose to use
 getSimilarity = function(feature.distance, var_s, var_p){#the def values in the paper
   pairwise.similarity = exp ( -feature.distance / var_s )^var_p 
@@ -195,4 +199,174 @@ runGuyLookup = function(espdata, lookup, gcm.id, my.participant_id, var_s, var_p
   participant.output$participant_id = my.participant_id
   
   return(participant.output)
+}
+
+###########################################################################
+# MGL functions
+###########################################################################
+
+## functions
+
+# takes in read-in tsv, returns list of verbs with conf of regular form
+getMGLRules <- function(output){
+  
+  output2 = output %>% 
+    mutate(
+      past.form = sub("»", "", form2),
+      present.form = sub("»", "", form1),
+      alternation = paste0(A,X8,B),
+      environment = paste0(P,X15,Q)
+    ) %>% 
+    rename(
+      related_forms = `related forms`
+    )
+  
+  output2 = output2 %>% 
+    select(
+      present.form,
+      past.form,
+      alternation,
+      Pres,
+      Pfeat,
+      environment,
+      Qfeat,
+      Qres,
+      scope,
+      hits,
+      reliability,
+      confidence,
+      related_forms,
+      exceptions
+    )
+  
+  foc.dat = left_join(output2, verb_handles, by = 'present.form')
+  
+  
+  foc.irreg.rules = subset(foc.dat, as.character(past.form) == as.character(irregular.form))
+  foc.reg.rules = subset(foc.dat, as.character(past.form) == as.character(regular.form))
+  foc.irreg.rules = subset(foc.irreg.rules, !duplicated(present.form))
+  foc.reg.rules = subset(foc.reg.rules, !duplicated(present.form))
+  
+  foc.irreg.rules$rule_type = 'irregular'
+  foc.reg.rules$rule_type = 'regular'
+  
+  regular.rules = foc.reg.rules %>% 
+    select(
+      present.form,
+      reliability,
+      confidence,
+      category,
+      base.print
+    ) %>% 
+    rename(
+      reg.rel = reliability, 
+      reg.conf = confidence
+    )
+  
+  irregular.rules = foc.irreg.rules %>% 
[…]

### 2020-01-28 models/mgl/esp_mgl_by_hand_autogui.py

@@ -1,0 +1,106 @@
+################################################################
+# MGL batch runner script
+# pracz
+################################################################
+
+# The MGL can run batch files. However, if it does, it does not use feature files, even if these are available under the same name as the input file.
+
+# To run the MGL using features properly, you need to run these by hand. You also need to quit and reopen the MGL every time, otherwise it fits without using features the second and subsequent times.
+
+# The only way to do it is by hand.
+
+# The script below moves the cursor and types.
+# It launches the MGL, sets parameters, fits one input file, quits the MGL.
+# It does this i times.
+
+# Good to know:
+#   - you can break an autogui script by moving the mouse in one corner of the screen
+#   - pyautogui.click doesn't really work on Mac Os mojave, the below script uses a workaround
+#   - even so, you need to add the interface that launches the script (e.g. pycharm, iterm2, terminal) as well as the appropriate python launcher to apps that can control the computer in Security & Privacy / Accessibility
+#   - it's not a good idea to allow anything to control the computer like that in the long run
+
+import pyautogui
+import time
+
+# takes two coordinates on the computer screen and a duration, returns a left mouse click on the coordinates after duration.
+def clickThing(x, y, duration = 2):
+
+    pyautogui.moveTo(x, y, duration = duration)
+    pyautogui.dragTo(x, y, button='left')
+    #pyautogui.click(x, y, button='left')
+
+# runs the MGL
+# assumes:
+#   - iMac (Retina 4K, 21.5-inch, 2019)
+#   - resolution: width=1680, height=945 (scaled one to Larger)
+#   - MinGenLearner.jar icon on desktop, bottom left corner, visible
+#   - MinGenLearner opens upper right corner
+
+def runGuy(input_file_name):
+
+    # click on MinGenLearner.jar and then double click
+    clickThing(1632, 858)
+    pyautogui.doubleClick(1632, 858)
+
+    # move SLOWLY to tick Show All Options box in MGL gui (so it has time to boot up)
+    clickThing(44, 120, duration = 10)
+
+    # click into Lower confidence limit, move it to .55, enter
+    clickThing(450, 244)
+    for i in range(1, 5):
+        pyautogui.press('up')
+
+    pyautogui.press('enter')
+
+    # click into Upper confidence limit, move it to .95, enter
+    clickThing(450, 268)
+    for i in range(1, 5):
[…]

### 2020-01-28 models/mgl/mgl_esp_create_input.R

@@ -1,203 +1,0 @@
-library(Hmisc)
-# library(dplyr)
-# library(stringr)
-# library(knitr)
-library(lme4)
-library(RePsychLing)
-library(tidyverse)
-setwd('~/Github/RaczBecknerHayPierrehumbert2019/')
-
-# MGL batch run doesn't pull in features. this has to be done by hand. 
-# I don't want to run 222 participants by hand.
-# Experiment regularisation rate, lexical typicality and list block order create the possible training + test data combinations for the individual MGL-s. 
-# Individual participants will move round slightly more but this is a reasonable approximation (and still takes hours)
-# I can pick the unique combinations of reg rate + lex typ + list_block_order and run one participant each.
-# this doesn't cover all participants, since 
-
-esp <- read_csv('data/convergence_paper_esp_data.txt') %>% filter(phase == 'test2')
-
-participants <- esp %>% select(participant_id,reg_rate,lex_typicality,list_block_order) %>% unique %>% arrange(reg_rate,lex_typicality,list_block_order) %>% mutate(part_type = paste(reg_rate, lex_typicality, list_block_order, sep = ' ')) %>% select(participant_id,part_type)
-participants <- participants[!duplicated(participants$part_type),]
-handles <- read.csv("data/participanthandles.txt")
-participants <- merge(participants,handles)
-
-# now I paste the relevant participants' in files into inputlists_wf_minimal
-
-# now I run the Minimal Generalisation Learner on each, 
-
-# you have to quit the program EVERY TIME after you fit an individual file, otherwise it doesn't work.
-
-# this works: turn mingenlearner off and on again every time, set lower and upper to 55 95 every time.
-
-# I check every sum file to see if it specifies a set of segments for P environment. that is indicative of features being used. (length of fit is also indicative: with features it takes a noticable amount of time, w/o it fits almost instantly)
-# close mgl. check sum output for features. open mgl. set 55 and 95. select next in. run. takes about 50 sec to fit.
-
-# This takes a couple hours.
-
-# I gather predictions.
-
-# participants2 <- participants %>% filter(participant < 100)
[…]

### 2020-01-28 models/mgl/mgl_process_output.R

@@ -1,0 +1,83 @@
+library(tidyverse)
+library(purrr)
+setwd('~/Github/RaczBecknerHayPierrehumbert2019/')
+
+## functions
+
+source('~/Github/RaczBecknerHayPierrehumbert2019/functions/grabfunctions.R')
+
+## data
+
+test2 = read_csv('~/Github/RaczBecknerHayPierrehumbert2019/data/convergence_paper_esp_test2_predictions.txt')
+
+espdata = read_csv('~/Github/RaczBecknerHayPierrehumbert2019/data/convergence_paper_esp_data.txt')
+baseline = read_csv('~/Github/RaczBecknerHayPierrehumbert2019/data/convergence_paper_baseline_data.txt')
+
+participant_handles = read_csv('~/Github/RaczBecknerHayPierrehumbert2019/data/participanthandles.txt')
+
+verb_handles = read_csv("~/Github/RaczBecknerHayPierrehumbert2019/data/256verbs_print_forms.txt") %>% 
+  select(
+    base.form,
+    category,
+    base.print,
+    regular.form,
+    irregular.form
+    ) %>% 
+  rename(present.form = base.form) %>% 
+  mutate(
+    irregular.form = str_replace_all(irregular.form, "5","o"),
+    regular.form = str_replace_all(regular.form, "Id","@d")
+  )
+
+############# run everything #############################
+
+mgl_baseline_rules_raw = read_tsv('~/Github/RaczBecknerHayPierrehumbert2019/models/mgl/baseline_mgl/CELEXFull3.sum')
+mgl_baseline_rules = getMGLRules(mgl_baseline_rules_raw)
+
+write_tsv(mgl_baseline_rules, '~/Github/RaczBecknerHayPierrehumbert2019/models/mgl/baseline_mgl/baseline_rules_features.tsv')
+
+participant_handles_edits = getNoFeatures(participant_handles)
+participant_handles_features = getWithFeatures(participant_handles)
+
+participant_handles_edits = participant_handles_edits %>% 
+  rename(disc = present.form)
+
+participant_handles_features = participant_handles_features %>% 
+  rename(disc = present.form)
+
+test2 = test2 %>% 
+  select(
+    -individual_mgl_edits,
+    -individual_mgl_features
+  ) %>% 
+  left_join(
+    participant_handles_features, by = c("disc", "participant_id")
+  ) %>% 
+  left_join(
[…]

### 2020-01-28 models/mgl/mgl_processor.R

@@ -1,69 +1,0 @@
-# code to process an MGL sum file
-setwd('~/Github/RaczBecknerHayPierrehumbert2019/models/gcm')
-
-library(tidyverse)
-
-handles <- read.delim("data/256verbs_print_forms.txt", sep=",") %>% select(base.form,category,base.print,regular.form,irregular.form) %>% rename(present.form = base.form)
-# get formatting right
-handles$irregular.form <- sub("5","o",handles$irregular.form)
-handles$regular.form <- sub("Id","@d",handles$regular.form)
-
-
-getMGLRules <- function(output){
-output2 <- output
-
-output2$past.form <- sub("»", "", output2$V5)
-output2$present.form <- sub("»", "", output2$V3)
-
-output2$alternation <- paste(as.character(output2$V7),as.character(output2$V8),as.character(output2$V9), sep="")
-output2$Pres <- output2$V12
-output2$Pfeat <- output2$V13
-output2$environment <- paste(as.character(output2$V14),as.character(output2$V15),as.character(output$V16), sep = ' ')
-output2$Qfeat <- output2$V17
-output2$Qres <- output2$V18
-output2$scope <- output2$V19
-output2$hits <- output2$V20
-# output2$rule2 <- paste(as.character(output2$rule),as.character(output2$structural_description), sep= ' / ')
-output2$reliability <- as.numeric(as.character(output2$V21))
-output2$confidence <- as.numeric(as.character(output2$V22))
-output2$related_forms <- output2$V23
-output2$exceptions <- output2$V24
-
-# output2$past.form <- output2$V5
-# output2$present.form <- output2$V3
-# It does the thing where it inserts an onset-rhyme boundary marker
-
-output3 <- output2[,c("present.form","past.form","alternation","Pres","Pfeat","environment","Qfeat","Qres","scope","hits","reliability","confidence","related_forms","exceptions")]
-output3 <- output3[-1,]
-
-output3 <- output3[order(-output3$confidence),]
-
-foc.dat <- merge(output3,handles)
-
-foc.dat <- foc.dat[order(-foc.dat$confidence),]
-foc.dat <- foc.dat[order(foc.dat$present.form),]
-
-foc.irreg.rules <- subset(foc.dat, as.character(past.form) == as.character(irregular.form))
-foc.reg.rules <- subset(foc.dat, as.character(past.form) == as.character(regular.form))
[…]

### 2020-01-29 supplementary_information/si.Rnw
haladunk

@@
 library(xtable)
+library(lme4)
 
@@
   mutate(
-    c.baseline_gcm_features = scales::rescale(baseline_gcm_features)
+    r_baseline_gcm_features = scales::rescale(baseline_gcm_features)
   ) %>% 
   filter(word == 'splive') %>% 
-  pull(c.baseline_gcm_features) %>% 
+  pull(r_baseline_gcm_features) %>% 
   unique %>% 
@@
   mutate(
-    c.baseline_mgl_features = scales::rescale(baseline_mgl_features)
+    r_baseline_mgl_features = scales::rescale(baseline_mgl_features)
   ) %>% 
   filter(word == 'splive') %>% 
-  select(baseline_mgl_features,c.baseline_mgl_features) %>% 
+  select(baseline_mgl_features,r_baseline_mgl_features) %>% 
   distinct()
@@
 
-ITT TARTOK
+\section{Regression modeling procedures}\label{appendixregression}
 
-\section{Residualization of the two categorization models on the baseline data}\label{appendixgcmmglresidualization}
+<<main_model1, echo=F,eval=T>>=
+test2 = test2 %>% 
+  mutate(
+    verb_baseline_average = baseline_av - mean(baseline_av, na.rm = T),
+    participant_prereg_average = pre_reg_av - mean(pre_reg_av, na.rm = T),
+    bot_lexical_typicality = case_when(
+        lex_typicality == 'typical' ~ 'typical',
+        lex_typicality == 'random' ~ 'random',
+        lex_typicality == 'atypical' ~ 'reversed'
+      ) %>% 
+      factor(levels = c('typical', 'random', 'reversed')),
+    bot_regularisation_rate = reg_rate %>% 
+      factor(levels = c('nc', '-40%', '+40%'))
+  )
 
-Since the regularity scores of the GCM and the MGL for the nonce verbs are correlated  (r $=$ 0.48), we use residualization to see whether the two models make separate significant contributions to explaining variation in the baseline dataset (Section 6.3 in the main text.)
+##########################################
+# from Florian Jaeger
+
+vif.mer <- function (fit) {
+  ## adapted from rms::vif
+  v <- vcov(fit)
+  nam <- names(fixef(fit))
+  ## exclude intercepts
+  ns <- sum(1 * (nam == "Intercept" | nam == "(Intercept)"))
+  if (ns > 0) {
+    v <- v[-(1:ns), -(1:ns), drop = FALSE]
+    nam <- nam[-(1:ns)] }
+  d <- diag(v)^0.5
+  v <- diag(solve(v/(d %o% d)))
+  names(v) <- nam 
+  v 
+  }
+@
+
+<<main_model2, echo=F,eval=F>>=
+
+lme.fit1 = glmer(resp_post_reg ~ 1 + verb_baseline_average * bot_lexical_typicality + participant_prereg_average + bot_regularisation_rate +
+    (1 + verb_baseline_average | participant_id) + 
+    (1  | word), 
+  data = test2, 
+  family = binomial (link = logit),
+  control=glmerControl(optimizer="bobyqa")
+)
+
+save(lme.fit1, file = 'si_fit1.rda')
+vif.mer(lme.fit1)
+@
+
+<<main_model3, echo=F,eval=T, results=tex>>=
+load('si_fit1.rda')
+print(xtable(summary(lme.fit1)$coef, digits = 3, label = 'regressiontable1', caption = 'Experimental Data, Post-test Regression Model Summary.
+Model formula: post-test regular response ~ verb baseline mean x lexical typicality + participant pretest mean + regularization shift + (1 + verb.baseline.mean | participant) + (1 | verb)'))
+@
+
+Our regression model selection in this article proceeded as follows.
+Except where stated otherwise, we fit logistic mixed-effects regression models, using R's lme4 package \citep{bateslme4}.
+Non-significant predictors are removed from our models, as determined by a series of comparisons of nested models, using likelihood ratio tests in R's anova function.
+However, our model tables (e.g., Tables 5 and 6 in the main text) include the significance levels generated by lme4 summaries.
+
+We investigate maximal random effects structures, allowing for random intercepts for participants and stimulus items (verbs), as well as random slopes for all within-unit factors.
+We compare nested random effects structures as above, and retain any random slopes and intercepts which are supported by likelihood ratio tests \citep{baayen2008mixed}.
+
+In the model in Table 5 in the main text, repeated here as Table \ref{regressiontable1}, we elect to mean center continuous variables to counteract collinearity that becomes evident when including the interaction between and {\em lexical typicality} .
+Although some debate exists regarding the practice of transforming variables \citep{belsley1991conditioning,echambadi2007mean}, centering variables is common practice to remove nonessential ill conditioning, that is, relationships that inevitably exist between main and product terms \citep{aiken1991multiple,jaccard2003interaction, jaeger2008categorical}.
+We note that if {\em verb baseline mean} is {\em not} centered, collinearity is high (VIF=15.79) in this model.
+A non-centered model also yields a significant main effect for {\em lexical typicality}, with increased post-test regularization in the random and reversed conditions.
+However, we elect to take a more conservative approach by not including a potentially spurious main effect.
+Note that other than this main effect difference, models with and without centering yield results that are qualitatively indistinguishable; all other main effects and interactions are present in either case.
+The collinearity for the centered model (Table 5 in the main text) is acceptable, with a maximum VIF score of 2.9.
+
+<<Bayes factor, echo=F, eval=F>>=
+# random vs atypical
+
+test2b = test2 %>% 
+  filter(
+    bot_lexical_typicality != 'typical'
+  )
+
+fit_power1 = glmer(resp_post_reg ~ 1 + verb_baseline_average * bot_lexical_typicality + participant_prereg_average + bot_regularisation_rate +
+    (1 + verb_baseline_average | participant_id) + 
+    (1  | word), 
+  data = test2b, 
+  family = binomial ( link = logit ),
+  control=glmerControl(optimizer="bobyqa")
+  )
+
+fit_power2 = glmer(resp_post_reg ~ 1 + verb_baseline_average + bot_lexical_typicality + participant_prereg_average + bot_regularisation_rate +
+    (1 + verb_baseline_average | participant_id) + 
+    (1  | word), 
+  data = test2b, 
+  family = binomial ( link = logit ),
+  control=glmerControl(optimizer="bobyqa")
+  )
+
+# summary(fit_power1)
+# summary(fit_power2)
+
+exp((BIC(fit_power2) - BIC(fit_power1))/2)
+@
+
+Since it is unclear whether the absence of a difference between random and reversed conditions is due to a lack of data, we refit the model with the baseline average : typicality interaction on the random and reversed conditions only.
+Then we calculated the Bayes Factor of the null model (with no interaction) over the full model (with interaction) using Bayesian Information Criteria calculated by lme4.
+The result strongly supports the null model over the full model, indicating that the preference for the null model is {\em not} due to the lack of evidence to reject it.
+
+\section{Fits of the two categorization models on the baseline data}\label{appendixgcmmglresidualization}
+
+We fit the Celex-GCM and the Celex-MGL on the baseline data to see how well they predict regularisation responses by the baseline participants.
+The Celex-MGL outperforms the Celex-GCM, if slightly.
+Both models contribute independently to explaining variation in the data.
+
+<<celex_gcm_mgl1, eval=T, echo=F>>=
+
+baseline = baseline %>% 
+  mutate(
+    r_baseline_mgl_features = scales::rescale(baseline_mgl_features),
+    r_baseline_gcm_features = scales::rescale(baseline_gcm_features),
+  )
+
+C_baseline_mgl_features = with(baseline, Hmisc::somers2(r_baseline_mgl_features,regular)['C']) %>% 
+  round(3)
+C_baseline_gcm_features = with(baseline, Hmisc::somers2(r_baseline_gcm_features,regular)['C']) %>% 
+  round(3)
+
+mgl_gcm_cor = baseline %>% 
+  distinct(word,r_baseline_mgl_features,baseline_gcm_features) %>% 
+  summarise(
+    cor = cor(r_baseline_mgl_features,baseline_gcm_features)
+  ) %>% 
+  pull() %>% 
+  round(3)
+
+@
+
+The concordance indices of the Celex-MGL (\Sexpr{C_baseline_mgl_features}) and the Celex-MGL (\Sexpr{C_baseline_gcm_features}) on the baseline data are similar, with some edge retained by the MGL, as in the case of \citep{albright2003rules}.
+Our concordance indices are calculated using the somers2 function in R's \texttt{Hmisc} package \citep{harrell2014hmisc}.
+
+Since the regularity scores of the GCM and the MGL for the nonce verbs are correlated  (r $=$ \Sexpr{mgl_gcm_cor}), we use residualization to see whether the two models make separate significant contributions to explaining variation in the baseline dataset (Section 6.3 in the main text.)
 
@@
 We rescale and center GCM and MGL scores for the model fits.
+
+<<celex_gcm_mgl2, eval=F, echo=F>>=
+
+baseline_words = baseline %>% 
+  distinct(word, r_baseline_mgl_features, r_baseline_gcm_features)
+
+celex_m1 = lm(r_baseline_gcm_features ~ 1 + r_baseline_mgl_features, data = baseline_words)
+celex_m2 = lm(r_baseline_mgl_features ~ 1 + r_baseline_gcm_features, data = baseline_words)
+
+baseline_words = baseline_words %>% 
+  mutate(
+    r_baseline_gcm_features_residual = resid(celex_m1),
+    r_baseline_mgl_features_residual = resid(celex_m2)
+  )
+
+baseline = left_join(baseline, baseline_words, by = c("word", "r_baseline_mgl_features", "r_baseline_gcm_features"))
+
+celex_m3 = glmer(regular ~ 1 + r_baseline_mgl_features + r_baseline_gcm_features_residual + 
+                (1 + r_baseline_mgl_features + r_baseline_gcm_features_residual | subject), 
+                data = baseline, 
+                family = binomial ( link = logit ),
+                control=glmerControl(optimizer="bobyqa"))
+
+celex_m4 = glmer(regular ~ 1 + r_baseline_gcm_features + r_baseline_mgl_features_residual + 
+                (1 + r_baseline_gcm_features + r_baseline_mgl_features_residual | subject), 
+                data = baseline, 
+                family = binomial ( link = logit ),
+                control=glmerControl(optimizer="bobyqa"))
+
+save(celex_m3, file = 'si_celex_m3.rda')
+save(celex_m4, file = 'si_celex_m4.rda')
+
+@
+
+<<celex_resid1, echo=F,eval=T, results=tex>>=
+load('si_celex_m3.rda')
+print(xtable(summary(celex_m3)$coef, digits = 3, label = 'resid1', caption = 'MGL scores and GCM residuals, baseline data'))
+@
+
+<<celex_resid2, echo=F,eval=T, results=tex>>=
+load('si_celex_m4.rda')
+print(xtable(summary(celex_m4)$coef, digits = 3, label = 'resid2', caption = 'GCM scores and MGL residuals, baseline data'))
+@
 
@@
 
- The residuals remain significant predictors in both models (M3: MGL score ($\beta$ $=$ 1.27, p $<$ 0.0001), GCM residual ($\beta$ $=$ 1.37, p $<$ 0.0001); M4: GCM score ($\beta$ $=$ 2.29, p $<$ 0.0001), MGL residual ($\beta$ $=$ 0.95, p $<$ 0.0001).
+The model summaries can be seen in Table \ref{resid1} (M3) and Table \ref{resid2} (M4).
+The residuals remain significant predictors in both models.
 Note that while this paper reports values fit to the baseline experiment data, the same pattern of results also holds for models fit to the pre-test data.
@@
 
-\section{Additional information regarding  the role of the GCM and the MGL in the post-test \label{posttestimplementation}}
+\section{Model fits on the post-test data}
 
-\subsection{Concordance Indices}
+We fit the Celex-GCM,the Celex-MGL, as well as the Individual-GCM and the Individual-MGL on the post-test data.
+Recall that the Celex-based models were trained on verbs in Celex.
+The individual-based models were trained for each participant, on Celex verbs as well as the co-player's responses in the ESP test.
+They were tested on the post-test for each participant, individually.
+
+In the paper, we list four techniques of evaluating model fit. We go through these in turn.
+
+\subsection{Comparing concordance indices}
+
+First, we calculated concordance indices for the Celex- and individual-models on the post-test data of the ESP experiment.
+
+<<first, echo=F,eval=T>>=
+
+test2 = test2 %>% 
+ mutate(
+    r_baseline_mgl_features = scales::rescale(baseline_mgl_features),
+    r_baseline_gcm_features = scales::rescale(baseline_gcm_features),
+    r_individual_mgl_features = scales::rescale(individual_mgl_features),
+    r_individual_gcm_features = scales::rescale(individual_gcm_features)
+ )
+
+# C_baseline_mgl_features # celex-mgl on baseline
+# C_baseline_gcm_features # celex-gcm on baseline
+
+C_test2_baseline_mgl_features = with(test2, Hmisc::somers2(r_baseline_mgl_features,resp_post_reg)['C']) %>% round(3)
+C_test2_baseline_gcm_features = with(test2, Hmisc::somers2(r_baseline_gcm_features,resp_post_reg)['C']) %>% round(3)
+C_test2_individual_mgl_features = with(test2, Hmisc::somers2(r_individual_mgl_features,resp_post_reg)['C']) %>% round(3)
+C_test2_individual_gcm_features = with(test2, Hmisc::somers2(r_individual_gcm_features,resp_post_reg)['C']) %>% round(3)
+
+@
+
 Table \ref{modellingsummary} summarizes the concordance indices between our data and the various categorization models.
-The individual-GCM improves on the \textsc{Celex}-GCM in terms of explaining the post-test data.
-The individual-MGL does not improve on the \textsc{Celex}-MGL.
+The \textsc{Celex}-MGL, outperforms the \textsc{Celex}-GCM on the post-test data to some extent, much like in the case of our baseline data.
+The individual-GCM considerably improves on the \textsc{Celex}-GCM in terms of explaining the post-test data.
+The individual-MGL also improves on the \textsc{Celex}-MGL, but to a lesser extent.
 
@@
     \begin{tabular}{lllll}
-            & training data & method & test data & C (test data)\\
+            & training data & method & test data & C\\
             \hline
-\textsc{Celex}-GCM & real verbs & `analogy' & verbs in baseline &{\bf 0.58 }\\
-\textsc{Celex}-MGL & real verbs & `rules' & verbs in baseline &{\bf  0.59 }\\ \hline
-\textsc{Celex}-GCM & real verbs & `analogy' & verbs in post-test &{\bf  0.6}\\
-individual-GCM & real verbs + verbs seen in ESP & `analogy' & verbs in post-test &{\bf  0.68}\\
+\textsc{Celex}-GCM & real verbs & `analogy' & verbs in baseline &\Sexpr{C_baseline_gcm_features }\\
+\textsc{Celex}-MGL & real verbs & `rules' & verbs in baseline &\Sexpr{C_baseline_mgl_features }\\ \hline
+\textsc{Celex}-GCM & real verbs & `analogy' & verbs in post-test &\Sexpr{C_test2_baseline_gcm_features}\\
+individual-GCM & real verbs + verbs seen in ESP & `analogy' & verbs in post-test &\Sexpr{C_test2_individual_gcm_features}\\
 
-\textsc{Celex}-MGL & real verbs & `rules' & verbs in post-test &{\bf  0.61}\\
-individual-MGL & real verbs + verbs seen in ESP & `rules' & verbs in post-test &{\bf  0.55}\\
+\textsc{Celex}-MGL & real verbs & `rules' & verbs in post-test &\Sexpr{C_test2_baseline_mgl_features}\\
+individual-MGL & real verbs + verbs seen in ESP & `rules' & verbs in post-test &\Sexpr{C_test2_individual_mgl_features}\\
 \hline
@@
 
- Our concordance indices are calculated using the somers2 function in R's \texttt{Hmisc} package \citep{harrell2014hmisc}.
+The issue here is that a concordance index is not a statistical test, and does not provide a margin of error.
+The trends are apparent: the individual-GCM is more of an improvement than the individual-MGL.
+But the strength of these trends is less clear.
 
- Note that the indices in Table \ref{modellingsummary} directly calculate C between the relevant algorithm score (MGL or GCM) and the binary post-test outcome (regular or irregular).
-As a diagnostic for our regression models, we also calculated concordance indices between model predictions and participants' regular/irregular responses.
-The concordance indices for multiple regression models are, unsurprisingly, considerably higher than the indices calculated on the basis of a single variable.
+\subsection{Extra information provided by the individual model over the \sc{Celex}-model}
 
- The models in Tables 5 and 6 of the main text both have an index of 0.84, indicating an acceptable level of predictive ability \citep{harrell2013regression}.
+Second, we considered the amount of extra variation explained by the individual-model over the \sc{Celex}-model, for the MGL and the GCM, respectively.
 
- Since the random effects likely account for a large amount of this variance, we calculated the marginal r${^2}$ values for both models: these are 0.27 and 0.22, respectively (the conditional values are 0.45 and 0.47).
-Conditions and learning models explain some amount of the variation, but the data
- are noisy.
+Section 6.4 in the main text explains that we calculate the extra information given by the individual models by subtracting the \sc{Celex} model prediction from the individual model prediction for each item for each participant.
 
-\subsection{Assessing separate contribution of individual models over \textsc{Celex} models}
-The text (Section 6.4 in the main text) explains that we calculate the extra information given by the individual models by subtracting the \textsc{Celex} model prediction from the individual model prediction for each item for each participant, and that we find that the extra information from the individual-GCM  models is predictive over and above the \textsc{Celex}-GCM model.
+<<modeldiff1, echo=F,eval=F>>=
+
+test2 = test2 %>% 
+  mutate(
+    gcm_score_diff = r_individual_gcm_features - r_baseline_gcm_features,
+    mgl_score_diff = r_individual_mgl_features - r_baseline_mgl_features
+)
+
+resp_post_reg ~ 1 + verb_baseline_average * bot_lexical_typicality + participant_prereg_average + bot_regularisation_rate
+
+indiv_m5 = glmer(resp_post_reg ~ r_baseline_gcm_features + gcm_score_diff + participant_prereg_average + 
+                   (1 + r_baseline_gcm_features + gcm_score_diff | participant_id) + (1 | word), 
+                 data = test2, 
+                 family = binomial (link = logit), 
+                 control=glmerControl(optimizer="bobyqa"))
+
+save(indiv_m5, file = 'si_indiv_m5.rda')
+
+indiv_m6 = glmer(resp_post_reg ~ r_baseline_mgl_features + mgl_score_diff + participant_prereg_average  + 
+                   (1 + r_baseline_mgl_features + mgl_score_diff | participant_id) + (1 | word), 
+                 data = test2, 
+                 family = binomial (link = logit), 
+                 control=glmerControl(optimizer="bobyqa"))
+
+save(indiv_m6, file = 'si_indiv_m6.rda')
+
+@
+
+<<indiv_diff1, echo=F, eval=T, results=tex>>=
+
+load('si_indiv_m5.rda')
+print(xtable(summary(indiv_m5)$coef, digits = 3, label = 'indiv1', caption = 'Celex-GCM and individual score difference, post-test data'))
+
+@
+
+<<indiv_diff2, echo=F,eval=T, results=tex>>=
+load('si_indiv_m6.rda')
+print(xtable(summary(indiv_m6)$coef, digits = 3, label = 'indiv2', caption = 'Celex-MGL and individual score difference, post-test data'))
+@
 
@@
 
-In Model 5, the extra information from the individual-GCM is a significant predictor of post-test responses  ($\beta$ $=$ 2.61, p $<$ 0.001),  above and beyond the \textsc{Celex}-GCM contribution.
-In similar Model 6, however, the extra information from the individual-MGL is not a significant predictor of post-test responses ($\beta$ $=$ 0.21, p $=$ 0.23).
-This means that the individual-GCM has some success in modeling post-test behavior when trained with real English verbs and information from the ESP test, whereas adjustments made to the individual-MGL do not improve predictive power.
+The extra information is a significant predictor of post-test responses in both models.
+However, this effect is much more robust for the GCM than for the MGL, as seen in Table \ref{indiv1} (M5) and Table \ref{indiv2} (M6).
+The individual-GCM has some success in modeling post-test behavior when trained with real English verbs and information from the ESP test, whereas adjustments made to the individual-MGL do not vastly improve predictive power.
 
-\section{Regression modeling procedures}\label{appendixregression}
+These techniques are indicative of the relative strengths of the individual-MGL and -GCM respective to their \sc{Celex}-counterparts and each other.
+However, they do not compare them directly using statistical tests.
 
-Our regression model selection in this article proceeded as follows.
-Except where stated otherwise, we fit logistic mixed-effects regression models, using R's lme4 package \citep{bateslme4}.
-Non-significant predictors are removed from our models, as determined by a series of comparisons of nested models, using likelihood ratio tests in R's anova function.
-However, our model tables (e.g., Tables 5 and 6 in the main text) include the significance levels generated by lme4 summaries.
-
-We investigate maximal random effects structures, allowing for random intercepts for participants and stimulus items (verbs), as well as random slopes for all within-unit factors.
-We compare nested random effects structures as above, and retain any random slopes and intercepts which are supported by likelihood ratio tests \citep{baayen2008mixed}.
-
-In the model in Table 5 in the main text, we elect to mean center continuous variables to counteract collinearity that becomes evident when including the interaction between and {\em lexical typicality} .
-Although some debate exists regarding the practice of transforming variables \citep{belsley1991conditioning,echambadi2007mean}, centering variables is common practice to remove nonessential ill conditioning, that is, relationships that inevitably exist between main and product terms \citep{aiken1991multiple,jaccard2003interaction, jaeger2008categorical}.
-We note that if {\em verb baseline mean} is {\em not} centered, collinearity is high (VIF=15.79) in this model.
-A non-centered model also yields a significant main effect for {\em lexical typicality}, with increased post-test regularization in the random and reversed conditions.
-However, we elect to take a more conservative approach by not including a potentially spurious main effect.
-Note that other than this main effect difference, models with and without centering yield results that are qualitatively indistinguishable; all other main effects and interactions are present in either case.
-The collinearity for the centered model (Table 5 in the main text) is acceptable, with a maximum VIF score of 2.89.
-
-Since it is unclear whether the absence of a difference between random and reversed conditions is due to a lack of data, we refit the model with the baseline average : typicality interaction on the random and reversed conditions only.
-Then we calculated the Bayes Factor of the null model (with no interaction) over the full model (with interaction) using Bayesian Information Criteria calculated by lme4.
-The result strongly supports the null model over the full model, indicating that the preference for the null model is {\em not} due to the lack of evidence to reject it.
+\subsection{Using combined model predictions to explain variation in participant post-test responses}
 

### 2020-01-29 supplementary_information/si.tex
haladunk

@@
 % latex table generated in R 3.6.0 by xtable 1.8-4 package
-% Mon Jan 27 16:35:17 2020
+% Wed Jan 29 18:07:06 2020
 \begin{table}[ht]
@@
 % latex table generated in R 3.6.0 by xtable 1.8-4 package
-% Mon Jan 27 16:35:17 2020
+% Wed Jan 29 18:07:07 2020
 \begin{table}[ht]
@@
 
-ITT TARTOK
+\section{Regression modeling procedures}\label{appendixregression}
 
-\section{Residualization of the two categorization models on the baseline data}\label{appendixgcmmglresidualization}
+% latex table generated in R 3.6.0 by xtable 1.8-4 package
+% Wed Jan 29 18:07:07 2020
+\begin{table}[ht]
+\centering
+\begin{tabular}{rrrrr}
+  \hline
+ & Estimate & Std. Error & z value & Pr($>$$|$z$|$) \\ 
+  \hline
+(Intercept) & -0.108 & 0.161 & -0.672 & 0.501 \\ 
+  verb\_baseline\_average & 8.259 & 0.686 & 12.040 & 0.000 \\ 
+  bot\_lexical\_typicalityrandom & 0.043 & 0.169 & 0.253 & 0.800 \\ 
+  bot\_lexical\_typicalityreversed & 0.035 & 0.173 & 0.204 & 0.839 \\ 
+  participant\_prereg\_average & 7.125 & 0.563 & 12.667 & 0.000 \\ 
+  bot\_regularisation\_rate-40\% & -0.456 & 0.185 & -2.468 & 0.014 \\ 
+  bot\_regularisation\_rate+40\% & 0.549 & 0.166 & 3.307 & 0.001 \\ 
+  verb\_baseline\_average:bot\_lexical\_typicalityrandom & -2.402 & 0.927 & -2.591 & 0.010 \\ 
+  verb\_baseline\_average:bot\_lexical\_typicalityreversed & -2.127 & 0.955 & -2.226 & 0.026 \\ 
+   \hline
+\end{tabular}
+\caption{Experimental Data, Post-test Regression Model Summary.
+Model formula: post-test regular response ~ verb baseline mean x lexical typicality + participant pretest mean + regularization shift + (1 + verb.baseline.mean | participant) + (1 | verb)}
+\label{regressiontable1}
+\end{table}
+Our regression model selection in this article proceeded as follows.
+Except where stated otherwise, we fit logistic mixed-effects regression models, using R's lme4 package \citep{bateslme4}.
+Non-significant predictors are removed from our models, as determined by a series of comparisons of nested models, using likelihood ratio tests in R's anova function.
+However, our model tables (e.g., Tables 5 and 6 in the main text) include the significance levels generated by lme4 summaries.
 
-Since the regularity scores of the GCM and the MGL for the nonce verbs are correlated  (r $=$ 0.48), we use residualization to see whether the two models make separate significant contributions to explaining variation in the baseline dataset (Section 6.3 in the main text.)
+We investigate maximal random effects structures, allowing for random intercepts for participants and stimulus items (verbs), as well as random slopes for all within-unit factors.
+We compare nested random effects structures as above, and retain any random slopes and intercepts which are supported by likelihood ratio tests \citep{baayen2008mixed}.
+
+In the model in Table 5 in the main text, repeated here as Table \ref{regressiontable1}, we elect to mean center continuous variables to counteract collinearity that becomes evident when including the interaction between and {\em lexical typicality} .
+Although some debate exists regarding the practice of transforming variables \citep{belsley1991conditioning,echambadi2007mean}, centering variables is common practice to remove nonessential ill conditioning, that is, relationships that inevitably exist between main and product terms \citep{aiken1991multiple,jaccard2003interaction, jaeger2008categorical}.
+We note that if {\em verb baseline mean} is {\em not} centered, collinearity is high (VIF=15.79) in this model.
+A non-centered model also yields a significant main effect for {\em lexical typicality}, with increased post-test regularization in the random and reversed conditions.
+However, we elect to take a more conservative approach by not including a potentially spurious main effect.
+Note that other than this main effect difference, models with and without centering yield results that are qualitatively indistinguishable; all other main effects and interactions are present in either case.
+The collinearity for the centered model (Table 5 in the main text) is acceptable, with a maximum VIF score of 2.9.
+
+Since it is unclear whether the absence of a difference between random and reversed conditions is due to a lack of data, we refit the model with the baseline average : typicality interaction on the random and reversed conditions only.
+Then we calculated the Bayes Factor of the null model (with no interaction) over the full model (with interaction) using Bayesian Information Criteria calculated by lme4.
+The result strongly supports the null model over the full model, indicating that the preference for the null model is {\em not} due to the lack of evidence to reject it.
+
+\section{Fits of the two categorization models on the baseline data}\label{appendixgcmmglresidualization}
+
+We fit the Celex-GCM and the Celex-MGL on the baseline data to see how well they predict regularisation responses by the baseline participants.
+The Celex-MGL outperforms the Celex-GCM, if slightly.
+Both models contribute independently to explaining variation in the data.
+
+The concordance indices of the Celex-MGL (0.586) and the Celex-MGL (0.581) on the baseline data are similar, with some edge retained by the MGL, as in the case of \citep{albright2003rules}.
+Our concordance indices are calculated using the somers2 function in R's \texttt{Hmisc} package \citep{harrell2014hmisc}.
+
+Since the regularity scores of the GCM and the MGL for the nonce verbs are correlated  (r $=$ 0.478), we use residualization to see whether the two models make separate significant contributions to explaining variation in the baseline dataset (Section 6.3 in the main text.)
 
@@
 
+% latex table generated in R 3.6.0 by xtable 1.8-4 package
+% Wed Jan 29 18:07:07 2020
+\begin{table}[ht]
+\centering
+\begin{tabular}{rrrrr}
+  \hline
+ & Estimate & Std. Error & z value & Pr($>$$|$z$|$) \\ 
+  \hline
+(Intercept) & -0.525 & 0.124 & -4.219 & 0.000 \\ 
+  r\_baseline\_mgl\_features & 1.347 & 0.093 & 14.482 & 0.000 \\ 
+  r\_baseline\_gcm\_features\_residual & 1.423 & 0.128 & 11.100 & 0.000 \\ 
+   \hline
+\end{tabular}
+\caption{MGL scores and GCM residuals, baseline data} 
+\label{resid1}
+\end{table}
+% latex table generated in R 3.6.0 by xtable 1.8-4 package
+% Wed Jan 29 18:07:07 2020
+\begin{table}[ht]
+\centering
+\begin{tabular}{rrrrr}
+  \hline
+ & Estimate & Std. Error & z value & Pr($>$$|$z$|$) \\ 
+  \hline
+(Intercept) & -1.225 & 0.160 & -7.682 & 0.000 \\ 
+  r\_baseline\_gcm\_features & 2.391 & 0.149 & 16.063 & 0.000 \\ 
+  r\_baseline\_mgl\_features\_residual & 1.008 & 0.090 & 11.154 & 0.000 \\ 
+   \hline
+\end{tabular}
+\caption{GCM scores and MGL residuals, baseline data} 
+\label{resid2}
+\end{table}
 Model 1 predicts the GCM scores from the MGL scores (M1: GCM regular score  $\sim$ MGL regular score).
@@
 
-\section{Additional information regarding  the role of the GCM and the MGL in the post-test \label{posttestimplementation}}
+\section{Model fits on the post-test data}
 
-\subsection{Concordance Indices}
+We fit the Celex-GCM,the Celex-MGL, as well as the Individual-GCM and the Individual-MGL on the post-test data.
+Recall that the Celex-based models were trained on verbs in Celex.
+The individual-based models were trained for each participant, on Celex verbs as well as the co-player's responses in the ESP test.
+They were tested on the post-test for each participant, individually.
+
+In the paper, we list four techniques of evaluating model fit. We go through these in turn.
+
+\subsection{Comparing concordance indices}
+
+First, we calculated concordance indices for the Celex- and individual-models on the post-test data of the ESP experiment.
+
 Table \ref{modellingsummary} summarizes the concordance indices between our data and the various categorization models.
-The individual-GCM improves on the \textsc{Celex}-GCM in terms of explaining the post-test data.
-The individual-MGL does not improve on the \textsc{Celex}-MGL.
+The \textsc{Celex}-MGL, outperforms the \textsc{Celex}-GCM on the post-test data to some extent, much like in the case of our baseline data.
+The individual-GCM considerably improves on the \textsc{Celex}-GCM in terms of explaining the post-test data.
+The individual-MGL also improves on the \textsc{Celex}-MGL, but to a lesser extent.
 
@@
             \hline
-\textsc{Celex}-GCM & real verbs & `analogy' & verbs in baseline &{\bf 0.58 }\\
-\textsc{Celex}-MGL & real verbs & `rules' & verbs in baseline &{\bf  0.59 }\\ \hline
-\textsc{Celex}-GCM & real verbs & `analogy' & verbs in post-test &{\bf  0.6}\\
-individual-GCM & real verbs + verbs seen in ESP & `analogy' & verbs in post-test &{\bf  0.68}\\
+\textsc{Celex}-GCM & real verbs & `analogy' & verbs in baseline &0.581\\
+\textsc{Celex}-MGL & real verbs & `rules' & verbs in baseline &0.586\\ \hline
+\textsc{Celex}-GCM & real verbs & `analogy' & verbs in post-test &0.602\\
+individual-GCM & real verbs + verbs seen in ESP & `analogy' & verbs in post-test &0.681\\
 
-\textsc{Celex}-MGL & real verbs & `rules' & verbs in post-test &{\bf  0.61}\\
-individual-MGL & real verbs + verbs seen in ESP & `rules' & verbs in post-test &{\bf  0.55}\\
+\textsc{Celex}-MGL & real verbs & `rules' & verbs in post-test &0.61\\
+individual-MGL & real verbs + verbs seen in ESP & `rules' & verbs in post-test &0.625\\
 \hline
@@
 
- Our concordance indices are calculated using the somers2 function in R's \texttt{Hmisc} package \citep{harrell2014hmisc}.
+The issue here is that a concordance index is not a statistical test, and does not provide a margin of error.
+The trends are apparent: the individual-GCM is more of an improvement than the individual-MGL.
+But the strength of these trends is less clear.
 
- Note that the indices in Table \ref{modellingsummary} directly calculate C between the relevant algorithm score (MGL or GCM) and the binary post-test outcome (regular or irregular).
-As a diagnostic for our regression models, we also calculated concordance indices between model predictions and participants' regular/irregular responses.
-The concordance indices for multiple regression models are, unsurprisingly, considerably higher than the indices calculated on the basis of a single variable.
+\subsection{Extra information provided by the individual model over the \sc{Celex}-model}
 
- The models in Tables 5 and 6 of the main text both have an index of 0.84, indicating an acceptable level of predictive ability \citep{harrell2013regression}.
+Second, we considered the amount of extra variation explained by the individual-model over the \sc{Celex}-model, for the MGL and the GCM, respectively.
 
- Since the random effects likely account for a large amount of this variance, we calculated the marginal r${^2}$ values for both models: these are 0.27 and 0.22, respectively (the conditional values are 0.45 and 0.47).
-Conditions and learning models explain some amount of the variation, but the data
- are noisy.
+Section 6.4 in the main text explains that we calculate the extra information given by the individual models by subtracting the \textsc{Celex} model prediction from the individual model prediction for each item for each participant.
 
-\subsection{Assessing separate contribution of individual models over \textsc{Celex} models}
-The text (Section 6.4 in the main text) explains that we calculate the extra information given by the individual models by subtracting the \textsc{Celex} model prediction from the individual model prediction for each item for each participant, and that we find that the extra information from the individual-GCM  models is predictive over and above the \textsc{Celex}-GCM model.
-
+% latex table generated in R 3.6.0 by xtable 1.8-4 package
+% Wed Jan 29 18:07:07 2020
+\begin{table}[ht]
+\centering
+\begin{tabular}{rrrrr}
+  \hline
+ & Estimate & Std. Error & z value & Pr($>$$|$z$|$) \\ 
+  \hline
+(Intercept) & -2.951 & 0.295 & -10.007 & 0.000 \\ 
+  r\_baseline\_gcm\_features & 4.369 & 0.425 & 10.284 & 0.000 \\ 
+  gcm\_score\_diff & 2.612 & 0.470 & 5.563 & 0.000 \\ 
+  participant\_prereg\_average & 5.289 & 0.602 & 8.788 & 0.000 \\ 
+   \hline
+\end{tabular}
+\caption{Celex-GCM and individual score difference, post-test data} 
+\label{indiv1}
+\end{table}
+% latex table generated in R 3.6.0 by xtable 1.8-4 package
+% Wed Jan 29 18:07:07 2020
+\begin{table}[ht]
+\centering
+\begin{tabular}{rrrrr}
+  \hline
+ & Estimate & Std. Error & z value & Pr($>$$|$z$|$) \\ 
+  \hline
+(Intercept) & -1.318 & 0.152 & -8.663 & 0.000 \\ 
+  r\_baseline\_mgl\_features & 1.944 & 0.176 & 11.045 & 0.000 \\ 
+  mgl\_score\_diff & 0.304 & 0.139 & 2.181 & 0.029 \\ 
+  participant\_prereg\_average & 7.203 & 0.591 & 12.190 & 0.000 \\ 
+   \hline
+\end{tabular}
+\caption{Celex-MGL and individual score difference, post-test data} 
+\label{indiv2}
+\end{table}
 We do this for both the GCM  (Model 5: regular response  $\sim$ \textsc{Celex} GCM score + extra GCM information + participant pre-test mean + (1 + \textsc{Celex} GCM + extra GCM information$|$participant) + (1$|$verb) and the MGL (Model 6: regular response  $\sim$ \textsc{Celex} MGL score + extra MGL information + participant pre-test mean + (1 + \textsc{Celex} MGL + extra MGL information$|$participant)+ (1$|$verb)).
 
-In Model 5, the extra information from the individual-GCM is a significant predictor of post-test responses  ($\beta$ $=$ 2.61, p $<$ 0.001),  above and beyond the \textsc{Celex}-GCM contribution.
-In similar Model 6, however, the extra information from the individual-MGL is not a significant predictor of post-test responses ($\beta$ $=$ 0.21, p $=$ 0.23).
-This means that the individual-GCM has some success in modeling post-test behavior when trained with real English verbs and information from the ESP test, whereas adjustments made to the individual-MGL do not improve predictive power.
+The extra information is a significant predictor of post-test responses in both models.
+However, this effect is much more robust for the GCM than for the MGL, as seen in Table \ref{indiv1} (M5) and Table \ref{indiv2} (M6).
+The individual-GCM has some success in modeling post-test behavior when trained with real English verbs and information from the ESP test, whereas adjustments made to the individual-MGL do not vastly improve predictive power.
 
-\section{Regression modeling procedures}\label{appendixregression}
+These techniques are indicative of the relative strengths of the individual-MGL and -GCM respective to their \sc{Celex}-counterparts and each other.
+However, they do not compare them directly using statistical tests.
 
-Our regression model selection in this article proceeded as follows.
-Except where stated otherwise, we fit logistic mixed-effects regression models, using R's lme4 package \citep{bateslme4}.
-Non-significant predictors are removed from our models, as determined by a series of comparisons of nested models, using likelihood ratio tests in R's anova function.
-However, our model tables (e.g., Tables 5 and 6 in the main text) include the significance levels generated by lme4 summaries.
-
-We investigate maximal random effects structures, allowing for random intercepts for participants and stimulus items (verbs), as well as random slopes for all within-unit factors.
-We compare nested random effects structures as above, and retain any random slopes and intercepts which are supported by likelihood ratio tests \citep{baayen2008mixed}.
-
-In the model in Table 5 in the main text, we elect to mean center continuous variables to counteract collinearity that becomes evident when including the interaction between and {\em lexical typicality} .
-Although some debate exists regarding the practice of transforming variables \citep{belsley1991conditioning,echambadi2007mean}, centering variables is common practice to remove nonessential ill conditioning, that is, relationships that inevitably exist between main and product terms \citep{aiken1991multiple,jaccard2003interaction, jaeger2008categorical}.
-We note that if {\em verb baseline mean} is {\em not} centered, collinearity is high (VIF=15.79) in this model.
-A non-centered model also yields a significant main effect for {\em lexical typicality}, with increased post-test regularization in the random and reversed conditions.
-However, we elect to take a more conservative approach by not including a potentially spurious main effect.
-Note that other than this main effect difference, models with and without centering yield results that are qualitatively indistinguishable; all other main effects and interactions are present in either case.
-The collinearity for the centered model (Table 5 in the main text) is acceptable, with a maximum VIF score of 2.89.
-
-Since it is unclear whether the absence of a difference between random and reversed conditions is due to a lack of data, we refit the model with the baseline average : typicality interaction on the random and reversed conditions only.
-Then we calculated the Bayes Factor of the null model (with no interaction) over the full model (with interaction) using Bayesian Information Criteria calculated by lme4.
-The result strongly supports the null model over the full model, indicating that the preference for the null model is {\em not} due to the lack of evidence to reject it.
+\subsection{Using combined model predictions to explain variation in participant post-test responses}
 

### 2020-01-29 supplementary_information/si.Rnw
some progress

@@
 library(lme4)
+library(patchwork) # take this out
 
@@
 source('~/Github/RaczBecknerHayPierrehumbert2019/functions/grabdata.R')
+
+# setting up columns in baseline and post-test data
+baseline = baseline %>% 
+  mutate(
+    r_baseline_mgl_features = scales::rescale(baseline_mgl_features), # rescale regularity weights
+    r_baseline_gcm_features = scales::rescale(baseline_gcm_features),
+  )
+
+test2 = test2 %>% 
+  mutate(
+    verb_baseline_average = baseline_av - mean(baseline_av, na.rm = T), # center verb average
+    participant_prereg_average = pre_reg_av - mean(pre_reg_av, na.rm = T), # center participant average
+    bot_lexical_typicality = case_when( # set factor levels used in paper, order factor
+        lex_typicality == 'typical' ~ 'typical',
+        lex_typicality == 'random' ~ 'random',
+        lex_typicality == 'atypical' ~ 'reversed'
+      ) %>% 
+      factor(levels = c('typical', 'random', 'reversed')),
+    bot_regularisation_rate = reg_rate %>% # order factor
+      factor(levels = c('nc', '-40%', '+40%')),
+    r_baseline_mgl_features = scales::rescale(baseline_mgl_features), # scale regular weights
+    r_baseline_gcm_features = scales::rescale(baseline_gcm_features),
+    r_individual_mgl_features = scales::rescale(individual_mgl_features),
+    r_individual_gcm_features = scales::rescale(individual_gcm_features),
+    gcm_score_diff = r_individual_gcm_features - r_baseline_gcm_features, # calculate score diff
+    mgl_score_diff = r_individual_mgl_features - r_baseline_mgl_features
+  )
+
+# mgl illustrations
+
+## get the raw output of the baseline feature-distance mgl
+mgl_baseline_rules_raw = read_tsv('~/Github/RaczBecknerHayPierrehumbert2019/models/mgl/baseline_mgl/CELEXFull3.sum') %>% 
+  mutate(
+    form1 = str_replace_all(form1, '»', ''), # get rid of syllable boundary markers
+    form2 = str_replace_all(form2, '»', '')
+  )
+
+## get verb handles
+verb_handles = read_csv('~/Github/RaczBecknerHayPierrehumbert2019/data/256verbs_print_forms.txt')
 
@@
 splive_gcm = test2 %>% 
-  mutate(
-    r_baseline_gcm_features = scales::rescale(baseline_gcm_features)
-  ) %>% 
   filter(word == 'splive') %>% 
@@
 
-# get the raw output of the baseline feature-distance mgl
-mgl_baseline_rules_raw = read_tsv('~/Github/RaczBecknerHayPierrehumbert2019/models/mgl/baseline_mgl/CELEXFull3.sum') %>% 
-  mutate(
-    form1 = str_replace_all(form1, '»', ''), # get rid of syllable boundary markers
-    form2 = str_replace_all(form2, '»', '')
-  )
-
 disc_category = test2 %>% # get words and categories from test2. match colnames with above
@@
   )
-
-verb_handles = read_csv('~/Github/RaczBecknerHayPierrehumbert2019/data/256verbs_print_forms.txt')
 
@@
 <<main_model1, echo=F,eval=T>>=
-test2 = test2 %>% 
-  mutate(
-    verb_baseline_average = baseline_av - mean(baseline_av, na.rm = T),
-    participant_prereg_average = pre_reg_av - mean(pre_reg_av, na.rm = T),
-    bot_lexical_typicality = case_when(
-        lex_typicality == 'typical' ~ 'typical',
-        lex_typicality == 'random' ~ 'random',
-        lex_typicality == 'atypical' ~ 'reversed'
-      ) %>% 
-      factor(levels = c('typical', 'random', 'reversed')),
-    bot_regularisation_rate = reg_rate %>% 
-      factor(levels = c('nc', '-40%', '+40%'))
-  )
 
@@
 
-lme.fit1 = glmer(resp_post_reg ~ 1 + verb_baseline_average * bot_lexical_typicality + participant_prereg_average + bot_regularisation_rate +
+# the model reported in the paper
+
+lme.fit1 = glmer(resp_post_reg ~ 
+    1 + verb_baseline_average * bot_lexical_typicality + participant_prereg_average + bot_regularisation_rate +
     (1 + verb_baseline_average | participant_id) + 
@@
 
+# the first comparison that comes to mind (no interaction):
+
+lme.fit1b = glmer(resp_post_reg ~ 
+    1 + verb_baseline_average + bot_lexical_typicality + participant_prereg_average + bot_regularisation_rate +
+    (1 + verb_baseline_average | participant_id) + 
+    (1  | word), 
+  data = test2, 
+  family = binomial (link = logit),
+  control=glmerControl(optimizer="bobyqa")
+)
+
+# the second comparison that comes to mind (three-way interaction):
+
+lme.fit1c = glmer(resp_post_reg ~ 
+    1 + verb_baseline_average * bot_lexical_typicality * bot_regularisation_rate + participant_prereg_average + 
+    (1 + verb_baseline_average | participant_id) + 
+    (1  | word), 
+  data = test2, 
+  family = binomial (link = logit),
+  control=glmerControl(optimizer="bobyqa", optCtrl = list(maxfun = 100000))
+)
+
 save(lme.fit1, file = 'si_fit1.rda')
+save(lme.fit1b, file = 'si_fit1b.rda')
+save(lme.fit1c, file = 'si_fit1c.rda')
 vif.mer(lme.fit1)
+vif.mer(lme.fit1c)
+
+anova(lme.fit1,lme.fit1b,lme.fit1c)
+
 @
@@
 
-fit_power1 = glmer(resp_post_reg ~ 1 + verb_baseline_average * bot_lexical_typicality + participant_prereg_average + bot_regularisation_rate +
+fit_power1 = glmer(resp_post_reg ~ 
+    1 + verb_baseline_average * bot_lexical_typicality + participant_prereg_average + bot_regularisation_rate +
     (1 + verb_baseline_average | participant_id) + 
@@
 
-fit_power2 = glmer(resp_post_reg ~ 1 + verb_baseline_average + bot_lexical_typicality + participant_prereg_average + bot_regularisation_rate +
+fit_power2 = glmer(resp_post_reg ~ 
+    1 + verb_baseline_average + bot_lexical_typicality + participant_prereg_average + bot_regularisation_rate +
     (1 + verb_baseline_average | participant_id) + 
@@
 # summary(fit_power2)
+save(fit_power1, file = 'si_fit_power1.rda')
+save(fit_power2, file = 'si_fit_power2.rda')
 
@@
 <<celex_gcm_mgl1, eval=T, echo=F>>=
-
-baseline = baseline %>% 
-  mutate(
-    r_baseline_mgl_features = scales::rescale(baseline_mgl_features),
-    r_baseline_gcm_features = scales::rescale(baseline_gcm_features),
-  )
 
@@
 
-celex_m3 = glmer(regular ~ 1 + r_baseline_mgl_features + r_baseline_gcm_features_residual + 
+celex_m3 = glmer(regular ~ 
+                1 + r_baseline_mgl_features + r_baseline_gcm_features_residual + 
                 (1 + r_baseline_mgl_features + r_baseline_gcm_features_residual | subject), 
@@
 
-celex_m4 = glmer(regular ~ 1 + r_baseline_gcm_features + r_baseline_mgl_features_residual + 
+celex_m4 = glmer(regular ~ 
+                1 + r_baseline_gcm_features + r_baseline_mgl_features_residual + 
                 (1 + r_baseline_gcm_features + r_baseline_mgl_features_residual | subject), 
@@
 
-test2 = test2 %>% 
- mutate(
-    r_baseline_mgl_features = scales::rescale(baseline_mgl_features),
-    r_baseline_gcm_features = scales::rescale(baseline_gcm_features),
-    r_individual_mgl_features = scales::rescale(individual_mgl_features),
-    r_individual_gcm_features = scales::rescale(individual_gcm_features)
- )
-
 # C_baseline_mgl_features # celex-mgl on baseline
@@
 
-\subsection{Extra information provided by the individual model over the \sc{Celex}-model}
+\subsection{Extra information provided by the individual model over the {\sc Celex}-model}
 
-Second, we considered the amount of extra variation explained by the individual-model over the \sc{Celex}-model, for the MGL and the GCM, respectively.
+Second, we considered the amount of extra variation explained by the individual-model over the {\sc Celex}-model, for the MGL and the GCM, respectively.
 
-Section 6.4 in the main text explains that we calculate the extra information given by the individual models by subtracting the \sc{Celex} model prediction from the individual model prediction for each item for each participant.
+Section 6.4 in the main text explains that we calculate the extra information given by the individual models by subtracting the {\sc Celex} model prediction from the individual model prediction for each item for each participant.
 
@@
 
-test2 = test2 %>% 
-  mutate(
-    gcm_score_diff = r_individual_gcm_features - r_baseline_gcm_features,
-    mgl_score_diff = r_individual_mgl_features - r_baseline_mgl_features
-)
-
-resp_post_reg ~ 1 + verb_baseline_average * bot_lexical_typicality + participant_prereg_average + bot_regularisation_rate
-
-indiv_m5 = glmer(resp_post_reg ~ r_baseline_gcm_features + gcm_score_diff + participant_prereg_average + 
+indiv_m5 = glmer(resp_post_reg ~
+    1 + r_baseline_gcm_features + gcm_score_diff + participant_prereg_average + 
     (1 + r_baseline_gcm_features + gcm_score_diff | participant_id) + (1 | word), 
@@
 
-indiv_m6 = glmer(resp_post_reg ~ r_baseline_mgl_features + mgl_score_diff + participant_prereg_average  + 
+indiv_m6 = glmer(resp_post_reg ~ 
+    1 + r_baseline_mgl_features + mgl_score_diff + participant_prereg_average  + 
     (1 + r_baseline_mgl_features + mgl_score_diff | participant_id) + (1 | word), 
@@
 
-These techniques are indicative of the relative strengths of the individual-MGL and -GCM respective to their \sc{Celex}-counterparts and each other.
-However, they do not compare them directly using statistical tests.
+These techniques are indicative of the relative strengths of the individual-MGL and -GCM respective to their {\sc Celex}-counterparts and each other. However, they do not compare them directly.
 
 \subsection{Using combined model predictions to explain variation in participant post-test responses}
+
+<<comparison_model1, echo=F,eval=F>>=
+
+# the model reported in the paper 
+comp_fit1 = glmer(resp_post_reg ~ 
+    1 + r_baseline_mgl_features + r_individual_gcm_features + participant_prereg_average  + 
+    (1 + r_baseline_mgl_features + r_individual_gcm_features | participant_id) + (1 | word), 
+                 data = test2, 
+                 family = binomial (link = logit), 
+                 control=glmerControl(optimizer="bobyqa"))
+
+# model without baseline mgl
+comp_fit1b = glmer(resp_post_reg ~ 
+    1 + r_individual_gcm_features + participant_prereg_average  + 
+    (1 + r_individual_gcm_features | participant_id) + (1 | word), 
+                 data = test2, 
+                 family = binomial (link = logit), 
+                 control=glmerControl(optimizer="bobyqa"))
+
+# model without individual gcm
+comp_fit1c = glmer(resp_post_reg ~ 
+    1 + r_baseline_mgl_features + participant_prereg_average  + 
+    (1 + r_baseline_mgl_features | participant_id) + (1 | word), 
+                 data = test2, 
+                 family = binomial (link = logit), 
+                 control=glmerControl(optimizer="bobyqa"))
+
+# model with individual mgl
+comp_fit1d = glmer(resp_post_reg ~ 
+    1 + r_baseline_mgl_features + r_individual_gcm_features + r_individual_mgl_features + participant_prereg_average  + 
+    (1 + r_baseline_mgl_features + r_individual_gcm_features | participant_id) + (1 | word), 
+                 data = test2, 
+                 family = binomial (link = logit), 
+                 control=glmerControl(optimizer="bobyqa"))
+
+# model with celex gcm
+comp_fit1e = glmer(resp_post_reg ~ 
+    1 + r_baseline_mgl_features + r_individual_gcm_features + r_baseline_gcm_features + participant_prereg_average  + 
+    (1 + r_baseline_mgl_features + r_individual_gcm_features | participant_id) + (1 | word), 
+                 data = test2, 
+                 family = binomial (link = logit), 
+                 control=glmerControl(optimizer="bobyqa"))
+
+save(comp_fit1, file = 'si_comp_fit1.rda')
+save(comp_fit1b, file = 'si_comp_fit1b.rda')
+save(comp_fit1c, file = 'si_comp_fit1c.rda')
+save(comp_fit1d, file = 'si_comp_fit1d.rda')
+save(comp_fit1e, file = 'si_comp_fit1e.rda')
+
+@
+
+We used the {\sc Celex}-GCM, {\sc Celex}-MGL, individual-GCM, and individual-MGL together to predict participant post-test responses.
+The best model, reported in the paper, contains the {\sc Celex}-MGL and the individal-GCM.
+The summary can be seen in Table \ref{comp3}.
+
+<<comparison_model2, echo=F,eval=T>>=
+
+load('si_comp_fit1.rda')
+load('si_comp_fit1b.rda')
+load('si_comp_fit1c.rda')
+load('si_comp_fit1d.rda')
+load('si_comp_fit1e.rda')
+
+chisq1 = anova(comp_fit1, comp_fit1b)[2,6] %>% round(2) # without baseline mgl
+chisq2 = anova(comp_fit1, comp_fit1c)[2,6] %>% round(2) # without individual gcm
+chisq3 = anova(comp_fit1, comp_fit1d)[2,6] %>% round(2) # with individual mgl
+chisq4 = anova(comp_fit1, comp_fit1e)[2,6] %>% round(2) # with celex gcm
+
+@
+
+<<comparison_model3, echo=F,eval=T, results=tex>>=
+print(xtable(summary(comp_fit1)$coef, digits = 3, label = 'comp3', caption = 'Celex-MGL and Individual-GCM predict variation in the post-test'))
+@
+
+Goodness-of-fit tests show that model fit is significantly worse if we exclude {\sc Celex}-MGL ($\chi^2 = $ \Sexpr{chisq1}) or individual-GCM ($\chi^2 = $ \Sexpr{chisq2}) as predictors.
+It is not significantly better if we include individual-MGL ($\chi^2 = $ \Sexpr{chisq3}) or {\sc Celex}-GCM ($\chi^2 = $ \Sexpr{chisq4}).
+
+\subsection{Visualising model contributions to explaining variation in post-test responses}
+
+Figure \ref{bigfigure} shows how baseline and post-test regularisation means of words compare for participant responses, the GCM, and the MGL.
+
+<<model_figure, echo=F,eval=T>>=
+
+# can't figure first panel out
+
+first_panel = test2 %>% 
+  group_by(
+    word,
+    bot_lexical_typicality,
+    baseline_av
+  ) %>%
+  summarise(
+    post_reg_mean = post_reg_av %>% 
+      mean()
+  ) %>% 
+  ungroup() %>% 
+  mutate(
+    baseline = scales::rescale(baseline_av),
+    posttest = scales::rescale(post_reg_mean),
+    type = 'Participant responses'
+  )
+
+second_panel = test2 %>% 
+  group_by(
+    word,
+    bot_lexical_typicality,
+    r_baseline_gcm_features
+  ) %>% 
+  rename(
+    baseline = r_baseline_gcm_features
+  ) %>% 
+  summarise(
+    posttest = mean(r_individual_gcm_features)
+  ) %>% 
+  mutate(
+    type = 'GCM'
+  )
+
+third_panel = test2 %>% 
+  group_by(
+    word,
+    bot_lexical_typicality,
+    r_baseline_mgl_features
+  ) %>% 
+  rename(
+    baseline = r_baseline_mgl_features
+  ) %>% 
+  summarise(
+    posttest = mean(r_individual_mgl_features)
+  ) %>% 
+  mutate(
+    type = 'MGL'
+  )
+
+# three_studies = bind_rows(first_panel,second_panel,third_panel)  
+three_studies = bind_rows(second_panel,third_panel)
+
+p1 = ggplot(three_studies, aes(x = baseline, y = posttest, colour = bot_lexical_typicality, shape = bot_lexical_typicality, linetype = bot_lexical_typicality)) +
+  guides(size = F) +
+  geom_point() +
+  geom_smooth(se = F, method = 'loess', formula = 'y ~ x') +
+  facet_wrap( ~ type) +
+  scale_shape_discrete(name="Lexical\ntypicality") +
+  # scale_colour_discrete(name="Lexical\ntypicality") +
+  xlab('verb baseline regularization') +
+  ylab('verb ESP regularization') +
+  guides(linetype = F) +
+  scale_colour_brewer(palette = 'Set2', name="Lexical\ntypicality")
+
+@
+
+\begin{figure}
+<<model_figure2, fig=T,echo=F,eval=T, height = 4, width = 13>>=
+wrap_elements(grid::textGrob("I'm a \n giant \n turnip")) + p1
+@
+\caption{Baseline and individual regularisation weights: participants, GCM, MGL}
+\label{bigfigure}
+\end{figure}
+
+\newpage
 

### 2020-01-29 supplementary_information/si.tex
some progress

@@
 
-% latex table generated in R 3.6.0 by xtable 1.8-4 package
-% Wed Jan 29 18:07:06 2020
+% latex table generated in R 3.6.2 by xtable 1.8-4 package
+% Thu Jan 30 00:23:25 2020
 \begin{table}[ht]
@@
 
-% latex table generated in R 3.6.0 by xtable 1.8-4 package
-% Wed Jan 29 18:07:07 2020
+% latex table generated in R 3.6.2 by xtable 1.8-4 package
+% Thu Jan 30 00:23:25 2020
 \begin{table}[ht]
@@
 
-% latex table generated in R 3.6.0 by xtable 1.8-4 package
-% Wed Jan 29 18:07:07 2020
+% latex table generated in R 3.6.2 by xtable 1.8-4 package
+% Thu Jan 30 00:23:26 2020
 \begin{table}[ht]
@@
 \end{table}
-% latex table generated in R 3.6.0 by xtable 1.8-4 package
-% Wed Jan 29 18:07:07 2020
+% latex table generated in R 3.6.2 by xtable 1.8-4 package
+% Thu Jan 30 00:23:26 2020
 \begin{table}[ht]
@@
     \begin{tabular}{lllll}
-            & training data & method & test data & C (test data)\\
+            & training data & method & test data & C\\
             \hline
@@
 
-\subsection{Extra information provided by the individual model over the \sc{Celex}-model}
+\subsection{Extra information provided by the individual model over the {\sc Celex}-model}
 
-Second, we considered the amount of extra variation explained by the individual-model over the \sc{Celex}-model, for the MGL and the GCM, respectively.
+Second, we considered the amount of extra variation explained by the individual-model over the {\sc Celex}-model, for the MGL and the GCM, respectively.
 
-Section 6.4 in the main text explains that we calculate the extra information given by the individual models by subtracting the \textsc{Celex} model prediction from the individual model prediction for each item for each participant.
+Section 6.4 in the main text explains that we calculate the extra information given by the individual models by subtracting the {\sc Celex} model prediction from the individual model prediction for each item for each participant.
 
-% latex table generated in R 3.6.0 by xtable 1.8-4 package
-% Wed Jan 29 18:07:07 2020
+% latex table generated in R 3.6.2 by xtable 1.8-4 package
+% Thu Jan 30 00:23:27 2020
 \begin{table}[ht]
@@
 \end{table}
-% latex table generated in R 3.6.0 by xtable 1.8-4 package
-% Wed Jan 29 18:07:07 2020
+% latex table generated in R 3.6.2 by xtable 1.8-4 package
+% Thu Jan 30 00:23:27 2020
 \begin{table}[ht]
@@
 
-These techniques are indicative of the relative strengths of the individual-MGL and -GCM respective to their \sc{Celex}-counterparts and each other.
-However, they do not compare them directly using statistical tests.
+These techniques are indicative of the relative strengths of the individual-MGL and -GCM respective to their {\sc Celex}-counterparts and each other. However, they do not compare them directly.
 
 \subsection{Using combined model predictions to explain variation in participant post-test responses}
+
+We used the {\sc Celex}-GCM, {\sc Celex}-MGL, individual-GCM, and individual-MGL together to predict participant post-test responses.
+The best model, reported in the paper, contains the {\sc Celex}-MGL and the individal-GCM.
+The summary can be seen in Table \ref{comp3}.
+
+% latex table generated in R 3.6.2 by xtable 1.8-4 package
+% Thu Jan 30 00:23:27 2020
+\begin{table}[ht]
+\centering
+\begin{tabular}{rrrrr}
+  \hline
+ & Estimate & Std. Error & z value & Pr($>$$|$z$|$) \\ 
+  \hline
+(Intercept) & -2.942 & 0.270 & -10.896 & 0.000 \\ 
+  r\_baseline\_mgl\_features & 1.600 & 0.171 & 9.361 & 0.000 \\ 
+  r\_individual\_gcm\_features & 2.786 & 0.378 & 7.369 & 0.000 \\ 
+  participant\_prereg\_average & 5.437 & 0.585 & 9.298 & 0.000 \\ 
+   \hline
+\end{tabular}
+\caption{Celex-MGL and Individual-GCM predict variation in the post-test} 
+\label{comp3}
+\end{table}
+Goodness-of-fit tests show that model fit is significantly worse if we exclude {\sc Celex}-MGL ($\chi^2 = $ 201.86) or individual-GCM ($\chi^2 = $ 81.82) as predictors.
+It is not significantly better if we include individual-MGL ($\chi^2 = $ 0.09) or {\sc Celex}-GCM ($\chi^2 = $ 2.03).
+
+\subsection{Visualising model contributions to explaining variation in post-test responses}
+
+Figure \ref{bigfigure} shows how baseline and post-test regularisation means of words compare for participant responses, the GCM, and the MGL.
+
+\begin{figure}
+\includegraphics{si-model_figure2}
+\caption{Baseline and individual regularisation weights: participants, GCM, MGL}
+\label{bigfigure}
+\end{figure}
+
+\newpage
 

### 2020-01-30 supplementary_information/si.Rnw
for clay

@@
 
-# libs
-library(tidyverse)
-library(xtable)
-library(lme4)
-library(patchwork) # take this out
+# this SI uses tidy data manipulation -- for a quick primer, visit https://dplyr.tidyverse.org/
+
+# libraries
+library(tidyverse) # for data wrangling, figures
+library(purrr) # for functional programming
+library(stringr) # for string manipulation
+library(readr) # for reading
+library(scales) # for rescaling
+library(xtable) # for latex tables
+library(lme4) # for GLMM-s
 
@@
 baseline = baseline %>% 
-  mutate(
+  dplyr::mutate(
     r_baseline_mgl_features = scales::rescale(baseline_mgl_features), # rescale regularity weights
@@
 test2 = test2 %>% 
-  mutate(
+  dplyr::mutate(
     verb_baseline_average = baseline_av - mean(baseline_av, na.rm = T), # center verb average
     participant_prereg_average = pre_reg_av - mean(pre_reg_av, na.rm = T), # center participant average
-    bot_lexical_typicality = case_when( # set factor levels used in paper, order factor
+    bot_lexical_typicality = dplyr::case_when( # set factor levels used in paper, order factor
         lex_typicality == 'typical' ~ 'typical',
@@
 ## get the raw output of the baseline feature-distance mgl
-mgl_baseline_rules_raw = read_tsv('~/Github/RaczBecknerHayPierrehumbert2019/models/mgl/baseline_mgl/CELEXFull3.sum') %>% 
-  mutate(
-    form1 = str_replace_all(form1, '»', ''), # get rid of syllable boundary markers
-    form2 = str_replace_all(form2, '»', '')
+mgl_baseline_rules_raw = readr::read_tsv('~/Github/RaczBecknerHayPierrehumbert2019/models/mgl/baseline_mgl/CELEXFull3.sum') %>% 
+  dplyr::mutate(
+    form1 = stringr::str_replace_all(form1, '»', ''), # get rid of syllable boundary markers
+    form2 = stringr::str_replace_all(form2, '»', '')
   )
@@
 ## get verb handles
-verb_handles = read_csv('~/Github/RaczBecknerHayPierrehumbert2019/data/256verbs_print_forms.txt')
+verb_handles = readr::read_csv('~/Github/RaczBecknerHayPierrehumbert2019/data/256verbs_print_forms.txt')
 
@@
 
-We also discuss (\ref{appendixgcmmglresidualization}) the residualization method used to compare them, (\ref{posttestimplementation}) how well they fit our ESP test, and (\ref{appendixregression}) the regression methods used.
+We also discuss model selection in analysing our regression data (\ref{appendixregression}), how the two learning models compare on our baseline data (\ref{appendixgcmmglresidualization}), and how they compare on our ESP post-test data (\ref{mglgcmposttest}).
 
@@
 baseline %>% 
-  group_by(word) %>% 
-  summarise(mean = mean(regular)) %>% 
-  arrange(-mean)
+  dplyr::group_by(word) %>% 
+  dplyr::summarise(mean = mean(regular)) %>% 
+  dplyr::arrange(-mean)
 
@@
 gcm_table = celex %>% # take celex
-  filter(!is.na(ahpa)) %>% # remove words not in Albright and Hayes celex
-  mutate(
-    vowels = str_extract_all(ahpa, '[euVoEIi2U{Q34@a6}]'), # match vowels in the albright and hayes transcription
-    nvowels = map(vowels, ~ length(.)) # calculate how many vowels the verb has
-  ) %>% 
-  select(-vowels) %>% # we don't need vowels now
-  unnest(cols = c(nvowels)) %>%  # unnest nvowels
-  filter(word != 'forsake', nvowels == 1) %>% # drop any verb with more than 1 vowel also "forsake which survives this
-  select(category, moder, disc) %>% # keep category, moder, disc columns (= reg/irreg, drove/sang etc, transcription)
-  distinct() %>% # repetitions! 
-  count(category, moder) %>% # count drove/sang etc across reg/irreg
-  mutate( # if it has no category, it's a misc verb
-    moder = replace_na(moder, 'other')
-  ) %>% 
-  pivot_wider( # stretch table
+  dplyr::filter(!is.na(ahpa)) %>% # remove words not in Albright and Hayes celex
+  dplyr::mutate(
+    vowels = stringr::str_extract_all(ahpa, '[euVoEIi2U{Q34@a6}]'), # match vowels in the albright and hayes transcription
+    nvowels = purrr::map(vowels, ~ length(.)) # calculate how many vowels the verb has
+  ) %>% 
+  dplyr::select(-vowels) %>% # we don't need vowels now
+  tidyr::unnest(cols = c(nvowels)) %>%  # unnest nvowels
+  dplyr::filter(word != 'forsake', nvowels == 1) %>% # drop any verb with more than 1 vowel also "forsake which survives this
+  dplyr::select(category, moder, disc) %>% # keep category, moder, disc columns (= reg/irreg, drove/sang etc, transcription)
+  dplyr::distinct() %>% # repetitions! 
+  dplyr::count(category, moder) %>% # count drove/sang etc across reg/irreg
+  dplyr::mutate( # if it has no category, it's a misc verb
+    moder = tidyr::replace_na(moder, 'other')
+  ) %>% 
+  tidyr::pivot_wider( # stretch table
     names_from = category, values_from = n
   ) %>% 
-  rename(
+  dplyr::rename(
    `verb class` = moder,
@@
 
-reg_count = filter(gcm_table, `verb class` == 'other') %>% 
-  pull(`regular set`)
+reg_count = dplyr::filter(gcm_table, `verb class` == 'other') %>% 
+  dplyr::pull(`regular set`)
 
 gcm_table = gcm_table[1:4,] %>% 
-  mutate(
+  dplyr::mutate(
     `miscellaneous regular verbs` = reg_count
@@
 <<gcmtable, echo=F, eval=T, results=tex>>=
-print(xtable(gcm_table, digits=2, label="gcmsetsize", caption="Number of forms in each verb class, GCM training data"))
+print(xtable::xtable(gcm_table, digits=2, label="gcmsetsize", caption="Number of forms in each verb class, GCM training data"))
 
@@
 irreg_set = gcm_table %>% 
-  filter(`verb class` == 'DROVE') %>% 
-  pull(`irregular set`) %>% 
+  dplyr::filter(`verb class` == 'DROVE') %>% 
+  dplyr::pull(`irregular set`) %>% 
   as.double
@@
 reg_set = gcm_table %>% 
-  filter(`verb class` == 'DROVE') %>% 
-  pull(`regular set`) %>% 
+  dplyr::filter(`verb class` == 'DROVE') %>% 
+  dplyr::pull(`regular set`) %>% 
   as.double
@@
 splive_gcm = test2 %>% 
-  filter(word == 'splive') %>% 
-  pull(r_baseline_gcm_features) %>% 
+  dplyr::filter(word == 'splive') %>% 
+  dplyr::pull(r_baseline_gcm_features) %>% 
   unique %>% 
@@
 
-%The nonce form {\em splive} belongs to the DROVE class in our model.
+The nonce form {\em splive} belongs to the DROVE class in our model.
 The two past forms of {\em splive} in the experiment are regular {\em splived} and irregular {\em splove}.
@@
 disc_category = test2 %>% # get words and categories from test2. match colnames with above
-  distinct(disc, category) %>% 
-  mutate(
+  dplyr::distinct(disc, category) %>% 
+  dplyr::mutate(
     form1 = disc
@@
 disc_reg_past_irreg_past = verb_handles %>% # get reg and irreg transcriptions from verb handles. match colnames with above.
-  select(base.form,regular.form,irregular.form) %>% 
-  rename(
+  dplyr::select(base.form,regular.form,irregular.form) %>% 
+  dplyr::rename(
     form1 = base.form
@@
 mgl_baseline_rules_raw = mgl_baseline_rules_raw %>% 
-  filter(
+  dplyr::filter(
     !is.na(category)
@@
 mgl_baseline_rules_raw = mgl_baseline_rules_raw %>%  # group rules per output: reg / irreg!
-  mutate(
-    rule.type = case_when(
+  dplyr::mutate(
+    rule.type = dplyr::case_when(
       form2 == regular.form ~ 'regular',
@@
 mgl_baseline_rules_raw = mgl_baseline_rules_raw %>%
-  filter(
+  dplyr::filter(
 !is.na(rule.type)
@@
 mgl_count = mgl_baseline_rules_raw %>% # clean-up
-  mutate(
-    `related forms` = str_replace_all(`related forms`, '»', ''),
-    exceptions = str_replace_all(exceptions, '»', ''),
-    related_list = strsplit(`related forms`, ', '), # split
-    exception_list = strsplit(exceptions, ', ')
+  dplyr::mutate(
+    `related forms` = stringr::str_replace_all(`related forms`, '»', ''),
+    exceptions = stringr::str_replace_all(exceptions, '»', ''),
+    related_list = strsplit(`related forms`, ', '), # split related forms list, a long string, into a character vector, store the vector
+    exception_list = strsplit(exceptions, ', ') # same for exceptions
   )
@@
 mgl_count1 = mgl_count %>%
-  select(category,rule.type,related_list) %>% 
-  unnest(related_list) %>% 
-  distinct() %>% 
-  count(category, rule.type) %>% 
-  mutate(
-    type = 'related forms'
-  )
-
-mgl_count2 = mgl_count %>%
-  select(category,rule.type,exception_list) %>% 
-  unnest(exception_list) %>% 
-  distinct() %>% 
-  count(category, rule.type) %>% 
-  mutate(
+  dplyr::select(category,rule.type,related_list) %>% 
+  tidyr::unnest(related_list) %>% # we now have the related forms as one verb per line, which is tidy. but the verbs repeat, since the same verb can be a related form for several target verbs
+  dplyr::distinct() %>% # so we distinct them!
+  dplyr::count(category, rule.type) %>% # since one distinct verb related form == one row, we can count the rows
+  dplyr::mutate(
+    type = 'related forms' # by the way, this is related forms
+  )
+
+mgl_count2 = mgl_count %>% # we repeat this for exceptions
+  dplyr::select(category,rule.type,exception_list) %>% 
+  tidyr::unnest(exception_list) %>% 
+  dplyr::distinct() %>% 
+  dplyr::count(category, rule.type) %>% 
+  dplyr::mutate(
     type = 'exceptions'
@@
 
-mgl_table = rbind(mgl_count1,mgl_count2) %>% 
-  pivot_wider(
+mgl_table = rbind(mgl_count1,mgl_count2) %>%  # we combine the two
+  tidyr::pivot_wider( # stretch the data frame
     names_from = type, values_from = n
   ) %>% 
-  arrange(rule.type, category)
+  dplyr::arrange(rule.type, category)
 
@@
 <<mgltable, echo=F, eval=T, results=tex>>=
-print(xtable(mgl_table, digits=2, label="mglsetsize", caption="Number of forms in each verb class, MGL training data"))
+print(xtable::xtable(mgl_table, digits=2, label="mglsetsize", caption="Number of forms in each verb class, MGL training data"))
 
@@
 splive_mgl_rules = mgl_baseline_rules_raw %>% 
-  filter(
+  dplyr::filter(
     form1 == 'spl2v'
@@
 test2 %>% 
-  mutate(
-    r_baseline_mgl_features = scales::rescale(baseline_mgl_features)
-  ) %>% 
-  filter(word == 'splive') %>% 
-  select(baseline_mgl_features,r_baseline_mgl_features) %>% 
-  distinct()
+  dplyr::filter(word == 'splive') %>% 
+  dplyr::select(baseline_mgl_features,r_baseline_mgl_features) %>% 
+  dplyr::distinct()
 
@@
 
+# now we count rules, not words. easier.
 mgl_rule_count = mgl_baseline_rules_raw %>% # take raw rules
-  filter( # throw out non-category verbs and non reg/irreg rules (in our definition)
+  dplyr::filter( # throw out non-category verbs and non reg/irreg rules (in our definition)
     !is.na(category),
@@
   ) %>% 
-  mutate( # make a rule column of the form A -> B / C
+  dplyr::mutate( # make a rule column of the form A -> B / C
     rule = paste0(A, ' ->', B, ' / ', Pfeat)
   ) %>% 
-  distinct(rule,category,rule.type) %>% # keep DISTINCT rules
-  count(category, rule.type) %>%  # count them
-  pivot_wider(names_from = rule.type, values_from = n) %>% # stretch table
-  mutate( # make row sum
+  dplyr::distinct(rule,category,rule.type) %>% # keep DISTINCT rules
+  dplyr::count(category, rule.type) %>%  # count them
+  tidyr::pivot_wider(names_from = rule.type, values_from = n) %>% # stretch table
+  dplyr::mutate( # make row sum
     sum = irregular + regular
   ) %>% 
-  rename(
+  dplyr::rename( # make nice names
     `regular rule` = regular,
@@
 
-# glue on colsums
+# glue on column sums
 mgl_rule_count = rbind(mgl_rule_count, c('sum', sum(mgl_rule_count$`irregular rule`), sum(mgl_rule_count$`regular rule`), sum(mgl_rule_count$sum)))
@@
 <<mgltable, echo=F, eval=T, results=tex>>=
-print(xtable(mgl_table, digits=2, label="mglruleses", caption="The number of unique rules posited by the MGL in each verb class"))
+print(xtable::xtable(mgl_table, digits=2, label="mglruleses", caption="The number of unique rules posited by the MGL in each verb class"))
 
@@
 
+% directly piping in this table from R would be difficult, but it's lookupable
 \begin{table}[ht]
@@
 ##########################################
-# from Florian Jaeger
+# VIF for GLMM-s, from Florian Jaeger
 
@@
 
+# while the glm syntax estimates an intercept even if you don't spell it out, I spell it out, because, along with the syntax for the random intercept, it makes what's happening more evident to me.
+
+# all fits use the bobyqa optimiser which tends to have fewer convergence issues in my experience
+
 lme.fit1 = glmer(resp_post_reg ~ 
@@
 
-save(lme.fit1, file = 'si_fit1.rda')
-save(lme.fit1b, file = 'si_fit1b.rda')
-save(lme.fit1c, file = 'si_fit1c.rda')
+save(lme.fit1, file = 'models/si_fit1.rda')
+save(lme.fit1b, file = 'models/si_fit1b.rda')
+save(lme.fit1c, file = 'models/si_fit1c.rda')
 vif.mer(lme.fit1)
@@
 
+# we don't report this comparison here or in the main text, but the models are stored for reference
+
 @
@@
 <<main_model3, echo=F,eval=T, results=tex>>=
-load('si_fit1.rda')
-print(xtable(summary(lme.fit1)$coef, digits = 3, label = 'regressiontable1', caption = 'Experimental Data, Post-test Regression Model Summary.
+load('models/si_fit1.rda')
+print(xtable::xtable(summary(lme.fit1)$coef, digits = 3, label = 'regressiontable1', caption = 'Experimental Data, Post-test Regression Model Summary.
 Model formula: post-test regular response ~ verb baseline mean x lexical typicality + participant pretest mean + regularization shift + (1 + verb.baseline.mean | participant) + (1 | verb)'))
@@
 <<Bayes factor, echo=F, eval=F>>=
-# random vs atypical
+
+# we keep random vs atypical
 
 test2b = test2 %>% 
-  filter(
+  dplyr::filter(
     bot_lexical_typicality != 'typical'
@@
 # summary(fit_power2)
-save(fit_power1, file = 'si_fit_power1.rda')
-save(fit_power2, file = 'si_fit_power2.rda')
-
+save(fit_power1, file = 'models/si_fit_power1.rda')
+save(fit_power2, file = 'models/si_fit_power2.rda')
+
+# a simple way to calculate Bayes factors, should you wish to do so:
 exp((BIC(fit_power2) - BIC(fit_power1))/2)
@@
 <<celex_gcm_mgl1, eval=T, echo=F>>=
+
+# calculating concordance indices.
+# we DON'T load Hmisc in itself as it messes up function names
 
@@
 
+# get correlation between mgl and gcm scores in baseline
 mgl_gcm_cor = baseline %>% 
-  distinct(word,r_baseline_mgl_features,baseline_gcm_features) %>% 
-  summarise(
+  dplyr::distinct(word,r_baseline_mgl_features,baseline_gcm_features) %>% 
+  dplyr::summarise(
     cor = cor(r_baseline_mgl_features,baseline_gcm_features)
   ) %>% 
-  pull() %>% 
+  dplyr::pull() %>% 
   round(3)
@@
 
+# simple baseline words data frame
 baseline_words = baseline %>% 
-  distinct(word, r_baseline_mgl_features, r_baseline_gcm_features)
-
+  dplyr::distinct(word, r_baseline_mgl_features, r_baseline_gcm_features)
+
+# fitting models 1 and 2 on it
 celex_m1 = lm(r_baseline_gcm_features ~ 1 + r_baseline_mgl_features, data = baseline_words)
@@
 
+# getting residuals from models and putting them back in this data frame
 baseline_words = baseline_words %>% 
-  mutate(
+  dplyr::mutate(
     r_baseline_gcm_features_residual = resid(celex_m1),
@@
 
+# joining it with entire baseline data
 baseline = left_join(baseline, baseline_words, by = c("word", "r_baseline_mgl_features", "r_baseline_gcm_features"))
 
+# fitting models 3 and 4
 celex_m3 = glmer(regular ~ 
@@
 
-save(celex_m3, file = 'si_celex_m3.rda')
-save(celex_m4, file = 'si_celex_m4.rda')
+save(celex_m3, file = 'models/si_celex_m3.rda')
+save(celex_m4, file = 'models/si_celex_m4.rda')
 
@@
 <<celex_resid1, echo=F,eval=T, results=tex>>=
-load('si_celex_m3.rda')
-print(xtable(summary(celex_m3)$coef, digits = 3, label = 'resid1', caption = 'MGL scores and GCM residuals, baseline data'))
+load('models/si_celex_m3.rda')
+print(xtable::xtable(summary(celex_m3)$coef, digits = 3, label = 'resid1', caption = 'MGL scores and GCM residuals, baseline data'))
 @
@@
 <<celex_resid2, echo=F,eval=T, results=tex>>=
-load('si_celex_m4.rda')
-print(xtable(summary(celex_m4)$coef, digits = 3, label = 'resid2', caption = 'GCM scores and MGL residuals, baseline data'))
+load('models/si_celex_m4.rda')
+print(xtable::xtable(summary(celex_m4)$coef, digits = 3, label = 'resid2', caption = 'GCM scores and MGL residuals, baseline data'))
 @
@@
 
-\section{Model fits on the post-test data}
+\section{Model fits on the post-test data}\label{mglgcmposttest}
 
@@
 <<first, echo=F,eval=T>>=
+
+# concordance indices of everything on everything, saving into variables, piping into table below
 
@@
 
+# fitting models 5 and 6
+
 indiv_m5 = glmer(resp_post_reg ~
@@
 
-save(indiv_m5, file = 'si_indiv_m5.rda')
+save(indiv_m5, file = 'models/si_indiv_m5.rda')
 
@@
 
-save(indiv_m6, file = 'si_indiv_m6.rda')
+save(indiv_m6, file = 'models/si_indiv_m6.rda')
 
@@
 
-load('si_indiv_m5.rda')
-print(xtable(summary(indiv_m5)$coef, digits = 3, label = 'indiv1', caption = 'Celex-GCM and individual score difference, post-test data'))
+load('models/si_indiv_m5.rda')
+print(xtable::xtable(summary(indiv_m5)$coef, digits = 3, label = 'indiv1', caption = 'Celex-GCM and individual score difference, post-test data'))
 
@@
 <<indiv_diff2, echo=F,eval=T, results=tex>>=
-load('si_indiv_m6.rda')
-print(xtable(summary(indiv_m6)$coef, digits = 3, label = 'indiv2', caption = 'Celex-MGL and individual score difference, post-test data'))
+load('models/si_indiv_m6.rda')
+print(xtable::xtable(summary(indiv_m6)$coef, digits = 3, label = 'indiv2', caption = 'Celex-MGL and individual score difference, post-test data'))
 @
@@
 
-save(comp_fit1, file = 'si_comp_fit1.rda')
-save(comp_fit1b, file = 'si_comp_fit1b.rda')
-save(comp_fit1c, file = 'si_comp_fit1c.rda')
-save(comp_fit1d, file = 'si_comp_fit1d.rda')
-save(comp_fit1e, file = 'si_comp_fit1e.rda')
+save(comp_fit1, file = 'models/si_comp_fit1.rda')
+save(comp_fit1b, file = 'models/si_comp_fit1b.rda')
+save(comp_fit1c, file = 'models/si_comp_fit1c.rda')
+save(comp_fit1d, file = 'models/si_comp_fit1d.rda')
+save(comp_fit1e, file = 'models/si_comp_fit1e.rda')
 
@@
 
-load('si_comp_fit1.rda')
-load('si_comp_fit1b.rda')
-load('si_comp_fit1c.rda')
-load('si_comp_fit1d.rda')
-load('si_comp_fit1e.rda')
+# getting a whole lot of chisq values. p values are not reported, but can be guessed from chisq values / are lookupable.
+
+load('models/si_comp_fit1.rda')
+load('models/si_comp_fit1b.rda')
+load('models/si_comp_fit1c.rda')
+load('models/si_comp_fit1d.rda')
+load('models/si_comp_fit1e.rda')
 
@@
 <<comparison_model3, echo=F,eval=T, results=tex>>=
-print(xtable(summary(comp_fit1)$coef, digits = 3, label = 'comp3', caption = 'Celex-MGL and Individual-GCM predict variation in the post-test'))
+print(xtable::xtable(summary(comp_fit1)$coef, digits = 3, label = 'comp3', caption = 'Celex-MGL and Individual-GCM predict variation in the post-test'))
 @
@@
 
-# can't figure first panel out
-
+# big combination figure.
+
+# first, verb baseline regularisation averages and post-test averages across lex typicality
 first_panel = test2 %>% 
-  group_by(
+  dplyr::group_by(
     word,
     bot_lexical_typicality,
-    baseline_av
-  ) %>%
-  summarise(
-    post_reg_mean = post_reg_av %>% 
+    baseline_av # baseline average is there already
+  ) %>%
+  dplyr::summarise(
+    post_reg_mean = resp_post_reg %>% # we get mean post test regularisation per lex typ
       mean()
@@
   ungroup() %>% 
-  mutate(
-    baseline = scales::rescale(baseline_av),
+  dplyr::mutate( # we scale everything
+    baseline = scales::rescale(baseline_av), # we call them baseline and posttest respectively
     posttest = scales::rescale(post_reg_mean),
-    type = 'Participant responses'
-  )
-
+    type = 'Participant responses' # we name this
+  )
+
+# same for gcm weights. 
 second_panel = test2 %>% 
-  group_by(
+  dplyr::group_by(
     word,
@@
   ) %>% 
-  rename(
+  dplyr::rename(
     baseline = r_baseline_gcm_features
   ) %>% 
-  summarise(
+  dplyr::summarise(
     posttest = mean(r_individual_gcm_features)
   ) %>% 
-  mutate(
+  dplyr::mutate(
     type = 'GCM'
@@
 
+# same for MGL weights
 third_panel = test2 %>% 
-  group_by(
+  dplyr::group_by(
     word,
@@
   ) %>% 
-  rename(
+  dplyr::rename(
     baseline = r_baseline_mgl_features
   ) %>% 
-  summarise(
+  dplyr::summarise(
     posttest = mean(r_individual_mgl_features)
   ) %>% 
-  mutate(
+  dplyr::mutate(
     type = 'MGL'
@@
 
-# three_studies = bind_rows(first_panel,second_panel,third_panel)  
-three_studies = bind_rows(second_panel,third_panel)
-
+# bind the three data frames, set factor order for type so pretty plot
+three_studies = bind_rows(first_panel,second_panel,third_panel) %>% 
+  dplyr::mutate(
+    type = factor(type, levels = c('Participant responses', 'GCM', 'MGL'))
+  )
+
+# plot this, x is baseline, y is posttest.
+# colour AND shape are lexical typicality.
+# split panels across raw data / gcm / mgl.
+# add loess smooths.
 p1 = ggplot(three_studies, aes(x = baseline, y = posttest, colour = bot_lexical_typicality, shape = bot_lexical_typicality, linetype = bot_lexical_typicality)) +
@@
 \begin{figure}
-<<model_figure2, fig=T,echo=F,eval=T, height = 4, width = 13>>=
-wrap_elements(grid::textGrob("I'm a \n giant \n turnip")) + p1
+<<model_figure2, fig=T,echo=F,eval=T, width = 12.5, height = 4, fig.margin=T, fig.align = ‘center’>>=
+p1
 @
@@
 
+The Figure shows the average baseline regularity (x axis) and the post-test regularity (y axis) of the verbs in the ESP experiment.
+The post-test values vary across the dimension of co-player lexical typicality.
+Loess smooths show the general trajectory of the effects of typicality.
+In the first panel, regularity translates to aggregate participant responses in the baseline and in the ESP post-test.
+In the second and third panels, regularity translates to regularity scores provided by the {\sc Celex}- and individual-GCM and -MGL.
+Values are rescaled for better comparison.
+
+While the loess smooths are likely slightly overfit, they show the general trend in the raw data, which is echoed in the regression analysis in the main text: the "typical" trend of baseline-post-test difference is stronger than either the "random" or "reversed" trends.
+These are not very much different from each other.
+The differences are gradual.
+The GCM, while generally overestimating regularity, captures this gradual shift, since verb behaviour is shifted individually in the post-test.
+The MGL also overestimates regularisation.
+Since rules affect groups of verbs, rather than individual verbs, the post-test shifts are abrupt.
+The "typical" trend continues climbing after a longer plateau.
+However, the "random" and "reversed" trends trail off, where the "reversed" trend turns negative towards the higher tail of the distribution.
+
 \newpage

### 2020-01-30 supplementary_information/si.tex
for clay

@@
 % latex table generated in R 3.6.2 by xtable 1.8-4 package
-% Thu Jan 30 00:23:25 2020
+% Thu Jan 30 09:32:35 2020
 \begin{table}[ht]
@@
 % latex table generated in R 3.6.2 by xtable 1.8-4 package
-% Thu Jan 30 00:23:25 2020
+% Thu Jan 30 09:32:36 2020
 \begin{table}[ht]
@@
 % latex table generated in R 3.6.2 by xtable 1.8-4 package
-% Thu Jan 30 00:23:26 2020
+% Thu Jan 30 09:32:37 2020
 \begin{table}[ht]
@@
 % latex table generated in R 3.6.2 by xtable 1.8-4 package
-% Thu Jan 30 00:23:27 2020
+% Thu Jan 30 09:32:37 2020
 \begin{table}[ht]
@@
 % latex table generated in R 3.6.2 by xtable 1.8-4 package
-% Thu Jan 30 00:23:27 2020
+% Thu Jan 30 09:32:38 2020
 \begin{table}[ht]

### 2020-02-18 supplementary_information/si.Rnw
after Clays first round of comments

@@
   dplyr::mutate(
-    r_baseline_mgl_features = scales::rescale(baseline_mgl_features), # rescale regularity weights
-    r_baseline_gcm_features = scales::rescale(baseline_gcm_features),
+    rescaled_CELEX_mgl_feature_distance = scales::rescale(baseline_mgl_features), # rescale regularity weights
+    rescaled_CELEX_gcm_feature_distance = scales::rescale(baseline_gcm_features),
   )
@@
       factor(levels = c('nc', '-40%', '+40%')),
-    r_baseline_mgl_features = scales::rescale(baseline_mgl_features), # scale regular weights
-    r_baseline_gcm_features = scales::rescale(baseline_gcm_features),
-    r_individual_mgl_features = scales::rescale(individual_mgl_features),
-    r_individual_gcm_features = scales::rescale(individual_gcm_features),
-    gcm_score_diff = r_individual_gcm_features - r_baseline_gcm_features, # calculate score diff
-    mgl_score_diff = r_individual_mgl_features - r_baseline_mgl_features
+    rescaled_CELEX_mgl_feature_distance = scales::rescale(baseline_mgl_features), # scale regular weights
+    rescaled_CELEX_gcm_feature_distance = scales::rescale(baseline_gcm_features),
+    rescaled_individual_mgl_feature_distance = scales::rescale(individual_mgl_features),
+    rescaled_individual_gcm_feature_distance = scales::rescale(individual_gcm_features),
+    gcm_score_diff = rescaled_individual_gcm_feature_distance - rescaled_CELEX_gcm_feature_distance, # calculate score diff
+    mgl_score_diff = rescaled_individual_mgl_feature_distance - rescaled_CELEX_mgl_feature_distance
   )
@@
 However, this material is included here so that this Supplement can serve as a standalone document for readers who are interested in the technical details of our study.
+Additional material is reproduced from the main text, revisiting specific regression model summaries and additional methods we used to investigate contributions from the GCM and MGL.
 
@@
   dplyr::filter(word == 'splive') %>% 
-  dplyr::pull(r_baseline_gcm_features) %>% 
+  dplyr::pull(rescaled_CELEX_gcm_feature_distance) %>% 
   unique %>% 
@@
   dplyr::filter(word == 'splive') %>% 
-  dplyr::select(baseline_mgl_features,r_baseline_mgl_features) %>% 
+  dplyr::select(baseline_mgl_features,rescaled_CELEX_mgl_feature_distance) %>% 
   dplyr::distinct()
@@
 
-In the model in Table 5 in the main text, repeated here as Table \ref{regressiontable1}, we elect to mean center continuous variables to counteract collinearity that becomes evident when including the interaction between and {\em lexical typicality} .
+In the model in Table 5 in the main text, repeated here as Table \ref{regressiontable1}, we elect to mean center continuous variables to counteract collinearity that becomes evident when including the interaction between and {\em lexical typicality}\footnote{In the SI, all p values rounded to 0.00 should be taken to indicate p $<$ 0.001.}.
 Although some debate exists regarding the practice of transforming variables \citep{belsley1991conditioning,echambadi2007mean}, centering variables is common practice to remove nonessential ill conditioning, that is, relationships that inevitably exist between main and product terms \citep{aiken1991multiple,jaccard2003interaction, jaeger2008categorical}.
@@
 
-We fit the Celex-GCM and the Celex-MGL on the baseline data to see how well they predict regularisation responses by the baseline participants.
-The Celex-MGL outperforms the Celex-GCM, if slightly.
+We fit the \textsc{Celex-GCM} and the \textsc{Celex-MGL} on the baseline data to see how well they predict regularization responses by the baseline participants.
+The \textsc{Celex-MGL} outperforms the \textsc{Celex-GCM}, if slightly.
 Both models contribute independently to explaining variation in the data.
@@
 
-C_baseline_mgl_features = with(baseline, Hmisc::somers2(r_baseline_mgl_features,regular)['C']) %>% 
+C_CELEX_mgl_features = with(baseline, Hmisc::somers2(rescaled_CELEX_mgl_feature_distance,regular)['C']) %>% 
   round(3)
-C_baseline_gcm_features = with(baseline, Hmisc::somers2(r_baseline_gcm_features,regular)['C']) %>% 
+C_CELEX_gcm_features = with(baseline, Hmisc::somers2(rescaled_CELEX_gcm_feature_distance,regular)['C']) %>% 
   round(3)
@@
 mgl_gcm_cor = baseline %>% 
-  dplyr::distinct(word,r_baseline_mgl_features,baseline_gcm_features) %>% 
+  dplyr::distinct(word,rescaled_CELEX_mgl_feature_distance,baseline_gcm_features) %>% 
   dplyr::summarise(
-    cor = cor(r_baseline_mgl_features,baseline_gcm_features)
+    cor = cor(rescaled_CELEX_mgl_feature_distance,baseline_gcm_features)
   ) %>% 
@@
 
-The concordance indices of the Celex-MGL (\Sexpr{C_baseline_mgl_features}) and the Celex-MGL (\Sexpr{C_baseline_gcm_features}) on the baseline data are similar, with some edge retained by the MGL, as in the case of \citep{albright2003rules}.
+The concordance indices of the \textsc{Celex}-MGL (\Sexpr{C_CELEX_mgl_features}) and the \textsc{Celex}-MGL (\Sexpr{C_CELEX_gcm_features}) on the baseline data are similar, with some edge retained by the MGL, as in the case of \citep{albright2003rules}.
 Our concordance indices are calculated using the somers2 function in R's \texttt{Hmisc} package \citep{harrell2014hmisc}.
@@
 baseline_words = baseline %>% 
-  dplyr::distinct(word, r_baseline_mgl_features, r_baseline_gcm_features)
+  dplyr::distinct(word, rescaled_CELEX_mgl_feature_distance, rescaled_CELEX_gcm_feature_distance)
 
 # fitting models 1 and 2 on it
-celex_m1 = lm(r_baseline_gcm_features ~ 1 + r_baseline_mgl_features, data = baseline_words)
-celex_m2 = lm(r_baseline_mgl_features ~ 1 + r_baseline_gcm_features, data = baseline_words)
+celex_m1 = lm(rescaled_CELEX_gcm_feature_distance ~ 1 + rescaled_CELEX_mgl_feature_distance, data = baseline_words)
+celex_m2 = lm(rescaled_CELEX_mgl_feature_distance ~ 1 + rescaled_CELEX_gcm_feature_distance, data = baseline_words)
 
@@
   dplyr::mutate(
-    r_baseline_gcm_features_residual = resid(celex_m1),
-    r_baseline_mgl_features_residual = resid(celex_m2)
+    rescaled_CELEX_gcm_feature_distance_residual = resid(celex_m1),
+    rescaled_CELEX_mgl_feature_distance_residual = resid(celex_m2)
   )
@@
 # joining it with entire baseline data
-baseline = left_join(baseline, baseline_words, by = c("word", "r_baseline_mgl_features", "r_baseline_gcm_features"))
+baseline = left_join(baseline, baseline_words, by = c("word", "rescaled_CELEX_mgl_feature_distance", "rescaled_CELEX_gcm_feature_distance"))
 
@@
 celex_m3 = glmer(regular ~ 
-                1 + r_baseline_mgl_features + r_baseline_gcm_features_residual + 
-                (1 + r_baseline_mgl_features + r_baseline_gcm_features_residual | subject), 
+                1 + rescaled_CELEX_mgl_feature_distance + rescaled_CELEX_gcm_feature_distance_residual + 
+                (1 + rescaled_CELEX_mgl_feature_distance + rescaled_CELEX_gcm_feature_distance_residual | subject), 
                 data = baseline, 
@@
 celex_m4 = glmer(regular ~ 
-                1 + r_baseline_gcm_features + r_baseline_mgl_features_residual + 
-                (1 + r_baseline_gcm_features + r_baseline_mgl_features_residual | subject), 
+                1 + rescaled_CELEX_gcm_feature_distance + rescaled_CELEX_mgl_feature_distance_residual + 
+                (1 + rescaled_CELEX_gcm_feature_distance + rescaled_CELEX_mgl_feature_distance_residual | subject), 
                 data = baseline, 
@@
 load('models/si_celex_m3.rda')
-print(xtable::xtable(summary(celex_m3)$coef, digits = 3, label = 'resid1', caption = 'MGL scores and GCM residuals, baseline data'))
+print(xtable::xtable(summary(celex_m3)$coef, digits = 3, label = 'resid1', caption = 'Model 3: MGL scores and GCM residuals, baseline data'))
 @
@@
 load('models/si_celex_m4.rda')
-print(xtable::xtable(summary(celex_m4)$coef, digits = 3, label = 'resid2', caption = 'GCM scores and MGL residuals, baseline data'))
+print(xtable::xtable(summary(celex_m4)$coef, digits = 3, label = 'resid2', caption = 'Model 4: GCM scores and MGL residuals, baseline data'))
 @
@@
 
-We fit the Celex-GCM,the Celex-MGL, as well as the Individual-GCM and the Individual-MGL on the post-test data.
-Recall that the Celex-based models were trained on verbs in Celex.
-The individual-based models were trained for each participant, on Celex verbs as well as the co-player's responses in the ESP test.
+We fit the \textsc{Celex-GCM},the \textsc{Celex-MGL}, as well as the Individual-GCM and the Individual-MGL on the post-test data.
+Recall that the \textsc{Celex}-based models were trained on verbs in \textsc{Celex}.
+The individual-based models were trained for each participant, on \textsc{Celex} verbs as well as the co-player's responses in the ESP test.
 They were tested on the post-test for each participant, individually.
@@
 
-First, we calculated concordance indices for the Celex- and individual-models on the post-test data of the ESP experiment.
+First, we calculated concordance indices for the \textsc{Celex}- and individual-models on the post-test data of the ESP experiment.
 
@@
 
-# C_baseline_mgl_features # celex-mgl on baseline
-# C_baseline_gcm_features # celex-gcm on baseline
-
-C_test2_baseline_mgl_features = with(test2, Hmisc::somers2(r_baseline_mgl_features,resp_post_reg)['C']) %>% round(3)
-C_test2_baseline_gcm_features = with(test2, Hmisc::somers2(r_baseline_gcm_features,resp_post_reg)['C']) %>% round(3)
-C_test2_individual_mgl_features = with(test2, Hmisc::somers2(r_individual_mgl_features,resp_post_reg)['C']) %>% round(3)
-C_test2_individual_gcm_features = with(test2, Hmisc::somers2(r_individual_gcm_features,resp_post_reg)['C']) %>% round(3)
+# C_CELEX_mgl_features # celex-mgl on baseline
+# C_CELEX_gcm_features # celex-gcm on baseline
+
+C_test2_baseline_mgl_features = with(test2, Hmisc::somers2(rescaled_CELEX_mgl_feature_distance,resp_post_reg)['C']) %>% round(3)
+C_test2_baseline_gcm_features = with(test2, Hmisc::somers2(rescaled_CELEX_gcm_feature_distance,resp_post_reg)['C']) %>% round(3)
+C_test2_individual_mgl_features = with(test2, Hmisc::somers2(rescaled_individual_mgl_feature_distance,resp_post_reg)['C']) %>% round(3)
+C_test2_individual_gcm_features = with(test2, Hmisc::somers2(rescaled_individual_gcm_feature_distance,resp_post_reg)['C']) %>% round(3)
 
@@
 Table \ref{modellingsummary} summarizes the concordance indices between our data and the various categorization models.
-The \textsc{Celex}-MGL, outperforms the \textsc{Celex}-GCM on the post-test data to some extent, much like in the case of our baseline data.
+The \textsc{Celex}-MGL outperforms the \textsc{Celex}-GCM on the post-test data to some extent, much like in the case of our baseline data.
 The individual-GCM considerably improves on the \textsc{Celex}-GCM in terms of explaining the post-test data.
@@
             \hline
-\textsc{Celex}-GCM & real verbs & `analogy' & verbs in baseline &\Sexpr{C_baseline_gcm_features }\\
-\textsc{Celex}-MGL & real verbs & `rules' & verbs in baseline &\Sexpr{C_baseline_mgl_features }\\ \hline
+\textsc{Celex}-GCM & real verbs & `analogy' & verbs in baseline &\Sexpr{C_CELEX_gcm_features }\\
+\textsc{Celex}-MGL & real verbs & `rules' & verbs in baseline &\Sexpr{C_CELEX_mgl_features }\\ \hline
 \textsc{Celex}-GCM & real verbs & `analogy' & verbs in post-test &\Sexpr{C_test2_baseline_gcm_features}\\
@@
 indiv_m5 = glmer(resp_post_reg ~
-    1 + r_baseline_gcm_features + gcm_score_diff + participant_prereg_average + 
-    (1 + r_baseline_gcm_features + gcm_score_diff | participant_id) + (1 | word), 
+    1 + rescaled_CELEX_gcm_feature_distance + gcm_score_diff + participant_prereg_average + 
+    (1 + rescaled_CELEX_gcm_feature_distance + gcm_score_diff | participant_id) + (1 | word), 
                  data = test2, 
@@
 indiv_m6 = glmer(resp_post_reg ~ 
-    1 + r_baseline_mgl_features + mgl_score_diff + participant_prereg_average  + 
-    (1 + r_baseline_mgl_features + mgl_score_diff | participant_id) + (1 | word), 
+    1 + rescaled_CELEX_mgl_feature_distance + mgl_score_diff + participant_prereg_average  + 
+    (1 + rescaled_CELEX_mgl_feature_distance + mgl_score_diff | participant_id) + (1 | word), 
                  data = test2, 
@@
 load('models/si_indiv_m5.rda')
-print(xtable::xtable(summary(indiv_m5)$coef, digits = 3, label = 'indiv1', caption = 'Celex-GCM and individual score difference, post-test data'))
+print(xtable::xtable(summary(indiv_m5)$coef, digits = 3, label = 'indiv1', caption = 'Model 5: Celex-GCM and individual score difference, post-test data'))
 
@@
 load('models/si_indiv_m6.rda')
-print(xtable::xtable(summary(indiv_m6)$coef, digits = 3, label = 'indiv2', caption = 'Celex-MGL and individual score difference, post-test data'))
+print(xtable::xtable(summary(indiv_m6)$coef, digits = 3, label = 'indiv2', caption = 'Model 6: Celex-MGL and individual score difference, post-test data'))
 @
@@
 comp_fit1 = glmer(resp_post_reg ~ 
-    1 + r_baseline_mgl_features + r_individual_gcm_features + participant_prereg_average  + 
-    (1 + r_baseline_mgl_features + r_individual_gcm_features | participant_id) + (1 | word), 
+    1 + rescaled_CELEX_mgl_feature_distance + rescaled_individual_gcm_feature_distance + participant_prereg_average  + 
+    (1 + rescaled_CELEX_mgl_feature_distance + rescaled_individual_gcm_feature_distance | participant_id) + (1 | word), 
                  data = test2, 
@@
 comp_fit1b = glmer(resp_post_reg ~ 
-    1 + r_individual_gcm_features + participant_prereg_average  + 
-    (1 + r_individual_gcm_features | participant_id) + (1 | word), 
+    1 + rescaled_individual_gcm_feature_distance + participant_prereg_average  + 
+    (1 + rescaled_individual_gcm_feature_distance | participant_id) + (1 | word), 
                  data = test2, 
@@
 comp_fit1c = glmer(resp_post_reg ~ 
-    1 + r_baseline_mgl_features + participant_prereg_average  + 
-    (1 + r_baseline_mgl_features | participant_id) + (1 | word), 
+    1 + rescaled_CELEX_mgl_feature_distance + participant_prereg_average  + 
+    (1 + rescaled_CELEX_mgl_feature_distance | participant_id) + (1 | word), 
                  data = test2, 
@@
 comp_fit1d = glmer(resp_post_reg ~ 
-    1 + r_baseline_mgl_features + r_individual_gcm_features + r_individual_mgl_features + participant_prereg_average  + 
-    (1 + r_baseline_mgl_features + r_individual_gcm_features | participant_id) + (1 | word), 
+    1 + rescaled_CELEX_mgl_feature_distance + rescaled_individual_gcm_feature_distance + rescaled_individual_mgl_feature_distance + participant_prereg_average  + 
+    (1 + rescaled_CELEX_mgl_feature_distance + rescaled_individual_gcm_feature_distance | participant_id) + (1 | word), 
                  data = test2, 
@@
 comp_fit1e = glmer(resp_post_reg ~ 
-    1 + r_baseline_mgl_features + r_individual_gcm_features + r_baseline_gcm_features + participant_prereg_average  + 
-    (1 + r_baseline_mgl_features + r_individual_gcm_features | participant_id) + (1 | word), 
+    1 + rescaled_CELEX_mgl_feature_distance + rescaled_individual_gcm_feature_distance + rescaled_CELEX_gcm_feature_distance + participant_prereg_average  + 
+    (1 + rescaled_CELEX_mgl_feature_distance + rescaled_individual_gcm_feature_distance | participant_id) + (1 | word), 
                  data = test2, 
@@
 
-Figure \ref{bigfigure} shows how baseline and post-test regularisation means of words compare for participant responses, the GCM, and the MGL.
+Figure \ref{bigfigure} shows how baseline and post-test regularization means of words compare for participant responses, the GCM, and the MGL.
 
@@
     bot_lexical_typicality,
-    r_baseline_gcm_features
+    rescaled_CELEX_gcm_feature_distance
   ) %>% 
   dplyr::rename(
-    baseline = r_baseline_gcm_features
+    baseline = rescaled_CELEX_gcm_feature_distance
   ) %>% 
   dplyr::summarise(
-    posttest = mean(r_individual_gcm_features)
+    posttest = mean(rescaled_individual_gcm_feature_distance)
   ) %>% 
@@
     bot_lexical_typicality,
-    r_baseline_mgl_features
+    rescaled_CELEX_mgl_feature_distance
   ) %>% 
   dplyr::rename(
-    baseline = r_baseline_mgl_features
+    baseline = rescaled_CELEX_mgl_feature_distance
   ) %>% 
   dplyr::summarise(
-    posttest = mean(r_individual_mgl_features)
+    posttest = mean(rescaled_individual_mgl_feature_distance)
   ) %>% 
@@
 @
-\caption{Baseline and individual regularisation weights: participants, GCM, MGL}
+\caption{Baseline and individual regularization weights: participants, GCM, MGL}
 \label{bigfigure}
@@
 
-While the loess smooths are likely slightly overfit, they show the general trend in the raw data, which is echoed in the regression analysis in the main text: the "typical" trend of baseline-post-test difference is stronger than either the "random" or "reversed" trends.
+While the loess smooths are likely slightly overfit, they show the general trend in the raw data, which is echoed in the regression analysis in the main text: the `typical' trend of baseline-post-test difference is stronger than either the "random" or "reversed" trends.
 These are not very much different from each other.
 The differences are gradual.
-The GCM, while generally overestimating regularity, captures this gradual shift, since verb behaviour is shifted individually in the post-test.
-The MGL also overestimates regularisation.
+The GCM, while generally overestimating regularity, captures this gradual shift, since verb behavior is shifted individually in the post-test.
+The MGL also overestimates regularization.
 Since rules affect groups of verbs, rather than individual verbs, the post-test shifts are abrupt.
 The "typical" trend continues climbing after a longer plateau.
-However, the "random" and "reversed" trends trail off, where the "reversed" trend turns negative towards the higher tail of the distribution.
+However, the `random' and `reversed' trends trail off, where the "reversed" trend turns negative towards the higher tail of the distribution.
 

### 2020-02-18 supplementary_information/si.tex
after Clays first round of comments

@@
 
-% latex table generated in R 3.6.2 by xtable 1.8-4 package
-% Thu Jan 30 09:32:35 2020
+% latex table generated in R 3.6.0 by xtable 1.8-4 package
+% Tue Feb 18 16:57:29 2020
 \begin{table}[ht]
@@
 
-% latex table generated in R 3.6.2 by xtable 1.8-4 package
-% Thu Jan 30 09:32:36 2020
+% latex table generated in R 3.6.0 by xtable 1.8-4 package
+% Tue Feb 18 16:57:29 2020
 \begin{table}[ht]
@@
 
-We fit the Celex-GCM and the Celex-MGL on the baseline data to see how well they predict regularisation responses by the baseline participants.
-The Celex-MGL outperforms the Celex-GCM, if slightly.
+We fit the \textsc{Celex-GCM} and the \textsc{Celex-MGL} on the baseline data to see how well they predict regularization responses by the baseline participants.
+The \textsc{Celex-MGL} outperforms the \textsc{Celex-GCM}, if slightly.
 Both models contribute independently to explaining variation in the data.
 
-The concordance indices of the Celex-MGL (0.586) and the Celex-MGL (0.581) on the baseline data are similar, with some edge retained by the MGL, as in the case of \citep{albright2003rules}.
+The concordance indices of the \textsc{Celex}-MGL (0.586) and the \textsc{Celex}-MGL (0.581) on the baseline data are similar, with some edge retained by the MGL, as in the case of \citep{albright2003rules}.
 Our concordance indices are calculated using the somers2 function in R's \texttt{Hmisc} package \citep{harrell2014hmisc}.
@@
 
-% latex table generated in R 3.6.2 by xtable 1.8-4 package
-% Thu Jan 30 09:32:37 2020
+% latex table generated in R 3.6.0 by xtable 1.8-4 package
+% Tue Feb 18 16:57:30 2020
 \begin{table}[ht]
@@
 (Intercept) & -0.525 & 0.124 & -4.219 & 0.000 \\ 
-  r\_baseline\_mgl\_features & 1.347 & 0.093 & 14.482 & 0.000 \\ 
-  r\_baseline\_gcm\_features\_residual & 1.423 & 0.128 & 11.100 & 0.000 \\ 
+  rescaled\_CELEX\_mgl\_feature\_distance & 1.347 & 0.093 & 14.482 & 0.000 \\ 
+  rescaled\_CELEX\_gcm\_feature\_distance\_residual & 1.423 & 0.128 & 11.100 & 0.000 \\ 
    \hline
 \end{tabular}
-\caption{MGL scores and GCM residuals, baseline data} 
+\caption{Model 3: MGL scores and GCM residuals, baseline data} 
 \label{resid1}
 \end{table}
-% latex table generated in R 3.6.2 by xtable 1.8-4 package
-% Thu Jan 30 09:32:37 2020
+% latex table generated in R 3.6.0 by xtable 1.8-4 package
+% Tue Feb 18 16:57:31 2020
 \begin{table}[ht]
@@
 (Intercept) & -1.225 & 0.160 & -7.682 & 0.000 \\ 
-  r\_baseline\_gcm\_features & 2.391 & 0.149 & 16.063 & 0.000 \\ 
-  r\_baseline\_mgl\_features\_residual & 1.008 & 0.090 & 11.154 & 0.000 \\ 
+  rescaled\_CELEX\_gcm\_feature\_distance & 2.391 & 0.149 & 16.063 & 0.000 \\ 
+  rescaled\_CELEX\_mgl\_feature\_distance\_residual & 1.008 & 0.090 & 11.154 & 0.000 \\ 
    \hline
 \end{tabular}
-\caption{GCM scores and MGL residuals, baseline data} 
+\caption{Model 4: GCM scores and MGL residuals, baseline data} 
 \label{resid2}
@@
 
-First, we calculated concordance indices for the Celex- and individual-models on the post-test data of the ESP experiment.
+First, we calculated concordance indices for the \textsc{Celex}- and individual-models on the post-test data of the ESP experiment.
 
 Table \ref{modellingsummary} summarizes the concordance indices between our data and the various categorization models.
-The \textsc{Celex}-MGL, outperforms the \textsc{Celex}-GCM on the post-test data to some extent, much like in the case of our baseline data.
+The \textsc{Celex}-MGL outperforms the \textsc{Celex}-GCM on the post-test data to some extent, much like in the case of our baseline data.
 The individual-GCM considerably improves on the \textsc{Celex}-GCM in terms of explaining the post-test data.
@@
 
-% latex table generated in R 3.6.2 by xtable 1.8-4 package
-% Thu Jan 30 09:32:37 2020
+% latex table generated in R 3.6.0 by xtable 1.8-4 package
+% Tue Feb 18 16:57:31 2020
 \begin{table}[ht]
@@
 (Intercept) & -2.951 & 0.295 & -10.007 & 0.000 \\ 
-  r\_baseline\_gcm\_features & 4.369 & 0.425 & 10.284 & 0.000 \\ 
+  rescaled\_CELEX\_gcm\_feature\_distance & 4.369 & 0.425 & 10.284 & 0.000 \\ 
   gcm\_score\_diff & 2.612 & 0.470 & 5.563 & 0.000 \\ 
@@
 \end{tabular}
-\caption{Celex-GCM and individual score difference, post-test data} 
+\caption{Model 5: Celex-GCM and individual score difference, post-test data} 
 \label{indiv1}
 \end{table}
-% latex table generated in R 3.6.2 by xtable 1.8-4 package
-% Thu Jan 30 09:32:37 2020
+% latex table generated in R 3.6.0 by xtable 1.8-4 package
+% Tue Feb 18 16:57:31 2020
 \begin{table}[ht]
@@
 (Intercept) & -1.318 & 0.152 & -8.663 & 0.000 \\ 
-  r\_baseline\_mgl\_features & 1.944 & 0.176 & 11.045 & 0.000 \\ 
+  rescaled\_CELEX\_mgl\_feature\_distance & 1.944 & 0.176 & 11.045 & 0.000 \\ 
   mgl\_score\_diff & 0.304 & 0.139 & 2.181 & 0.029 \\ 
@@
 \end{tabular}
-\caption{Celex-MGL and individual score difference, post-test data} 
+\caption{Model 6: Celex-MGL and individual score difference, post-test data} 
 \label{indiv2}
@@
 
-% latex table generated in R 3.6.2 by xtable 1.8-4 package
-% Thu Jan 30 09:32:38 2020
+% latex table generated in R 3.6.0 by xtable 1.8-4 package
+% Tue Feb 18 16:57:31 2020
 \begin{table}[ht]
@@
 \includegraphics{si-model_figure2}
-\caption{Baseline and individual regularisation weights: participants, GCM, MGL}
+\caption{Baseline and individual regularization weights: participants, GCM, MGL}
 \label{bigfigure}

### 2020-02-19 supplementary_information/si.Rnw
mgl sum typo

@@
     form1 = stringr::str_replace_all(form1, '»', ''), # get rid of syllable boundary markers
-    form2 = stringr::str_replace_all(form2, '»', '')
+    form2 = stringr::str_replace_all(form2, '»', ''),
+    form2 = stringr::str_replace_all(form2, 'o', '5') # transcription thing
   )
@@
   )
+
+# verb_handles
 

### 2020-02-19 supplementary_information/si.tex
mgl sum typo

@@
 % latex table generated in R 3.6.0 by xtable 1.8-4 package
-% Tue Feb 18 16:57:29 2020
+% Wed Feb 19 11:30:14 2020
 \begin{table}[ht]
@@
 1 & burnt & irregular &  24 &  41 \\ 
-  2 & kept & irregular &  10 &   9 \\ 
-  3 & sang & irregular &  11 &  77 \\ 
-  4 & burnt & regular &  38 &  21 \\ 
-  5 & drove & regular &  29 &  27 \\ 
-  6 & kept & regular &  67 &  41 \\ 
-  7 & sang & regular &  47 &  38 \\ 
+  2 & drove & irregular &  21 & 114 \\ 
+  3 & kept & irregular &  10 &   9 \\ 
+  4 & sang & irregular &  11 &  77 \\ 
+  5 & burnt & regular &  38 &  21 \\ 
+  6 & drove & regular &  29 &  27 \\ 
+  7 & kept & regular &  67 &  41 \\ 
+  8 & sang & regular &  47 &  38 \\ 
    \hline
@@
 % latex table generated in R 3.6.0 by xtable 1.8-4 package
-% Tue Feb 18 16:57:30 2020
+% Wed Feb 19 11:30:16 2020
 \begin{table}[ht]
@@
 % latex table generated in R 3.6.0 by xtable 1.8-4 package
-% Tue Feb 18 16:57:31 2020
+% Wed Feb 19 11:30:16 2020
 \begin{table}[ht]
@@
 % latex table generated in R 3.6.0 by xtable 1.8-4 package
-% Tue Feb 18 16:57:31 2020
+% Wed Feb 19 11:30:17 2020
 \begin{table}[ht]

### 2020-02-20 supplementary_information/si.Rnw
nice sig

@@
 library(lme4) # for GLMM-s
+library(broom) # for tidy summaries
+
+# functions
+
+niceCoefTable = function(fit){
+  my_table = fit %>% 
+    tidy %>% 
+    filter(
+      group == 'fixed'
+    ) %>% 
+    mutate(
+      sig = case_when(
+        p.value >= 0.05 ~ 'n.s.',
+        p.value < 0.05 & p.value >= 0.01 ~ '*',
+        p.value < 0.01 & p.value >= 0.001 ~ '**',
+        p.value < 0.001 ~ '***'
+      )
+    ) %>% 
+    select(
+      - p.value,
+      - group
+    )
+return(my_table)  
+}
+
+# VIF for GLMM-s, from Florian Jaeger
+
+vif.mer <- function (fit) {
+  ## adapted from rms::vif
+  v <- vcov(fit)
+  nam <- names(fixef(fit))
+  ## exclude intercepts
+  ns <- sum(1 * (nam == "Intercept" | nam == "(Intercept)"))
+  if (ns > 0) {
+    v <- v[-(1:ns), -(1:ns), drop = FALSE]
+    nam <- nam[-(1:ns)] }
+  d <- diag(v)^0.5
+  v <- diag(solve(v/(d %o% d)))
+  names(v) <- nam 
+  v 
+  }
 
@@
 # regular
-splive_mgl_rules[3,]$Pfeat
-splive_mgl_rules[3,]$hits
-splive_mgl_rules[3,]$scope
-splive_mgl_rules[3,]$reliability
-splive_mgl_rules[3,]$confidence
+splive_mgl_rules[splive_mgl_rules$B == 'd',]$Pfeat
+splive_mgl_rules[splive_mgl_rules$B == 'd',]$hits
+splive_mgl_rules[splive_mgl_rules$B == 'd',]$scope
+splive_mgl_rules[splive_mgl_rules$B == 'd',]$reliability
+splive_mgl_rules[splive_mgl_rules$B == 'd',]$confidence
 
 # irregular
-splive_mgl_rules[1,]$Pfeat
-splive_mgl_rules[1,]$hits
-splive_mgl_rules[1,]$scope
-splive_mgl_rules[1,]$`related forms`
-splive_mgl_rules[1,]$exceptions
-splive_mgl_rules[1,]$reliability
-splive_mgl_rules[1,]$confidence
+splive_mgl_rules[splive_mgl_rules$B == 'o',]$Pfeat # this kept the original transcription that we change to DISC 5 in form2
+splive_mgl_rules[splive_mgl_rules$B == 'o',]$hits
+splive_mgl_rules[splive_mgl_rules$B == 'o',]$scope
+splive_mgl_rules[splive_mgl_rules$B == 'o',]$`related forms`
+splive_mgl_rules[splive_mgl_rules$B == 'o',]$exceptions
+splive_mgl_rules[splive_mgl_rules$B == 'o',]$reliability
+splive_mgl_rules[splive_mgl_rules$B == 'o',]$confidence
 
@@
 The {\em relative (adjusted) confidence} of the predicted regularity of {\em splive} is $0.98/(0.98+0.37)=0.73$.
-The respective rescaled value is 0.363.
+(The respective rescaled value is 0.363 -- which, in itself, is not very informative.)
 
@@
 
-<<main_model1, echo=F,eval=T>>=
-
-##########################################
-# VIF for GLMM-s, from Florian Jaeger
-
-vif.mer <- function (fit) {
-  ## adapted from rms::vif
-  v <- vcov(fit)
-  nam <- names(fixef(fit))
-  ## exclude intercepts
-  ns <- sum(1 * (nam == "Intercept" | nam == "(Intercept)"))
-  if (ns > 0) {
-    v <- v[-(1:ns), -(1:ns), drop = FALSE]
-    nam <- nam[-(1:ns)] }
-  d <- diag(v)^0.5
-  v <- diag(solve(v/(d %o% d)))
-  names(v) <- nam 
-  v 
-  }
-@
-
 <<main_model2, echo=F,eval=F>>=
@@
 load('models/si_fit1.rda')
-print(xtable::xtable(summary(lme.fit1)$coef, digits = 3, label = 'regressiontable1', caption = 'Experimental Data, Post-test Regression Model Summary.
+lme.fit1.sum = niceCoefTable(lme.fit1)
+
+print(xtable::xtable(lme.fit1.sum, digits = 2, label = 'regressiontable1', caption = 'Experimental Data, Post-test Regression Model Summary.
 Model formula: post-test regular response ~ verb baseline mean x lexical typicality + participant pretest mean + regularization shift + (1 + verb.baseline.mean | participant) + (1 | verb)'))
@@
 load('models/si_celex_m3.rda')
-print(xtable::xtable(summary(celex_m3)$coef, digits = 3, label = 'resid1', caption = 'Model 3: MGL scores and GCM residuals, baseline data'))
+celex_m3_summary = niceCoefTable(celex_m3)
+print(xtable::xtable(celex_m3_summary, digits = 2, label = 'resid1', caption = 'Model 3: MGL scores and GCM residuals, baseline data'))
 @
@@
 load('models/si_celex_m4.rda')
-print(xtable::xtable(summary(celex_m4)$coef, digits = 3, label = 'resid2', caption = 'Model 4: GCM scores and MGL residuals, baseline data'))
+celex_m4_summary = niceCoefTable(celex_m4)
+print(xtable::xtable(celex_m4_summary, digits = 2, label = 'resid2', caption = 'Model 4: GCM scores and MGL residuals, baseline data'))
 @
@@
 <<indiv_diff1, echo=F, eval=T, results=tex>>=
-
 load('models/si_indiv_m5.rda')
-print(xtable::xtable(summary(indiv_m5)$coef, digits = 3, label = 'indiv1', caption = 'Model 5: Celex-GCM and individual score difference, post-test data'))
+indiv_m5_sum = niceCoefTable(indiv_m5)
+print(xtable::xtable(indiv_m5_sum, digits = 2, label = 'indiv1', caption = 'Model 5: Celex-GCM and individual score difference, post-test data'))
 
@@
 load('models/si_indiv_m6.rda')
-print(xtable::xtable(summary(indiv_m6)$coef, digits = 3, label = 'indiv2', caption = 'Model 6: Celex-MGL and individual score difference, post-test data'))
+indiv_m6_sum = niceCoefTable(indiv_m6)
+print(xtable::xtable(indiv_m6_sum, digits = 3, label = 'indiv2', caption = 'Model 6: Celex-MGL and individual score difference, post-test data'))
 @
@@
 <<comparison_model3, echo=F,eval=T, results=tex>>=
-print(xtable::xtable(summary(comp_fit1)$coef, digits = 3, label = 'comp3', caption = 'Celex-MGL and Individual-GCM predict variation in the post-test'))
+comp_fit1_sum = niceCoefTable(comp_fit1)
+print(xtable::xtable(comp_fit1_sum, digits = 2, label = 'comp3', caption = 'Celex-MGL and Individual-GCM predict variation in the post-test'))
 @

### 2020-02-20 supplementary_information/si.tex
nice sig

@@
 % latex table generated in R 3.6.0 by xtable 1.8-4 package
-% Wed Feb 19 11:30:14 2020
+% Thu Feb 20 11:18:10 2020
 \begin{table}[ht]
@@
 % latex table generated in R 3.6.0 by xtable 1.8-4 package
-% Wed Feb 19 11:30:14 2020
+% Thu Feb 20 11:18:10 2020
 \begin{table}[ht]
 \centering
-\begin{tabular}{rrrrr}
+\begin{tabular}{rlrrrl}
   \hline
- & Estimate & Std. Error & z value & Pr($>$$|$z$|$) \\ 
+ & term & estimate & std.error & statistic & sig \\ 
   \hline
-(Intercept) & -0.108 & 0.161 & -0.672 & 0.501 \\ 
-  verb\_baseline\_average & 8.259 & 0.686 & 12.040 & 0.000 \\ 
-  bot\_lexical\_typicalityrandom & 0.043 & 0.169 & 0.253 & 0.800 \\ 
-  bot\_lexical\_typicalityreversed & 0.035 & 0.173 & 0.204 & 0.839 \\ 
-  participant\_prereg\_average & 7.125 & 0.563 & 12.667 & 0.000 \\ 
-  bot\_regularisation\_rate-40\% & -0.456 & 0.185 & -2.468 & 0.014 \\ 
-  bot\_regularisation\_rate+40\% & 0.549 & 0.166 & 3.307 & 0.001 \\ 
-  verb\_baseline\_average:bot\_lexical\_typicalityrandom & -2.402 & 0.927 & -2.591 & 0.010 \\ 
-  verb\_baseline\_average:bot\_lexical\_typicalityreversed & -2.127 & 0.955 & -2.226 & 0.026 \\ 
+1 & (Intercept) & -0.11 & 0.16 & -0.67 & n.s. \\ 
+  2 & verb\_baseline\_average & 8.26 & 0.69 & 12.04 & *** \\ 
+  3 & bot\_lexical\_typicalityrandom & 0.04 & 0.17 & 0.25 & n.s. \\ 
+  4 & bot\_lexical\_typicalityreversed & 0.04 & 0.17 & 0.20 & n.s. \\ 
+  5 & participant\_prereg\_average & 7.13 & 0.56 & 12.67 & *** \\ 
+  6 & bot\_regularisation\_rate-40\% & -0.46 & 0.18 & -2.47 & * \\ 
+  7 & bot\_regularisation\_rate+40\% & 0.55 & 0.17 & 3.31 & *** \\ 
+  8 & verb\_baseline\_average:bot\_lexical\_typicalityrandom & -2.40 & 0.93 & -2.59 & ** \\ 
+  9 & verb\_baseline\_average:bot\_lexical\_typicalityreversed & -2.13 & 0.96 & -2.23 & * \\ 
    \hline
@@
 % latex table generated in R 3.6.0 by xtable 1.8-4 package
-% Wed Feb 19 11:30:16 2020
+% Thu Feb 20 11:18:11 2020
 \begin{table}[ht]
 \centering
-\begin{tabular}{rrrrr}
+\begin{tabular}{rlrrrl}
   \hline
- & Estimate & Std. Error & z value & Pr($>$$|$z$|$) \\ 
+ & term & estimate & std.error & statistic & sig \\ 
   \hline
-(Intercept) & -0.525 & 0.124 & -4.219 & 0.000 \\ 
-  rescaled\_CELEX\_mgl\_feature\_distance & 1.347 & 0.093 & 14.482 & 0.000 \\ 
-  rescaled\_CELEX\_gcm\_feature\_distance\_residual & 1.423 & 0.128 & 11.100 & 0.000 \\ 
+1 & (Intercept) & -0.53 & 0.12 & -4.22 & *** \\ 
+  2 & rescaled\_CELEX\_mgl\_feature\_distance & 1.35 & 0.09 & 14.48 & *** \\ 
+  3 & rescaled\_CELEX\_gcm\_feature\_distance\_residual & 1.42 & 0.13 & 11.10 & *** \\ 
    \hline
@@
 % latex table generated in R 3.6.0 by xtable 1.8-4 package
-% Wed Feb 19 11:30:16 2020
+% Thu Feb 20 11:18:12 2020
 \begin{table}[ht]
 \centering
-\begin{tabular}{rrrrr}
+\begin{tabular}{rlrrrl}
   \hline
- & Estimate & Std. Error & z value & Pr($>$$|$z$|$) \\ 
+ & term & estimate & std.error & statistic & sig \\ 
   \hline
-(Intercept) & -1.225 & 0.160 & -7.682 & 0.000 \\ 
-  rescaled\_CELEX\_gcm\_feature\_distance & 2.391 & 0.149 & 16.063 & 0.000 \\ 
-  rescaled\_CELEX\_mgl\_feature\_distance\_residual & 1.008 & 0.090 & 11.154 & 0.000 \\ 
+1 & (Intercept) & -1.23 & 0.16 & -7.68 & *** \\ 
+  2 & rescaled\_CELEX\_gcm\_feature\_distance & 2.39 & 0.15 & 16.06 & *** \\ 
+  3 & rescaled\_CELEX\_mgl\_feature\_distance\_residual & 1.01 & 0.09 & 11.15 & *** \\ 
    \hline
@@
 % latex table generated in R 3.6.0 by xtable 1.8-4 package
-% Wed Feb 19 11:30:16 2020
+% Thu Feb 20 11:18:12 2020
 \begin{table}[ht]
 \centering
-\begin{tabular}{rrrrr}
+\begin{tabular}{rlrrrl}
   \hline
- & Estimate & Std. Error & z value & Pr($>$$|$z$|$) \\ 
+ & term & estimate & std.error & statistic & sig \\ 
   \hline
-(Intercept) & -2.951 & 0.295 & -10.007 & 0.000 \\ 
-  rescaled\_CELEX\_gcm\_feature\_distance & 4.369 & 0.425 & 10.284 & 0.000 \\ 
-  gcm\_score\_diff & 2.612 & 0.470 & 5.563 & 0.000 \\ 
-  participant\_prereg\_average & 5.289 & 0.602 & 8.788 & 0.000 \\ 
+1 & (Intercept) & -2.95 & 0.29 & -10.01 & *** \\ 
+  2 & rescaled\_CELEX\_gcm\_feature\_distance & 4.37 & 0.42 & 10.28 & *** \\ 
+  3 & gcm\_score\_diff & 2.61 & 0.47 & 5.56 & *** \\ 
+  4 & participant\_prereg\_average & 5.29 & 0.60 & 8.79 & *** \\ 
    \hline
@@
 % latex table generated in R 3.6.0 by xtable 1.8-4 package
-% Wed Feb 19 11:30:17 2020
+% Thu Feb 20 11:18:12 2020
 \begin{table}[ht]
 \centering
-\begin{tabular}{rrrrr}
+\begin{tabular}{rlrrrl}
   \hline
- & Estimate & Std. Error & z value & Pr($>$$|$z$|$) \\ 
+ & term & estimate & std.error & statistic & sig \\ 
   \hline
-(Intercept) & -1.318 & 0.152 & -8.663 & 0.000 \\ 
-  rescaled\_CELEX\_mgl\_feature\_distance & 1.944 & 0.176 & 11.045 & 0.000 \\ 
-  mgl\_score\_diff & 0.304 & 0.139 & 2.181 & 0.029 \\ 
-  participant\_prereg\_average & 7.203 & 0.591 & 12.190 & 0.000 \\ 
+1 & (Intercept) & -1.318 & 0.152 & -8.663 & *** \\ 
+  2 & rescaled\_CELEX\_mgl\_feature\_distance & 1.944 & 0.176 & 11.045 & *** \\ 
+  3 & mgl\_score\_diff & 0.304 & 0.139 & 2.181 & * \\ 
+  4 & participant\_prereg\_average & 7.203 & 0.591 & 12.190 & *** \\ 
    \hline
@@
 % latex table generated in R 3.6.0 by xtable 1.8-4 package
-% Wed Feb 19 11:30:17 2020
+% Thu Feb 20 11:18:12 2020
 \begin{table}[ht]
 \centering
-\begin{tabular}{rrrrr}
+\begin{tabular}{rlrrrl}
   \hline
- & Estimate & Std. Error & z value & Pr($>$$|$z$|$) \\ 
+ & term & estimate & std.error & statistic & sig \\ 
   \hline
-(Intercept) & -2.942 & 0.270 & -10.896 & 0.000 \\ 
-  r\_baseline\_mgl\_features & 1.600 & 0.171 & 9.361 & 0.000 \\ 
-  r\_individual\_gcm\_features & 2.786 & 0.378 & 7.369 & 0.000 \\ 
-  participant\_prereg\_average & 5.437 & 0.585 & 9.298 & 0.000 \\ 
+1 & (Intercept) & -2.94 & 0.27 & -10.90 & *** \\ 
+  2 & r\_baseline\_mgl\_features & 1.60 & 0.17 & 9.36 & *** \\ 
+  3 & r\_individual\_gcm\_features & 2.79 & 0.38 & 7.37 & *** \\ 
+  4 & participant\_prereg\_average & 5.44 & 0.58 & 9.30 & *** \\ 
    \hline

### 2020-02-25 supplementary_information/si.Rnw
last round before Jen looks

@@
     participant_prereg_average = pre_reg_av - mean(pre_reg_av, na.rm = T), # center participant average
-    bot_lexical_typicality = dplyr::case_when( # set factor levels used in paper, order factor
+    co_player_lexical_typicality = dplyr::case_when( # set factor levels used in paper, order factor
         lex_typicality == 'typical' ~ 'typical',
@@
       factor(levels = c('typical', 'random', 'reversed')),
-    bot_regularisation_rate = reg_rate %>% # order factor
+    co_player_regularisation_rate = reg_rate %>% # order factor
       factor(levels = c('nc', '-40%', '+40%')),
@@
 
-\noindent In this supplementary information, we detail (\ref{appendixverbs}) the structure of our nonce verb stimuli, the setup of (\ref{appendixgcm}) the Generalized Context Model (GCM) and (\ref{appendixmgl}) the Minimal Generalization Learner (MGL), and (\ref{appendixgcmvsmgl}) how these models compare.
+In this supplementary information, we detail (\ref{appendixverbs}) the structure of our nonce verb stimuli, the setup of (\ref{appendixgcm}) the Generalized Context Model (GCM) and (\ref{appendixmgl}) the Minimal Generalization Learner (MGL), and (\ref{appendixgcmvsmgl}) how these models compare.
 We illustrate the models using fits on our \emph {baseline} data.
@@
 <<gcmtable, echo=F, eval=T, results=tex>>=
-print(xtable::xtable(gcm_table, digits=2, label="gcmsetsize", caption="Number of forms in each verb class, GCM training data"))
+print(xtable::xtable(gcm_table, digits=2, include.rownames = FALSE, rowlabel="gcmsetsize", caption="Number of forms in each verb class, GCM training data"), include.rownames = F)
 
@@
 
+mgl_table = mgl_table %>% 
+  mutate(
+    category = toupper(category)
+  )
+
 @
@@
 <<mgltable, echo=F, eval=T, results=tex>>=
-print(xtable::xtable(mgl_table, digits=2, label="mglsetsize", caption="Number of forms in each verb class, MGL training data"))
+print(xtable::xtable(mgl_table, digits=2, label="mglsetsize", caption="Number of forms in each verb class, MGL training data"), include.rownames = F)
 
@@
 In the MGL, the disjunction emerges despite the lack of an initial specification: while one rule accounts for the specific behavior of a given form, multiple rules cover a given class.
-The number of regular and irregular rules posited by the MGL for each verb class can be seen in Table \ref{mglruleses}.
+The number of regular and irregular rules posited by the MGL for each verb class can be seen in Table \ref{mgltable}.
 
@@
 mgl_rule_count = rbind(mgl_rule_count, c('sum', sum(mgl_rule_count$`irregular rule`), sum(mgl_rule_count$`regular rule`), sum(mgl_rule_count$sum)))
-
-@
-
-<<mgltable, echo=F, eval=T, results=tex>>=
-print(xtable::xtable(mgl_table, digits=2, label="mglruleses", caption="The number of unique rules posited by the MGL in each verb class"))
 
@@
 lme.fit1 = glmer(resp_post_reg ~ 
-    1 + verb_baseline_average * bot_lexical_typicality + participant_prereg_average + bot_regularisation_rate +
+    1 + verb_baseline_average * co_player_lexical_typicality + participant_prereg_average + co_player_regularisation_rate +
     (1 + verb_baseline_average | participant_id) + 
@@
 lme.fit1b = glmer(resp_post_reg ~ 
-    1 + verb_baseline_average + bot_lexical_typicality + participant_prereg_average + bot_regularisation_rate +
+    1 + verb_baseline_average + co_player_lexical_typicality + participant_prereg_average + co_player_regularisation_rate +
     (1 + verb_baseline_average | participant_id) + 
@@
 lme.fit1c = glmer(resp_post_reg ~ 
-    1 + verb_baseline_average * bot_lexical_typicality * bot_regularisation_rate + participant_prereg_average + 
+    1 + verb_baseline_average * co_player_lexical_typicality * co_player_regularisation_rate + participant_prereg_average + 
     (1 + verb_baseline_average | participant_id) + 
@@
 print(xtable::xtable(lme.fit1.sum, digits = 2, label = 'regressiontable1', caption = 'Experimental Data, Post-test Regression Model Summary.
-Model formula: post-test regular response ~ verb baseline mean x lexical typicality + participant pretest mean + regularization shift + (1 + verb.baseline.mean | participant) + (1 | verb)'))
+Model formula: post-test regular response ~ verb baseline mean x lexical typicality + participant pretest mean + regularization shift + (1 + verb.baseline.mean | participant) + (1 | verb)'), include.rownames = F)
 @
@@
   dplyr::filter(
-    bot_lexical_typicality != 'typical'
+    co_player_lexical_typicality != 'typical'
   )
@@
 fit_power1 = glmer(resp_post_reg ~ 
-    1 + verb_baseline_average * bot_lexical_typicality + participant_prereg_average + bot_regularisation_rate +
+    1 + verb_baseline_average * co_player_lexical_typicality + participant_prereg_average + co_player_regularisation_rate +
     (1 + verb_baseline_average | participant_id) + 
@@
 fit_power2 = glmer(resp_post_reg ~ 
-    1 + verb_baseline_average + bot_lexical_typicality + participant_prereg_average + bot_regularisation_rate +
+    1 + verb_baseline_average + co_player_lexical_typicality + participant_prereg_average + co_player_regularisation_rate +
     (1 + verb_baseline_average | participant_id) + 
@@
 celex_m3_summary = niceCoefTable(celex_m3)
-print(xtable::xtable(celex_m3_summary, digits = 2, label = 'resid1', caption = 'Model 3: MGL scores and GCM residuals, baseline data'))
+print(xtable::xtable(celex_m3_summary, digits = 2, label = 'resid1', caption = 'Model 3: MGL scores and GCM residuals, baseline data'), include.rownames = F)
 @
@@
 celex_m4_summary = niceCoefTable(celex_m4)
-print(xtable::xtable(celex_m4_summary, digits = 2, label = 'resid2', caption = 'Model 4: GCM scores and MGL residuals, baseline data'))
+print(xtable::xtable(celex_m4_summary, digits = 2, label = 'resid2', caption = 'Model 4: GCM scores and MGL residuals, baseline data'), include.rownames = F)
 @
@@
 indiv_m5_sum = niceCoefTable(indiv_m5)
-print(xtable::xtable(indiv_m5_sum, digits = 2, label = 'indiv1', caption = 'Model 5: Celex-GCM and individual score difference, post-test data'))
+print(xtable::xtable(indiv_m5_sum, digits = 2, label = 'indiv1', caption = 'Model 5: Celex-GCM and individual score difference, post-test data'), include.rownames = F)
 
@@
 indiv_m6_sum = niceCoefTable(indiv_m6)
-print(xtable::xtable(indiv_m6_sum, digits = 3, label = 'indiv2', caption = 'Model 6: Celex-MGL and individual score difference, post-test data'))
+print(xtable::xtable(indiv_m6_sum, digits = 3, label = 'indiv2', caption = 'Model 6: Celex-MGL and individual score difference, post-test data'), include.rownames = F)
 @
@@
 comp_fit1_sum = niceCoefTable(comp_fit1)
-print(xtable::xtable(comp_fit1_sum, digits = 2, label = 'comp3', caption = 'Celex-MGL and Individual-GCM predict variation in the post-test'))
+print(xtable::xtable(comp_fit1_sum, digits = 2, label = 'comp3', caption = 'Celex-MGL and Individual-GCM predict variation in the post-test'), include.rownames = F)
 @
@@
     word,
-    bot_lexical_typicality,
+    co_player_lexical_typicality,
     baseline_av # baseline average is there already
@@
     word,
-    bot_lexical_typicality,
+    co_player_lexical_typicality,
     rescaled_CELEX_gcm_feature_distance
@@
     word,
-    bot_lexical_typicality,
+    co_player_lexical_typicality,
     rescaled_CELEX_mgl_feature_distance
@@
 # add loess smooths.
-p1 = ggplot(three_studies, aes(x = baseline, y = posttest, colour = bot_lexical_typicality, shape = bot_lexical_typicality, linetype = bot_lexical_typicality)) +
+p1 = ggplot(three_studies, aes(x = baseline, y = posttest, colour = co_player_lexical_typicality, shape = co_player_lexical_typicality, linetype = co_player_lexical_typicality)) +
   guides(size = F) +

### 2020-02-25 supplementary_information/si.tex
last round before Jen looks

@@
 % latex table generated in R 3.6.0 by xtable 1.8-4 package
-% Thu Feb 20 11:18:10 2020
+% Tue Feb 25 14:32:38 2020
 \begin{table}[ht]
 \centering
-\begin{tabular}{rllll}
+\begin{tabular}{llll}
   \hline
- & verb class & irregular set & regular set & miscellaneous regular verbs \\ 
+verb class & irregular set & regular set & miscellaneous regular verbs \\ 
   \hline
-1 & BURNT & 6 & 42 & 1218 \\ 
-  2 & DROVE & 14 & 83 & 1218 \\ 
-  3 & KEPT & 12 & 31 & 1218 \\ 
-  4 & SANG & 8 & 13 & 1218 \\ 
-  5 & sum & 40 & 169 & 1218 \\ 
+BURNT & 6 & 42 & 1218 \\ 
+  DROVE & 14 & 83 & 1218 \\ 
+  KEPT & 12 & 31 & 1218 \\ 
+  SANG & 8 & 13 & 1218 \\ 
+  sum & 40 & 169 & 1218 \\ 
    \hline
@@
 \caption{Number of forms in each verb class, GCM training data} 
-\label{gcmsetsize}
 \end{table}
@@
 % latex table generated in R 3.6.0 by xtable 1.8-4 package
-% Thu Feb 20 11:18:10 2020
+% Tue Feb 25 14:32:38 2020
 \begin{table}[ht]
 \centering
-\begin{tabular}{rllrr}
+\begin{tabular}{llrr}
   \hline
- & category & rule.type & related forms & exceptions \\ 
+category & rule.type & related forms & exceptions \\ 
   \hline
-1 & burnt & irregular &  24 &  41 \\ 
-  2 & drove & irregular &  21 & 114 \\ 
-  3 & kept & irregular &  10 &   9 \\ 
-  4 & sang & irregular &  11 &  77 \\ 
-  5 & burnt & regular &  38 &  21 \\ 
-  6 & drove & regular &  29 &  27 \\ 
-  7 & kept & regular &  67 &  41 \\ 
-  8 & sang & regular &  47 &  38 \\ 
+BURNT & irregular &  24 &  41 \\ 
+  DROVE & irregular &  21 & 114 \\ 
+  KEPT & irregular &  10 &   9 \\ 
+  SANG & irregular &  11 &  77 \\ 
+  BURNT & regular &  38 &  21 \\ 
+  DROVE & regular &  29 &  27 \\ 
+  KEPT & regular &  67 &  41 \\ 
+  SANG & regular &  47 &  38 \\ 
    \hline
@@
 In the MGL, the disjunction emerges despite the lack of an initial specification: while one rule accounts for the specific behavior of a given form, multiple rules cover a given class.
-The number of regular and irregular rules posited by the MGL for each verb class can be seen in Table \ref{mglruleses}.
+The number of regular and irregular rules posited by the MGL for each verb class can be seen in Table \ref{mgltable}.
 
-% latex table generated in R 3.6.0 by xtable 1.8-4 package
-% Thu Feb 20 11:18:10 2020
-\begin{table}[ht]
-\centering
-\begin{tabular}{rllrr}
-  \hline
- & category & rule.type & related forms & exceptions \\ 
-  \hline
-1 & burnt & irregular &  24 &  41 \\ 
-  2 & drove & irregular &  21 & 114 \\ 
-  3 & kept & irregular &  10 &   9 \\ 
-  4 & sang & irregular &  11 &  77 \\ 
-  5 & burnt & regular &  38 &  21 \\ 
-  6 & drove & regular &  29 &  27 \\ 
-  7 & kept & regular &  67 &  41 \\ 
-  8 & sang & regular &  47 &  38 \\ 
-   \hline
-\end{tabular}
-\caption{The number of unique rules posited by the MGL in each verb class} 
-\label{mglruleses}
-\end{table}
 Table \ref{mglruleseskept} shows the list of (irregular and regular) rules for one of the classes, the KEPT class.
@@
 % latex table generated in R 3.6.0 by xtable 1.8-4 package
-% Thu Feb 20 11:18:10 2020
+% Tue Feb 25 14:32:38 2020
 \begin{table}[ht]
 \centering
-\begin{tabular}{rlrrrl}
+\begin{tabular}{lrrrl}
   \hline
- & term & estimate & std.error & statistic & sig \\ 
+term & estimate & std.error & statistic & sig \\ 
   \hline
-1 & (Intercept) & -0.11 & 0.16 & -0.67 & n.s. \\ 
-  2 & verb\_baseline\_average & 8.26 & 0.69 & 12.04 & *** \\ 
-  3 & bot\_lexical\_typicalityrandom & 0.04 & 0.17 & 0.25 & n.s. \\ 
-  4 & bot\_lexical\_typicalityreversed & 0.04 & 0.17 & 0.20 & n.s. \\ 
-  5 & participant\_prereg\_average & 7.13 & 0.56 & 12.67 & *** \\ 
-  6 & bot\_regularisation\_rate-40\% & -0.46 & 0.18 & -2.47 & * \\ 
-  7 & bot\_regularisation\_rate+40\% & 0.55 & 0.17 & 3.31 & *** \\ 
-  8 & verb\_baseline\_average:bot\_lexical\_typicalityrandom & -2.40 & 0.93 & -2.59 & ** \\ 
-  9 & verb\_baseline\_average:bot\_lexical\_typicalityreversed & -2.13 & 0.96 & -2.23 & * \\ 
+(Intercept) & -0.11 & 0.16 & -0.67 & n.s. \\ 
+  verb\_baseline\_average & 8.26 & 0.69 & 12.04 & *** \\ 
+  co\_player\_lexical\_typicalityrandom & 0.04 & 0.17 & 0.25 & n.s. \\ 
+  co\_player\_lexical\_typicalityreversed & 0.04 & 0.17 & 0.20 & n.s. \\ 
+  participant\_prereg\_average & 7.13 & 0.56 & 12.67 & *** \\ 
+  co\_player\_regularisation\_rate-40\% & -0.46 & 0.18 & -2.47 & * \\ 
+  co\_player\_regularisation\_rate+40\% & 0.55 & 0.17 & 3.31 & *** \\ 
+  verb\_baseline\_average:co\_player\_lexical\_typicalityrandom & -2.40 & 0.93 & -2.59 & ** \\ 
+  verb\_baseline\_average:co\_player\_lexical\_typicalityreversed & -2.13 & 0.96 & -2.23 & * \\ 
    \hline
@@
 % latex table generated in R 3.6.0 by xtable 1.8-4 package
-% Thu Feb 20 11:18:11 2020
+% Tue Feb 25 14:32:39 2020
 \begin{table}[ht]
 \centering
-\begin{tabular}{rlrrrl}
+\begin{tabular}{lrrrl}
   \hline
- & term & estimate & std.error & statistic & sig \\ 
+term & estimate & std.error & statistic & sig \\ 
   \hline
-1 & (Intercept) & -0.53 & 0.12 & -4.22 & *** \\ 
-  2 & rescaled\_CELEX\_mgl\_feature\_distance & 1.35 & 0.09 & 14.48 & *** \\ 
-  3 & rescaled\_CELEX\_gcm\_feature\_distance\_residual & 1.42 & 0.13 & 11.10 & *** \\ 
+(Intercept) & -0.53 & 0.12 & -4.22 & *** \\ 
+  rescaled\_CELEX\_mgl\_feature\_distance & 1.35 & 0.09 & 14.48 & *** \\ 
+  rescaled\_CELEX\_gcm\_feature\_distance\_residual & 1.42 & 0.13 & 11.10 & *** \\ 
    \hline
@@
 % latex table generated in R 3.6.0 by xtable 1.8-4 package
-% Thu Feb 20 11:18:12 2020
+% Tue Feb 25 14:32:40 2020
 \begin{table}[ht]
 \centering
-\begin{tabular}{rlrrrl}
+\begin{tabular}{lrrrl}
   \hline
- & term & estimate & std.error & statistic & sig \\ 
+term & estimate & std.error & statistic & sig \\ 
   \hline
-1 & (Intercept) & -1.23 & 0.16 & -7.68 & *** \\ 
-  2 & rescaled\_CELEX\_gcm\_feature\_distance & 2.39 & 0.15 & 16.06 & *** \\ 
-  3 & rescaled\_CELEX\_mgl\_feature\_distance\_residual & 1.01 & 0.09 & 11.15 & *** \\ 
+(Intercept) & -1.23 & 0.16 & -7.68 & *** \\ 
+  rescaled\_CELEX\_gcm\_feature\_distance & 2.39 & 0.15 & 16.06 & *** \\ 
+  rescaled\_CELEX\_mgl\_feature\_distance\_residual & 1.01 & 0.09 & 11.15 & *** \\ 
    \hline
@@
 % latex table generated in R 3.6.0 by xtable 1.8-4 package
-% Thu Feb 20 11:18:12 2020
+% Tue Feb 25 14:32:40 2020
 \begin{table}[ht]
 \centering
-\begin{tabular}{rlrrrl}
+\begin{tabular}{lrrrl}
   \hline
- & term & estimate & std.error & statistic & sig \\ 
+term & estimate & std.error & statistic & sig \\ 
   \hline
-1 & (Intercept) & -2.95 & 0.29 & -10.01 & *** \\ 
-  2 & rescaled\_CELEX\_gcm\_feature\_distance & 4.37 & 0.42 & 10.28 & *** \\ 
-  3 & gcm\_score\_diff & 2.61 & 0.47 & 5.56 & *** \\ 
-  4 & participant\_prereg\_average & 5.29 & 0.60 & 8.79 & *** \\ 
+(Intercept) & -2.95 & 0.29 & -10.01 & *** \\ 
+  rescaled\_CELEX\_gcm\_feature\_distance & 4.37 & 0.42 & 10.28 & *** \\ 
+  gcm\_score\_diff & 2.61 & 0.47 & 5.56 & *** \\ 
+  participant\_prereg\_average & 5.29 & 0.60 & 8.79 & *** \\ 
    \hline
@@
 % latex table generated in R 3.6.0 by xtable 1.8-4 package
-% Thu Feb 20 11:18:12 2020
+% Tue Feb 25 14:32:40 2020
 \begin{table}[ht]
 \centering
-\begin{tabular}{rlrrrl}
+\begin{tabular}{lrrrl}
   \hline
- & term & estimate & std.error & statistic & sig \\ 
+term & estimate & std.error & statistic & sig \\ 
   \hline
-1 & (Intercept) & -1.318 & 0.152 & -8.663 & *** \\ 
-  2 & rescaled\_CELEX\_mgl\_feature\_distance & 1.944 & 0.176 & 11.045 & *** \\ 
-  3 & mgl\_score\_diff & 0.304 & 0.139 & 2.181 & * \\ 
-  4 & participant\_prereg\_average & 7.203 & 0.591 & 12.190 & *** \\ 
+(Intercept) & -1.318 & 0.152 & -8.663 & *** \\ 
+  rescaled\_CELEX\_mgl\_feature\_distance & 1.944 & 0.176 & 11.045 & *** \\ 
+  mgl\_score\_diff & 0.304 & 0.139 & 2.181 & * \\ 
+  participant\_prereg\_average & 7.203 & 0.591 & 12.190 & *** \\ 
    \hline
@@
 % latex table generated in R 3.6.0 by xtable 1.8-4 package
-% Thu Feb 20 11:18:12 2020
+% Tue Feb 25 14:32:40 2020
 \begin{table}[ht]
 \centering
-\begin{tabular}{rlrrrl}
+\begin{tabular}{lrrrl}
   \hline
- & term & estimate & std.error & statistic & sig \\ 
+term & estimate & std.error & statistic & sig \\ 
   \hline
-1 & (Intercept) & -2.94 & 0.27 & -10.90 & *** \\ 
-  2 & r\_baseline\_mgl\_features & 1.60 & 0.17 & 9.36 & *** \\ 
-  3 & r\_individual\_gcm\_features & 2.79 & 0.38 & 7.37 & *** \\ 
-  4 & participant\_prereg\_average & 5.44 & 0.58 & 9.30 & *** \\ 
+(Intercept) & -2.94 & 0.27 & -10.90 & *** \\ 
+  r\_baseline\_mgl\_features & 1.60 & 0.17 & 9.36 & *** \\ 
+  r\_individual\_gcm\_features & 2.79 & 0.38 & 7.37 & *** \\ 
+  participant\_prereg\_average & 5.44 & 0.58 & 9.30 & *** \\ 
    \hline

### 2020-03-03 supplementary_information/si.Rnw
final commit? yeah right

@@
 
-\noindent This is an Sweave file.
-The Rnw source file contains the code used to compile the pdf output file.
-It also contains additional code sections that help replicate the results.
-The files also contain pointers to further detail in the supplementary depository that accompanies the paper.
-
-In this supplementary information, we detail (\ref{appendixverbs}) the structure of our nonce verb stimuli, the setup of (\ref{appendixgcm}) the Generalized Context Model (GCM) and (\ref{appendixmgl}) the Minimal Generalization Learner (MGL), and (\ref{appendixgcmvsmgl}) how these models compare.
+\noindent In this supplementary information, we detail (\ref{appendixverbs}) the structure of our nonce verb stimuli, the setup of (\ref{appendixgcm}) the Generalized Context Model (GCM) and (\ref{appendixmgl}) the Minimal Generalization Learner (MGL), and (\ref{appendixgcmvsmgl}) how these models compare.
 We illustrate the models using fits on our \emph {baseline} data.
@@
 Additional material is reproduced from the main text, revisiting specific regression model summaries and additional methods we used to investigate contributions from the GCM and MGL.
+
+\begin{quote}
+This is an Sweave file.
+The Rnw source file contains the code used to compile the pdf output file.
+It also contains additional code sections that help replicate the results.
+The file also contains pointers to further detail in the supplementary repository that accompanies the paper.
+\end{quote}
 
@@
 For instance, our DROVE class is a generalized version of Moder's class 4, described using the \textipa{[aI]}$\rightarrow{}$\textipa{[oU]} alternation but also including {\em weave}.
-(In contrast, Albright \& Hayes restrict this class to this alternation.) The verb classes could be defined in a number of ways and still be mostly consistent with the work of Moder, for instance.
-We have trialed a set of possible small changes and found that they have no major effect on our overall results.
+(In contrast, Albright \& Hayes restrict this class to the \textipa{[aI]}$\rightarrow{}$\textipa{[oU]} alternation.) The verb classes could be defined in a number of ways and still be mostly consistent with the work of Moder, for instance.
+We have trialed a set of small changes in defining the verb categories and found that they have no major effect on our overall results.
 
@@
 \begin{quote}
- \texttt{The implementation of the GCM is laid out in models/gcm/gcm.html}.
+ The implementation of the GCM is laid out in models/gcm/gcm.html.
 \end{quote}
@@
 The irregular set consists of verb types in \textsc{Celex} that form their past tense according to the pattern captured by the class (such as an \{\textipa{[aI]},\textipa{[i]}\} $\rightarrow{}$ \textipa{[oU]} alternation).
-The regular set consists of verb types that have base forms that are similar to these irregular forms but have a regular {\em -ed} past tense form as well as miscellaneous regular verbs -- those that do not belong to our schemata .
+The regular set consists of verb types that have base forms that are similar to these irregular forms but have a regular {\em -ed} past tense form as well as miscellaneous regular verbs -- those that do not belong to any of our schemata.
 We narrow the regular set to monosyllabic forms.
-However, all polysyllabic irregular forms that could serve as a point of comparison are compounds based on monosyllabic forms (an example is {\em overwrite}, a compound form of irregular {\em write}).
-(A compound form might be more regular than a simplex form, but \textsc{Celex} will list both the regular and the irregular variant in both cases.) Table \ref{regexesforgcm} shows the descriptions of the verb classes using regular expressions.
+However, all polysyllabic irregular forms that could serve as a point of comparison are compounds based on monosyllabic forms (an example is {\em overwrite}, a compound form of irregular {\em write}.
+A compound form might be more regular than a simplex form, but \textsc{Celex} will list both the regular and the irregular variant in both cases.) Table \ref{regexesforgcm} shows the descriptions of the verb classes using regular expressions.
 
@@
   gcm_table,
-  c('sum', 
+  c('count of unique forms', 
     sum(gcm_table$`irregular set`),
@@
 The miscellaneous set consists of monosyllabic verbs that do not belong to any of the schemata and are regular.
-These are included in the regular training set of \textsc{ each} verb class.
+These are included in the regular training set of each verb class.
 
@@
 <<gcmtable, echo=F, eval=T, results=tex>>=
-print(xtable::xtable(gcm_table, digits=2, include.rownames = FALSE, rowlabel="gcmsetsize", caption="Number of forms in each verb class, GCM training data"), include.rownames = F)
+print(xtable::xtable(gcm_table, digits=2, include.rownames = FALSE, label="gcmsetsize", caption="Number of forms in each verb class, GCM training data"), include.rownames = F)
 
@@
 
-\subsection{Output format}
+\subsection{Output format \label{gcm_scale}}
 
@@
 
-We standardize the regularity score to match the range of participant responses: [0,1].
+We rescale the regularity score to match the range of participant responses: [0,1].
 The modified score is interpretable as the probability that the outcome will be regular in aggregated data.
@@
 These include their Segmental Similarity Calculator, implementing the natural class based similarity metric due to \cite{frisch1997similarity}, also used in the GCM implementation.
-Due to issues with the MGL code, we had to fit the MGL separately for participants.
+Due to issues with the MGL code, we had to fit the MGL manually, using the graphical interface, for each separate participant.
+The pyautogui library in python was used to automate this, the code is available in the supplementary repository.
 
 \begin{quote}
- \texttt{The implementation of the MGL is laid out in models/mgl/mgl.html}.
+The implementation of the MGL is laid out in models/mgl/mgl.html.
 \end{quote}
@@
 
-Note that the sets of exceptions and related forms of each rule can overlap, both respectively and with each other.
-As a consequence, the MGL rules apply to fewer forms than apparent from the table: 456 (instead of 617) in total.
-
 <<mgl_form_counts, eval=T, echo=F>>=
@@
 
-# mgl_count$form1 %>% unique %>% length
-
 # mind you, not all of these verbs have both regular and irregular rules, as discussed in the text.
@@
 )
-
-# mgl_count$form1 %>% unique %>% length
 
@@
 
-@
+# we also want the TOTAL number of related forms and exceptions across rules.
+
+mgl_count_total1 = mgl_count %>% 
+  dplyr::select(related_list) %>% 
+  tidyr::unnest(related_list) %>% # we now have the related forms as one verb per line, which is tidy. but the verbs repeat, since the same verb can be a related form for several target verbs
+  dplyr::distinct() %>% # so we distinct them!
+  dplyr::mutate(
+    type = 'related forms' # by the way, this is related forms
+  ) %>% 
+  dplyr::rename(
+   forms = related_list 
+  )
+
+mgl_count_total2 = mgl_count %>% 
+  dplyr::select(exception_list) %>% 
+  tidyr::unnest(exception_list) %>% # we now have the related forms as one verb per line, which is tidy. but the verbs repeat, since the same verb can be a related form for several target verbs
+  dplyr::distinct() %>% # so we distinct them!
+  dplyr::mutate(
+    type = 'exceptions' # by the way, this is related forms
+  ) %>% 
+  dplyr::rename(
+   forms = exception_list 
+  )
+
+all_forms_n_mgl = mgl_count_total2 %>% 
+  dplyr::bind_rows(mgl_count_total1) %>% 
+  dplyr::distinct(forms) %>% 
+  dplyr::summarise(n =n()) %>% 
+  dplyr::pull(n)
+
+@
+
+Note that the sets of exceptions and related forms of each rule can overlap, both respectively and with each other.
+As a consequence, the MGL rules apply to fewer forms than apparent from the table: \Sexpr{all_forms_n_mgl} in total.
 
@@
 Therefore, we keep the cutoff criterion and assume that the adjusted confidence of the irregular rule for these 28 verbs is zero.
-We then take the relative confidence of the regular rule as compared to the regular {\em and} the irregular rule for each verb and took this as the adjusted regular confidence of the given verb.
+We then take the relative confidence of the regular rule as compared to the regular {\em and} the irregular rule for each verb and take this as the adjusted regular confidence of the given verb.
 (If the irregular rule is missing, the value of this adjusted regular confidence is 1.) This is given by the following equation:
@@
 The {\em relative (adjusted) confidence} of the predicted regularity of {\em splive} is $0.98/(0.98+0.37)=0.73$.
-(The respective rescaled value is 0.363 -- which, in itself, is not very informative.)
+(As with the GCM predictions -- see Section \ref{gcm_scale} of this SI -- we rescale the MGL predictions as well for statistical analysis.
+The respective rescaled value is 0.363.)
 
@@
 In contrast, the MGL starts establishing rules across all forms in the starting dictionary.
-
-As we see, however, the MGL also focuses the training set.
-The structural descriptions of the forms only cover a fraction of all verbs in the starting dictionary, organized by shared groups of segments.
-While disjunction increases the power of a theory (increasing the set of acceptable classes), disjunct classes frequently serve as input for phonological processes \citep{mielke2008emergence} .
-In the MGL, the disjunction emerges despite the lack of an initial specification: while one rule accounts for the specific behavior of a given form, multiple rules cover a given class.
-The number of regular and irregular rules posited by the MGL for each verb class can be seen in Table \ref{mgltable}.
 
@@
 
-# glue on column sums
-mgl_rule_count = rbind(mgl_rule_count, c('sum', sum(mgl_rule_count$`irregular rule`), sum(mgl_rule_count$`regular rule`), sum(mgl_rule_count$sum)))
+# we get a total rule count as well. since rules overlap, this isn't a sum of the rule counts above
+
+total_rules = mgl_baseline_rules_raw %>% # take raw rules
+  dplyr::filter( # throw out non-category verbs and non reg/irreg rules (in our definition)
+    !is.na(category),
+    !is.na(rule.type)
+  ) %>% 
+  dplyr::mutate( # make a rule column of the form A -> B / C
+    rule = paste0(A, ' ->', B, ' / ', Pfeat)
+  ) %>% 
+  dplyr::distinct(rule,rule.type) %>% # keep DISTINCT rules
+  dplyr::count(rule.type)
+
+irreg_total_count = total_rules[total_rules$rule.type == 'irregular',]$n
+reg_total_count = total_rules[total_rules$rule.type == 'regular',]$n
+
+@
+
+As we see, however, the MGL also focuses the training set.
+The structural descriptions of the forms only cover a fraction of all verbs in the starting dictionary, organized by shared groups of segments.
+While disjunction increases the power of a theory (increasing the set of acceptable classes), disjunct classes frequently serve as input for phonological processes \citep{mielke2008emergence}.
+In the MGL, the disjunction emerges despite the lack of an initial specification: while one rule accounts for the specific behavior of a given form, multiple rules cover a given class.
+The number of regular and irregular rules posited by the MGL for each verb class can be seen in Table \ref{mglruletable}.
+Note that the regular rules themselves overlap: A regular rule that accounts for \emph{weep-weeped} can also account for \emph{burn-burned}.
+In total, there are \Sexpr{irreg_total_count} irregular rules and \Sexpr{reg_total_count} regular rules posited by the MGL on the {\sc Celex} verbs.
+
+<<mglruletable, echo=F, eval=T, results=tex>>=
+print(xtable::xtable(mgl_rule_count, digits=2, label="mglruletable", caption="Number of regular and irregular rules in each verb class, posited by the MGL for the training data (Note that the regular rules themselves overlap)"), include.rownames = F)
 
@@
 
+# here are the effect estimate figures from the paper:
+# plot(effects::effect('co_player_regularisation_rate', lme.fit1), xlab = 'co-player regularization shift', ylab = 'player post-test regularization', main = 'Effect of co-player rate of regularization on pos-test responses')
+
+# plot(effects::effect('verb_baseline_average:co_player_lexical_typicality', lme.fit1), multiline = T, lines=list(lty=c(1,2,3)), xlab = 'verb baseline average', ylab = 'player post-test regularization', main = 'Effect of co-player lexical typicality on pos-test responses')
+
 @
@@
 Non-significant predictors are removed from our models, as determined by a series of comparisons of nested models, using likelihood ratio tests in R's anova function.
-However, our model tables (e.g., Tables 5 and 6 in the main text) include the significance levels generated by lme4 summaries.
-
+However, our model tables (e.g., Tables 5 and 6 in the main text) include the Wald significance levels generated by lme4 summaries (* p$<$0.05; ** p$<$0.01; *** p$<$0.001).
 We investigate maximal random effects structures, allowing for random intercepts for participants and stimulus items (verbs), as well as random slopes for all within-unit factors.
@@
 
-In the model in Table 5 in the main text, repeated here as Table \ref{regressiontable1}, we elect to mean center continuous variables to counteract collinearity that becomes evident when including the interaction between and {\em lexical typicality}\footnote{In the SI, all p values rounded to 0.00 should be taken to indicate p $<$ 0.001.}.
+In the model in Table 5 in the main text, repeated here as Table \ref{regressiontable1}, we elect to mean center continuous variables to counteract collinearity that becomes evident when including the interaction between and {\em lexical typicality}.
 Although some debate exists regarding the practice of transforming variables \citep{belsley1991conditioning,echambadi2007mean}, centering variables is common practice to remove nonessential ill conditioning, that is, relationships that inevitably exist between main and product terms \citep{aiken1991multiple,jaccard2003interaction, jaeger2008categorical}.
@@
 We fit the \textsc{Celex-GCM} and the \textsc{Celex-MGL} on the baseline data to see how well they predict regularization responses by the baseline participants.
-The \textsc{Celex-MGL} outperforms the \textsc{Celex-GCM}, if slightly.
+The \textsc{Celex-MGL} outperforms the \textsc{Celex-GCM}, albeit only slightly.
 Both models contribute independently to explaining variation in the data.
@@
 
-The concordance indices of the \textsc{Celex}-MGL (\Sexpr{C_CELEX_mgl_features}) and the \textsc{Celex}-MGL (\Sexpr{C_CELEX_gcm_features}) on the baseline data are similar, with some edge retained by the MGL, as in the case of \citep{albright2003rules}.
+The concordance indices of the \textsc{Celex}-MGL (\Sexpr{C_CELEX_mgl_features}) and the \textsc{Celex}-MGL (\Sexpr{C_CELEX_gcm_features}) on the baseline data are similar, with some edge retained by the MGL, as in the case of \cite{albright2003rules}.
 Our concordance indices are calculated using the somers2 function in R's \texttt{Hmisc} package \citep{harrell2014hmisc}.
@@
 
-We fit the \textsc{Celex-GCM},the \textsc{Celex-MGL}, as well as the Individual-GCM and the Individual-MGL on the post-test data.
+We fit the \textsc{Celex-GCM}, the \textsc{Celex-MGL}, as well as the Individual-GCM and the Individual-MGL on the post-test data.
 Recall that the \textsc{Celex}-based models were trained on verbs in \textsc{Celex}.
@@
 indiv_m6_sum = niceCoefTable(indiv_m6)
-print(xtable::xtable(indiv_m6_sum, digits = 3, label = 'indiv2', caption = 'Model 6: Celex-MGL and individual score difference, post-test data'), include.rownames = F)
+print(xtable::xtable(indiv_m6_sum, digits = 2, label = 'indiv2', caption = 'Model 6: Celex-MGL and individual score difference, post-test data'), include.rownames = F)
 @
@@
 
+p3 = anova(comp_fit1, comp_fit1d)[2,8] %>% round(2) # with individual mgl
+p4 = anova(comp_fit1, comp_fit1e)[2,8] %>% round(2) # with celex gcm
+
 @
@@
 
-Goodness-of-fit tests show that model fit is significantly worse if we exclude {\sc Celex}-MGL ($\chi^2 = $ \Sexpr{chisq1}) or individual-GCM ($\chi^2 = $ \Sexpr{chisq2}) as predictors.
-It is not significantly better if we include individual-MGL ($\chi^2 = $ \Sexpr{chisq3}) or {\sc Celex}-GCM ($\chi^2 = $ \Sexpr{chisq4}).
+Goodness-of-fit tests show that model fit is significantly worse if we exclude {\sc Celex}-MGL ($\chi^2 = $ \Sexpr{chisq1} ***) or individual-GCM ($\chi^2 = $ \Sexpr{chisq2} ***) as predictors.
+It is not significantly better if we include individual-MGL ($\chi^2 = $ \Sexpr{chisq3}, p = \Sexpr{p3}) or {\sc Celex}-GCM ($\chi^2 = $ \Sexpr{chisq4}, p = \Sexpr{p4}).
 
@@
   geom_smooth(se = F, method = 'loess', formula = 'y ~ x') +
-  facet_wrap( ~ type) +
+  facet_wrap( ~ type, ncol = 2) +
   scale_shape_discrete(name="Lexical\ntypicality") +
@@
   guides(linetype = F) +
+  theme_bw() +
   scale_colour_brewer(palette = 'Set2', name="Lexical\ntypicality")
 
+ggsave('si_figure.pdf', plot = p1, width = 9, height = 7)
+
 @
@@
 \begin{figure}
-<<model_figure2, fig=T,echo=F,eval=T, width = 12.5, height = 4, fig.margin=T, fig.align = ‘center’>>=
-p1
-@
+\centering
+\includegraphics[width = \textwidth]{si_figure.pdf}
 \caption{Baseline and individual regularization weights: participants, GCM, MGL}
@@
 
-While the loess smooths are likely slightly overfit, they show the general trend in the raw data, which is echoed in the regression analysis in the main text: the `typical' trend of baseline-post-test difference is stronger than either the "random" or "reversed" trends.
+While the loess smooths are likely slightly overfit, they show the general trend in the raw data, which is echoed in the regression analysis in the main text: the `typical' trend of baseline-post-test difference is stronger than either the `random' or `reversed' trends.
 These are not very much different from each other.
@@
 Since rules affect groups of verbs, rather than individual verbs, the post-test shifts are abrupt.
-The "typical" trend continues climbing after a longer plateau.
-However, the `random' and `reversed' trends trail off, where the "reversed" trend turns negative towards the higher tail of the distribution.
+The `typical' trend continues climbing after a longer plateau.
+However, the `random' and `reversed' trends trail off, where the `reversed' trend turns negative towards the higher tail of the distribution.
 

### 2020-03-03 supplementary_information/si.tex
final commit? yeah right

@@
 % latex table generated in R 3.6.0 by xtable 1.8-4 package
-% Tue Feb 25 14:32:38 2020
+% Tue Mar  3 13:14:57 2020
 \begin{table}[ht]
@@
   SANG & 8 & 13 & 1218 \\ 
-  sum & 40 & 169 & 1218 \\ 
+  count of unique forms & 40 & 169 & 1218 \\ 
    \hline
@@
 \caption{Number of forms in each verb class, GCM training data} 
+\label{gcmsetsize}
 \end{table}
@@
 Note that the sets of exceptions and related forms of each rule can overlap, both respectively and with each other.
-As a consequence, the MGL rules apply to fewer forms than apparent from the table: 456 (instead of 617) in total.
+As a consequence, the MGL rules apply to fewer forms than apparent from the table: 401 in total.
 
 % latex table generated in R 3.6.0 by xtable 1.8-4 package
-% Tue Feb 25 14:32:38 2020
+% Tue Mar  3 13:14:57 2020
 \begin{table}[ht]
@@
 The structural descriptions of the forms only cover a fraction of all verbs in the starting dictionary, organized by shared groups of segments.
-While disjunction increases the power of a theory (increasing the set of acceptable classes), disjunct classes frequently serve as input for phonological processes \citep{mielke2008emergence} .
+While disjunction increases the power of a theory (increasing the set of acceptable classes), disjunct classes frequently serve as input for phonological processes \citep{mielke2008emergence}.
 In the MGL, the disjunction emerges despite the lack of an initial specification: while one rule accounts for the specific behavior of a given form, multiple rules cover a given class.
-The number of regular and irregular rules posited by the MGL for each verb class can be seen in Table \ref{mgltable}.
+The number of regular and irregular rules posited by the MGL for each verb class can be seen in Table \ref{mglruletable}.
+Note that the regular rules themselves overlap: A regular rule that accounts for \emph{weep-weeped} can also account for \emph{burn-burned}.
+In total, there are 32 irregular rules and 12 regular rules posited by the MGL on the {\sc Celex} verbs.
+
+% latex table generated in R 3.6.0 by xtable 1.8-4 package
+% Tue Mar  3 13:14:57 2020
+\begin{table}[ht]
+\centering
+\begin{tabular}{lrrr}
+  \hline
+category & irregular rule & regular rule & sum \\ 
+  \hline
+burnt &   4 &   2 &   6 \\ 
+  drove &  13 &   3 &  16 \\ 
+  kept &   5 &   5 &  10 \\ 
+  sang &  10 &   4 &  14 \\ 
+   \hline
+\end{tabular}
+\caption{Number of regular and irregular rules in each verb class, posited by the MGL for the training data (Note that the regular rules themselves overlap)} 
+\label{mglruletable}
+\end{table}
 
@@
 We fit the \textsc{Celex-GCM} and the \textsc{Celex-MGL} on the baseline data to see how well they predict regularization responses by the baseline participants.
-The \textsc{Celex-MGL} outperforms the \textsc{Celex-GCM}, if slightly.
+The \textsc{Celex-MGL} outperforms the \textsc{Celex-GCM}, albeit only slightly.
 Both models contribute independently to explaining variation in the data.
 
-The concordance indices of the \textsc{Celex}-MGL (0.586) and the \textsc{Celex}-MGL (0.581) on the baseline data are similar, with some edge retained by the MGL, as in the case of \citep{albright2003rules}.
+The concordance indices of the \textsc{Celex}-MGL (0.586) and the \textsc{Celex}-MGL (0.581) on the baseline data are similar, with some edge retained by the MGL, as in the case of \cite{albright2003rules}.
 Our concordance indices are calculated using the somers2 function in R's \texttt{Hmisc} package \citep{harrell2014hmisc}.
@@
 % latex table generated in R 3.6.0 by xtable 1.8-4 package
-% Tue Feb 25 14:32:39 2020
+% Tue Mar  3 13:14:58 2020
 \begin{table}[ht]
@@
 % latex table generated in R 3.6.0 by xtable 1.8-4 package
-% Tue Feb 25 14:32:40 2020
+% Tue Mar  3 13:14:58 2020
 \begin{table}[ht]
@@
   \hline
-(Intercept) & -1.318 & 0.152 & -8.663 & *** \\ 
-  rescaled\_CELEX\_mgl\_feature\_distance & 1.944 & 0.176 & 11.045 & *** \\ 
-  mgl\_score\_diff & 0.304 & 0.139 & 2.181 & * \\ 
-  participant\_prereg\_average & 7.203 & 0.591 & 12.190 & *** \\ 
+(Intercept) & -1.32 & 0.15 & -8.66 & *** \\ 
+  rescaled\_CELEX\_mgl\_feature\_distance & 1.94 & 0.18 & 11.04 & *** \\ 
+  mgl\_score\_diff & 0.30 & 0.14 & 2.18 & * \\ 
+  participant\_prereg\_average & 7.20 & 0.59 & 12.19 & *** \\ 
    \hline
@@
 % latex table generated in R 3.6.0 by xtable 1.8-4 package
-% Tue Feb 25 14:32:40 2020
+% Tue Mar  3 13:14:59 2020
 \begin{table}[ht]
@@
 (Intercept) & -2.94 & 0.27 & -10.90 & *** \\ 
-  r\_baseline\_mgl\_features & 1.60 & 0.17 & 9.36 & *** \\ 
-  r\_individual\_gcm\_features & 2.79 & 0.38 & 7.37 & *** \\ 
+  rescaled\_CELEX\_mgl\_feature\_distance & 1.60 & 0.17 & 9.36 & *** \\ 
+  rescaled\_individual\_gcm\_feature\_distance & 2.79 & 0.38 & 7.37 & *** \\ 
   participant\_prereg\_average & 5.44 & 0.58 & 9.30 & *** \\ 
@@
 \end{table}
-Goodness-of-fit tests show that model fit is significantly worse if we exclude {\sc Celex}-MGL ($\chi^2 = $ 201.86) or individual-GCM ($\chi^2 = $ 81.82) as predictors.
-It is not significantly better if we include individual-MGL ($\chi^2 = $ 0.09) or {\sc Celex}-GCM ($\chi^2 = $ 2.03).
+Goodness-of-fit tests show that model fit is significantly worse if we exclude {\sc Celex}-MGL ($\chi^2 = $ 201.86 ***) or individual-GCM ($\chi^2 = $ 81.82 ***) as predictors.
+It is not significantly better if we include individual-MGL ($\chi^2 = $ 0.09, p = 0.77) or {\sc Celex}-GCM ($\chi^2 = $ 2.03, p = 0.15).
 
@@
 \begin{figure}
-\includegraphics{si-model_figure2}
+\centering
+\includegraphics[width = \textwidth]{si_figure.pdf}
 \caption{Baseline and individual regularization weights: participants, GCM, MGL}

### 2020-03-26 supplementary_information/si.Rnw
fix fig

@@
 library(broom) # for tidy summaries
+library(ggthemes) # for colourblind scale
 
@@
   dplyr::mutate(
-    type = factor(type, levels = c('Participant responses', 'GCM', 'MGL'))
+    type = factor(type, levels = c('Participant responses', 'GCM', 'MGL')),
+    linetype = case_when(
+      co_player_lexical_typicality == 'typical' ~ 'solid',
+      co_player_lexical_typicality == 'random' ~ 'dotdash',
+      co_player_lexical_typicality == 'reversed' ~ 'twodash'
+    )
   )
@@
 # add loess smooths.
-p1 = ggplot(three_studies, aes(x = baseline, y = posttest, colour = co_player_lexical_typicality, shape = co_player_lexical_typicality, linetype = co_player_lexical_typicality)) +
+p1 = ggplot(three_studies, aes(x = baseline, y = posttest, colour = co_player_lexical_typicality, shape = co_player_lexical_typicality, linetype = linetype)) +
   guides(size = F) +
-  geom_point() +
+  geom_point(alpha = .25) +
   geom_smooth(se = F, method = 'loess', formula = 'y ~ x') +
   facet_wrap( ~ type, ncol = 2) +
-  scale_shape_discrete(name="Lexical\ntypicality") +
+  scale_shape_discrete(name="Lexical\ntypicality", solid = FALSE) +
   # scale_colour_discrete(name="Lexical\ntypicality") +
@@
   theme_bw() +
-  scale_colour_brewer(palette = 'Set2', name="Lexical\ntypicality")
+  scale_colour_colorblind() +
+  scale_linetype_identity()
 

### 2020-04-20 supplementary_information/si.Rnw
Janet checked

@@
 
+# takes lmer fit, spits out coef table with ast
 niceCoefTable = function(fit){
@@
 # VIF for GLMM-s, from Florian Jaeger
-
+# takes lmer fit, spits out vif for coefs
 vif.mer <- function (fit) {
@@
 
-\noindent In this supplementary information, we detail (\ref{appendixverbs}) the structure of our nonce verb stimuli, the setup of (\ref{appendixgcm}) the Generalized Context Model (GCM) and (\ref{appendixmgl}) the Minimal Generalization Learner (MGL), and (\ref{appendixgcmvsmgl}) how these models compare.
+\noindent In this supplementary information, we detail (\ref{appendixverbs}) the structure of our nonce verb stimuli, (\ref{appendixgcm}) the setup of the Generalized Context Model (GCM) and (\ref{appendixmgl}) the Minimal Generalization Learner (MGL), and (\ref{appendixgcmvsmgl}) how these models compare.
 We illustrate the models using fits on our \emph {baseline} data.
@@
 
-\begin{quote}
 This is an Sweave file.
@@
 The file also contains pointers to further detail in the supplementary repository that accompanies the paper.
-\end{quote}
 
@@
 However, all polysyllabic irregular forms that could serve as a point of comparison are compounds based on monosyllabic forms (an example is {\em overwrite}, a compound form of irregular {\em write}.
-A compound form might be more regular than a simplex form, but \textsc{Celex} will list both the regular and the irregular variant in both cases.) Table \ref{regexesforgcm} shows the descriptions of the verb classes using regular expressions.
+A compound form might be more regular than a simplex form, but \textsc{Celex} will list both the regular and the irregular variant in both cases.
+Table \ref{regexesforgcm} shows the descriptions of the verb classes using regular expressions.
 
@@
 
-All parts of the word are weighted equally, because although there is evidence that past tense formation in English is predominantly driven by overlaps in word endings, onsets also play a role.
+All parts of the word are weighted equally, because despite evidence that past tense formation in English is predominantly driven by overlaps in word endings, onsets also play a role.
 (cf.
-the predominance of \emph {s}(+stop) onsets in irregular verbs forming the past tense with a vowel change, e.g.\ {\em stink}, {\em sink}, etc.
+the predominance of \emph {s}(-cont) onsets in irregular verbs forming the past tense with a vowel change, e.g.\ {\em stink}, {\em sink}, etc.
 -- see \citealt{bybee1983morphological}).
@@
 
-The overall score used in our analyses is the {\em regularity score}, which is the complement to the {\em irregularity score} and reaches a maximum of 1.0 when the output is most likely to be regular.
+The overall score used in our analyses is the {\em regularity score}, which is the complement to the {\em irregularity score} and reaches a maximum of 1.0 when the output is certain to be regular.
 Unlike \cite{dawdy2014learnability}, there is no decision rule on top of the scoring, such that any form that is more likely than not to be regular is predicted to surface as regular all the time.
@@
 The MGL generates multiple possible past tense forms for each nonce verb.
-We only consider those rules {\em relevant} that generate the past tense forms that appear in the experiment (e.g.\ {\em splive} : {\em splived} / {\em splove}).
+We only consider to be {\em relevant} those rules that generate the past tense forms that appear in the experiment (e.g.\ {\em splive} : {\em splived} / {\em splove}).
 There is at most one relevant regular rule and one relevant irregular rule for one verb, but multiple rules can generate the (ir)regular forms for each verb class.
@@
 
-Note that the sets of exceptions and related forms of each rule can overlap, both respectively and with each other.
-As a consequence, the MGL rules apply to fewer forms than apparent from the table: \Sexpr{all_forms_n_mgl} in total.
+Note that the sets of exceptions and related forms of each rule can overlap.  As a consequence, the MGL rules apply to fewer forms than apparent from the table: \Sexpr{all_forms_n_mgl} in total.
 
@@
 It does so by aligning the wordforms and analyzing shared phonetic features.
-For example,  merging the word-specific rules for {\em ring/rang} and for {\em stink/stank} yields a more general rule that expresses the information that they share: \textipa{[I]} $\rightarrow{}$\textipa{[\ae]} / [+coronal, - cont] \underline{\hspace{0.5cm}} \textipa{[N]}.
+For example,  merging the word-specific rules for {\em ring/rang} and for {\em stink/stank} yields a more general rule that expresses the information that they share: \textipa{[I]} $\rightarrow{}$\textipa{[\ae]} / [+coronal] \underline{\hspace{0.5cm}} \textipa{[N]}.
 Each rule inferred in this way is then further generalized on the basis of more comparisons; for instance, taking note of {\em swim/swam} expands the \textipa{[I]} $\rightarrow{}$\textipa{[\ae]} rule to specify that it occurs before all [+nasal] consonants.
@@
 Therefore, we keep the cutoff criterion and assume that the adjusted confidence of the irregular rule for these 28 verbs is zero.
-We then take the relative confidence of the regular rule as compared to the regular {\em and} the irregular rule for each verb and take this as the adjusted regular confidence of the given verb.
+
+For all forms, we then take the relative confidence of the regular rule as compared to the regular {\em and} the irregular rule for each verb and take this as the adjusted regular confidence of the given verb.
 (If the irregular rule is missing, the value of this adjusted regular confidence is 1.) This is given by the following equation:
@@
 The {\em relevant irregular rule} that generates the irregular form is `\textipa{[aI]} $\rightarrow{}$ \textipa{[oU]} / \{\textipa{D}, \textipa{Z}, \textipa{dZ}, d, l, n, r, z\} \underline{\hspace{0.5cm}}v'.
-It applies to \textipa{[aI]} in the nucleus {\em preceded by voiced anterior consonant and followed by \textipa{[v]}}.
+It applies to \textipa{[aI]} in the nucleus {\em preceded by a voiced anterior consonant and followed by \textipa{[v]}}.
 In \textsc{Celex},the rule applies to three forms ({\em drive, strive, dive}) and fails to apply to five ({\em arrive, thrive, contrive, rive, connive}).
@@
 The {\em relative (adjusted) confidence} of the predicted regularity of {\em splive} is $0.98/(0.98+0.37)=0.73$.
-(As with the GCM predictions -- see Section \ref{gcm_scale} of this SI -- we rescale the MGL predictions as well for statistical analysis.
+(As with the GCM predictions -- see Section \ref{gcm_scale} of this SI -- we rescale the MGL predictions for statistical analysis.
 The respective rescaled value is 0.363.)
@@
 
-The training set for the GCM is focused initially; it is grouped into four verb classes and based on formal similarity with base forms of existing irregular verbs playing a role within each class.
+The training set for the GCM is focused from the start; it is grouped into four verb classes and based on formal similarity with base forms of existing irregular verbs playing a role within each class.
 This, in effect, assumes lexical gangs as a starting point \citep{alegre1999rule}.
@@
 
-<<main_model3, echo=F,eval=T, results=tex>>=
-load('models/si_fit1.rda')
-lme.fit1.sum = niceCoefTable(lme.fit1)
-
-print(xtable::xtable(lme.fit1.sum, digits = 2, label = 'regressiontable1', caption = 'Experimental Data, Post-test Regression Model Summary.
-Model formula: post-test regular response ~ verb baseline mean x lexical typicality + participant pretest mean + regularization shift + (1 + verb.baseline.mean | participant) + (1 | verb)'), include.rownames = F)
-@
-
 Our regression model selection in this article proceeded as follows.
@@
 
-In the model in Table 5 in the main text, repeated here as Table \ref{regressiontable1}, we elect to mean center continuous variables to counteract collinearity that becomes evident when including the interaction between and {\em lexical typicality}.
+<<main_model3, echo=F,eval=T, results=tex>>=
+load('models/si_fit1.rda')
+lme.fit1.sum = niceCoefTable(lme.fit1)
+
+print(xtable::xtable(lme.fit1.sum, digits = 2, label = 'regressiontable1', caption = 'Experimental Data, Post-test Regression Model Summary.
+Model formula: post-test regular response ~ verb baseline mean x lexical typicality + participant pretest mean + regularization shift + (1 + verb.baseline.mean | participant) + (1 | verb)'), include.rownames = F)
+@
+
+In the model in Table 5 in the main text, repeated here as Table \ref{regressiontable1}, we elect to mean center continuous variables to counteract collinearity that becomes evident when including the interaction between {\em verb baseline average} and {\em lexical typicality}.
 Although some debate exists regarding the practice of transforming variables \citep{belsley1991conditioning,echambadi2007mean}, centering variables is common practice to remove nonessential ill conditioning, that is, relationships that inevitably exist between main and product terms \citep{aiken1991multiple,jaccard2003interaction, jaeger2008categorical}.
@@
 
-Since it is unclear whether the absence of a difference between random and reversed conditions is due to a lack of data, we refit the model with the baseline average : typicality interaction on the random and reversed conditions only.
+Since it is unclear whether the absence of a difference between random and reversed conditions is due to a lack of data, we refit the model with the {\em verb baseline average : lexical typicality} interaction on the random and reversed conditions only.
 Then we calculated the Bayes Factor of the null model (with no interaction) over the full model (with interaction) using Bayesian Information Criteria calculated by lme4.
@@
 The summary can be seen in Table \ref{comp3}.
+This is the model we report in detail in the main text.
 
@@
 
-While the loess smooths are likely slightly overfit, they show the general trend in the raw data, which is echoed in the regression analysis in the main text: the `typical' trend of baseline-post-test difference is stronger than either the `random' or `reversed' trends.
+The loess smooths show the general trend in the raw data, which is echoed in the regression analysis in the main text: the `typical' trend of baseline-post-test difference is stronger than either the `random' or `reversed' trends.
 These are not very much different from each other.
-The differences are gradual.
+The differences in participant behaviour across conditions are gradual.
 The GCM, while generally overestimating regularity, captures this gradual shift, since verb behavior is shifted individually in the post-test.
 The MGL also overestimates regularization.
-Since rules affect groups of verbs, rather than individual verbs, the post-test shifts are abrupt.
+Since rules affect groups of verbs, rather than individual verbs, the MGL predicts more clustered variation across conditions.
 The `typical' trend continues climbing after a longer plateau.

### 2020-04-20 supplementary_information/si.tex
fix fig
Janet checked

@@
 % latex table generated in R 3.6.0 by xtable 1.8-4 package
-% Tue Mar  3 13:14:57 2020
+% Mon Apr 20 20:04:26 2020
 \begin{table}[ht]
@@
 
-Note that the sets of exceptions and related forms of each rule can overlap, both respectively and with each other.
-As a consequence, the MGL rules apply to fewer forms than apparent from the table: 401 in total.
+Note that the sets of exceptions and related forms of each rule can overlap.  As a consequence, the MGL rules apply to fewer forms than apparent from the table: 401 in total.
 
 % latex table generated in R 3.6.0 by xtable 1.8-4 package
-% Tue Mar  3 13:14:57 2020
+% Mon Apr 20 20:04:26 2020
 \begin{table}[ht]
@@
 
+Our regression model selection in this article proceeded as follows.
+Except where stated otherwise, we fit logistic mixed-effects regression models, using R's lme4 package \citep{bateslme4}.
+Non-significant predictors are removed from our models, as determined by a series of comparisons of nested models, using likelihood ratio tests in R's anova function.
+However, our model tables (e.g., Tables 5 and 6 in the main text) include the Wald significance levels generated by lme4 summaries (* p$<$0.05; ** p$<$0.01; *** p$<$0.001).
+We investigate maximal random effects structures, allowing for random intercepts for participants and stimulus items (verbs), as well as random slopes for all within-unit factors.
+We compare nested random effects structures as above, and retain any random slopes and intercepts which are supported by likelihood ratio tests \citep{baayen2008mixed}.
+
 % latex table generated in R 3.6.0 by xtable 1.8-4 package
-% Tue Mar  3 13:14:57 2020
+% Mon Apr 20 20:04:26 2020
 \begin{table}[ht]
@@
 \end{table}
-Our regression model selection in this article proceeded as follows.
-Except where stated otherwise, we fit logistic mixed-effects regression models, using R's lme4 package \citep{bateslme4}.
-Non-significant predictors are removed from our models, as determined by a series of comparisons of nested models, using likelihood ratio tests in R's anova function.
-However, our model tables (e.g., Tables 5 and 6 in the main text) include the Wald significance levels generated by lme4 summaries (* p$<$0.05; ** p$<$0.01; *** p$<$0.001).
-We investigate maximal random effects structures, allowing for random intercepts for participants and stimulus items (verbs), as well as random slopes for all within-unit factors.
-We compare nested random effects structures as above, and retain any random slopes and intercepts which are supported by likelihood ratio tests \citep{baayen2008mixed}.
-
-In the model in Table 5 in the main text, repeated here as Table \ref{regressiontable1}, we elect to mean center continuous variables to counteract collinearity that becomes evident when including the interaction between and {\em lexical typicality}.
+In the model in Table 5 in the main text, repeated here as Table \ref{regressiontable1}, we elect to mean center continuous variables to counteract collinearity that becomes evident when including the interaction between {\em verb baseline average} and {\em lexical typicality}.
 Although some debate exists regarding the practice of transforming variables \citep{belsley1991conditioning,echambadi2007mean}, centering variables is common practice to remove nonessential ill conditioning, that is, relationships that inevitably exist between main and product terms \citep{aiken1991multiple,jaccard2003interaction, jaeger2008categorical}.
@@
 % latex table generated in R 3.6.0 by xtable 1.8-4 package
-% Tue Mar  3 13:14:58 2020
+% Mon Apr 20 20:04:27 2020
 \begin{table}[ht]
@@
 The summary can be seen in Table \ref{comp3}.
+This is the model we report in detail in the main text.
 
 % latex table generated in R 3.6.0 by xtable 1.8-4 package
-% Tue Mar  3 13:14:59 2020
+% Mon Apr 20 20:04:28 2020
 \begin{table}[ht]

### 2020-04-24 supplementary_information/si.Rnw
na

@@
     ) %>%
+    rename(`Z score` = statistic) %>% 
     select(
@@
 This is achieved using the string-edit (Levenshtein) distance, which is the smallest number of changes needed to transform one word into the other.
-Costs range from 0 (the corresponding segments are identical) to 1 (inserting or deleting an entire segment).
+For one unit of edit distance, these costs range from 0 (the corresponding segments are identical) to 1 (inserting or deleting an entire segment).
 Following \cite{albright2003rules} and \cite{dawdy2014learnability}, costs between 0 and 1 are assigned to corresponding segments that are not identical, based on how much the segments differ.

### 2020-10-15 supplementary_information/si.Rnw
proof check

@@
 library(broom) # for tidy summaries
+library(broom.mixed) # for tidy summaries, lmer
 library(ggthemes) # for colourblind scale
@@
     tidy %>% 
-    filter(
-      group == 'fixed'
-    ) %>% 
-    mutate(
+    dplyr::filter(
+      effect == 'fixed'
+    ) %>% 
+    dplyr::mutate(
       sig = case_when(
@@
     ) %>%
-    rename(`Z score` = statistic) %>% 
+    dplyr::rename(`Z score` = statistic) %>% 
     select(
+      - effect,
       - p.value,
@@
 
-This is an Sweave file.
+The source file for this SI, along with the data and the code for the paper, are hosted at https://github.com/petyaracz/RaczHayPierrehumbert2019.
+% This is an Sweave file.
 The Rnw source file contains the code used to compile the pdf output file.
@@
 # here are the effect estimate figures from the paper:
-# plot(effects::effect('co_player_regularisation_rate', lme.fit1), xlab = 'co-player regularization shift', ylab = 'player post-test regularization', main = 'Effect of co-player rate of regularization on pos-test responses')
-
-# plot(effects::effect('verb_baseline_average:co_player_lexical_typicality', lme.fit1), multiline = T, lines=list(lty=c(1,2,3)), xlab = 'verb baseline average', ylab = 'player post-test regularization', main = 'Effect of co-player lexical typicality on pos-test responses')
+plot(effects::effect('co_player_regularisation_rate', lme.fit1), xlab = 'co-player regularization shift', ylab = 'player post-test regularization', main = 'Effect of co-player rate of regularization on post-test responses', lwd = 2)
+
+plot(effects::effect('verb_baseline_average:co_player_lexical_typicality', lme.fit1), multiline = T, lines=list(lty=c(1,2,3)), lwd = 2, xlab = 'verb baseline average', ylab = 'player post-test regularization', main = 'Effect of co-player lexical typicality on post-test responses')
 
@@
 <<main_model3, echo=F,eval=T, results=tex>>=
-load('models/si_fit1.rda')
+load('~/Github/RaczBecknerHayPierrehumbert2019/supplementary_information/models/si_fit1.rda')
 lme.fit1.sum = niceCoefTable(lme.fit1)

### 2020-10-15 supplementary_information/si.tex
na
proof check

@@
 
-% latex table generated in R 3.6.0 by xtable 1.8-4 package
-% Mon Apr 20 20:04:26 2020
+% latex table generated in R 4.0.2 by xtable 1.8-4 package
+% Thu Oct 15 15:19:32 2020
 \begin{table}[ht]
@@
 
-% latex table generated in R 3.6.0 by xtable 1.8-4 package
-% Mon Apr 20 20:04:26 2020
+% latex table generated in R 4.0.2 by xtable 1.8-4 package
+% Thu Oct 15 15:19:33 2020
 \begin{table}[ht]
@@
   \hline
-term & estimate & std.error & statistic & sig \\ 
+term & estimate & std.error & Z score & sig \\ 
   \hline
@@
 
-% latex table generated in R 3.6.0 by xtable 1.8-4 package
-% Mon Apr 20 20:04:27 2020
+% latex table generated in R 4.0.2 by xtable 1.8-4 package
+% Thu Oct 15 15:19:34 2020
 \begin{table}[ht]
@@
 \end{table}
-% latex table generated in R 3.6.0 by xtable 1.8-4 package
-% Mon Apr 20 20:04:27 2020
+% latex table generated in R 4.0.2 by xtable 1.8-4 package
+% Thu Oct 15 15:19:34 2020
 \begin{table}[ht]
@@
 \end{table}
-% latex table generated in R 3.6.0 by xtable 1.8-4 package
-% Mon Apr 20 20:04:27 2020
+% latex table generated in R 4.0.2 by xtable 1.8-4 package
+% Thu Oct 15 15:19:35 2020
 \begin{table}[ht]
@@
 
-% latex table generated in R 3.6.0 by xtable 1.8-4 package
-% Mon Apr 20 20:04:28 2020
+% latex table generated in R 4.0.2 by xtable 1.8-4 package
+% Thu Oct 15 15:19:35 2020
 \begin{table}[ht]
